### 2023-01-12 Manuscript_v1.tex
first commit

\documentclass[journal,twoside,web]{ieeecolor}
\usepackage{generic}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}

\usepackage{multirow}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\markboth{\journalname, VOL. XX, NO. XX, XXXX 2017}
{Author \MakeLowercase{\textit{et al.}}: Point-process based representation learning for Electronic Health Records}
\begin{document}
\title{Point-process based representation learning for Electronic Health Records}
\author{Hojjat Karami, \IEEEmembership{Fellow, IEEE}, Anisoara Ionescu, and David Atienza, \IEEEmembership{Member, IEEE}
\thanks{This paragraph of the first footnote will contain the date on 
which you submitted your paper for review. It will also contain support 
information, including sponsor and financial support acknowledgment. For 
example, ``This work was supported in part by the U.S. Department of 
Commerce under Grant BS123456.'' }
\thanks{The next few paragraphs should contain 
the authors' current affiliations, including current address and e-mail. For 
example, F. A. Author is with the National Institute of Standards and 
Technology, Boulder, CO 80305 USA (e-mail: author@boulder.nist.gov). }
\thanks{S. B. Author, Jr., was with Rice University, Houston, TX 77005 USA. He is 
now with the Department of Physics, Colorado State University, Fort Collins, 
CO 80523 USA (e-mail: author@lamar.colostate.edu).}
\thanks{T. C. Author is with 
the Electrical Engineering Department, University of Colorado, Boulder, CO 
80309 USA, on leave from the National Research Institute for Metals, 
Tsukuba, Japan (e-mail: author@nrim.go.jp).}}

\maketitle

\begin{abstract}
    It is very well accepted that missingness in electronic health records (EHRs) are not at random which is regarded as informative missingness. The clinician’s decision on when to observe lab tests over time can be modeled using point processes. We propose a novel framework based on neural point process to analyze laboratory tests of ICU patients. This framework can take into account additional information for better characterization of conditional intensity function (CIF) as well as better accuracy in prediction of future timestamp and labels.
\end{abstract}

\begin{IEEEkeywords}
Enter keywords or phrases in alphabetical 
order, separated by commas. For a list of suggested keywords, send a blank 
e-mail to keywords@ieee.org or visit the site.
\end{IEEEkeywords}

\section{Introduction}
\label{sec:intro}

Writing an effective abstract is an indispensable part of any form of research, since it that can motivate the audience to read follow the rest of the text. In this task, I have analyzed five abstracts from the field of artificial intelligence (AI). The table in section 2 shows a summary of the most important aspects of the analyzed abstracts.

\section{Background}
\label{sec:Background}

\textbf{Temporal point process}

\textbf{Neural temporal point process}

\textbf{Handling irregular sampling}

\section{Related Works}
\label{sec:Related Works}

\section{Proposed Model}
\label{sec:Proposed Model}

Event sequence data consists of N sequences $\{\mathcal{S}_{i}\}_{i=1}^N$, where
 each sequence $\mathcal{S}_{i}$ is a series of $L_i$ events $\{ (t_j, e_j)
 \}_{j=1}^{L_i}$. Here, $e_j$ represents events that could be independent our mutually 
 exclusive occurring at $t_i$.
 \newline
 In addition to the event data, we might have additional information
  $\{\mathcal{D}_{i}\}_{i=1}^N$. Suppose that each state is represented as $\{
  (t_k, v_k, m_k) \}_{k=1}^{P_i}$, consisting of a time value $t_k\in
  \mathbb{R}^{+} $, an observed value  $z_k\in \mathbb{R} $ and a modality
  indicator  $m_k\in \{1, ..., M\} $.

\subsection{Event Encoder}
We use a similar transformer architecture [thp] for encoding events. The
advantage of the attention mechanism is that it discards recurrent architecture.
it is necessary to include temporal information. Similar to the original
positional encoding [vaswani], we use a temporal encoding procedure:

\begin{equation}\label{eq:1}
    [z(t_j)]_{k} = 
     \begin{cases} 
        \sin\left(  \frac{t_j}{\mathcal{T}^{(k-1)/d_t} }  \right) & \text{if } k \text{ is odd}\\
        \sin\left(  \frac{t_j}{\mathcal{T}^{k/d_t} }  \right) & \text{if } k \text{ is even}

        \end{cases} 
\end{equation}

Here, $d_t \in \mathbb{N} $ is the dimensionality of encoded timestamp and $z
\in \mathbb{R}^{d_t} $ is the embeding vector of timestamp.

Each event mark $e_j$ is projected to a sparse binary vector representation. We
add an embedding layer to achieve a more compact and efficient representation
$emb$. Here, $w$ and $b$ are weights and biases of the embedding layer which can be
learned during network training.

While previous works suggested adding temporal encoding to the embedded events,
we propose to concatenate these two vectors. The effectiveness of this
concatenation is further investigated in the results.

\begin{equation} x_{emb}=[y(k_j), z(t_j)] \end{equation}

Now that the $x_{emb}$ is ready for encoding, it is encoded through a standard
transformer encoder with multiple layers and attention heads.

\begin{equation}
    x_{enc}=TE(x_{emb})    
\end{equation}

\subsection{State Encoder}
Here, we propose a method to incorporate additional information for a
better representation. In healthcare, much data is available from different
modalities such as vital signs and laboratory values.

Similar to [setF], we use an attention-based aggregation approach for encoding all additional information. Each side information  $ (t_k, v_k, m_k) $ can be represented by $s_k=(z(t_k), v_k, m_k)$. we define attention $a(\mathcal{S}_k,s_k )$

We define $\mathcal{S}_p$ to be the set of the first $p$ available information.
The goal is to calculate $a(\mathcal{S}_p,s_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{S}_p$.
This is achieved by computing an embedding of the set elements using a smaller set functions $f^{\prime}$, and projecting the concatenation of the set representation and the individual set element into d-dimensional space:

\begin{align*}
    f^{\prime}(\mathcal{S}_p) = g^{\prime} \left(  \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) \right)  \\
    K_p=[f^{\prime}(\mathcal{S}_p), s_p]^T W^K   
\end{align*}

Furthermore, we define a query vector $w^q \in \mathbb{R}^d$, which allows the model to summarize different aspects of the dataset via

\begin{equation*}
    e_p=\frac{K_p.w^q}{\sqrt{d}  }
\end{equation*}

Now, the desired attention can be computed as follows:

\begin{equation*}
    a(\mathcal{S}_p,s_k)=\frac{\exp (e_p)}{\sum_{k\leq p} \exp (e_k)  } 
\end{equation*}

Finally, we compute a weighted aggregation of set elements:

\begin{equation*}
    f(\mathcal{S}_p) =     
    g_{\psi}\left(
    \sum_{s_k \in \mathcal{S}_p}  a(\mathcal{S}_p,s_k)h_{\theta}(s_k) 
    \right) 
\end{equation*}

Without loss of generality, we can consider multiple heads by adding an additional dimension to keys and queries.\\

All formulas are:

\begin{equation*}
    \begin{cases} 
f^{\prime}(\mathcal{S}_p) = g^{\prime} \left(   
 \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) 
\right) \\
K_p=[f^{\prime}(\mathcal{S}_p), s_p]^T W^K \\
e_p=\frac{K_p.w^q}{\sqrt{d}  }\\
a(\mathcal{S}_p,s_k)=\frac{\exp (e_p)}{\sum_{k\leq p} \exp (e_k)  } \\
f(\mathcal{S}_p) = \sum_{s_k \in \mathcal{S}_p}  a(\mathcal{S}_p,s_k)h_{\theta}(s_k) \\
    z_p=g_{\psi}\left(         f(\mathcal{S}_p)         \right) \\
    \end{cases} 
\end{equation*}

\subsection{Mark and Time Decoder}

by concatenating encoded events $x_{enc}$ and additional information $f^{\prime}(\mathcal{S}_p)$, we can predict next marks and times as follows:

\begin{equation}
    \hat{e}_{j+1}=MLP([x_j, f^{\prime}(\mathcal{S}_j)])
\end{equation}

\begin{equation}
    \hat{t}_{j+1}=MLP([x_j, f^{\prime}(\mathcal{S}_j)])
\end{equation}

\begin{equation}
    \mathcal{L}_{mark} =
\end{equation}
\begin{equation}
    \mathcal{L}_{time} = \sum_{j=1}^{L-1} \left(  (\hat{t}_{j+1}-t_j)-(t_{j+1}-t_j)  \right)^2
\end{equation}

\subsection{Event Decoder}
Once we obtain a representation of a patient using embedded events and states, we can try to parameterize conditional intensity functions (CIFs) of the events.

In neural point process literature, many approaches have been propose to decode either conditional or cumulative intensity function.

\begin{equation}
    \lambda_k(t|\mathcal{H}_t ) = f_k\left(\alpha_k \frac{t-t_j}{t_j}+\mathbf{w}_k^T \mathbf{x}_{enc}(t_j)+\mathbf{y}_k^T \mathbf{s}_{enc}(t_j) +b_k   \right)
\end{equation}

\begin{equation}
    \lambda_k(t|\mathcal{H}_t ) = f_k\left(\alpha_k \frac{t-t_j}{t_j}+\mathbf{w}_k^T \mathbf{x}_{enc}(t_j)+b_k   \right)
\end{equation}

\section{Experiments}
\label{sec:Experiments}

\subsection*{Datasets}

\textbf{Physionet 2019 Sepsis Early Prediction Challenge (\emph{P19})}.
This dataset contains clinical data of about 40k patients in ICU. Clinical data consist of demographics, vital signs and laboratory values as well as sepsis label in a one-hour time grid. Our objective is to predict the timestamp of next lab sampling events as well as measured variables (event marks) given the patient history.

% To show the effectiveness of the proposed method, we further consider the following scenarios:
% \begin{itemize}
%     \item Enc(Event)+Dec(NextEvent): Here
%     \item Enc(Event)+Dec(NextEvent+CIF): Here
%     \item Enc(Event+state)+Dec(NextEvent): Here
%     \item Enc(Event+state)+Dec(NextEvent+CIF): Here
% \end{itemize}

\textbf{MIMIC-IV (\emph{M4})}.
We selected Medical Information Mart for Intensive Care (MIMIC) IV [1], which is a real-world clinical database comprising health data relating to over 40,000 patients admitted to ICU at the Beth Israel Deaconess Medical Center.

\textbf{Synthea (\emph{Syn})}.
We used the Synthea simulator (Walonoski et al., 2018) which generates patient-level EHRs using human expert curated Markov processes. Here, we reused the already processed version of this data by Edgauard.

\textbf{Stackoverflow (\emph{SO})}.
StackOverflow is a question-answering website. The website rewards users with badges to promote engagement in the community, and the same badge can be rewarded multiple times to the same user. We collect data in a two-year period, and we treat each user’s reward history as a sequence. Each event in the sequence signifies receipt of a particular medal.

\subsection*{Scenarios}
To investigate the effectiveness of the proposed method, we consider three input scenarios (TE, DA, TE+DA) as well as three loss functions (next event, CIF, next event+CIF) which would result in nine scenarios.

The first series of experiments are conducted to investigate the advantage of encoding additional information for paramterization of intensity functions. We consider seven scenarios:
Here, the baseline models

To show the effectiveness of time concatenation we report.

\subsection*{Baselines}
we use NEURALTPP that is already developed pipeline by as they already considered a lot of comibinations.

\subsection*{Metrics}
We report the weighted AUPRC, AUROC of next predicted event as well as root mean sqaure error (RMSE) of next measurement interval.
For evaluating the goodness of fit for the parameterized point process, we report normalized negative likelihood normalized by number of ocurred event (NLL/events). Furthermore, we can also evaluate the learned representation of each patient to predict the sepsis label in a binary classification task.

\section{Results and Discussion}
\label{sec:Results and Discussion}

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
% \usepackage{graphicx}
\begin{table}[h!]
    \centering
    \caption{unsupervised learning}
    \label{tab:table1}
    \resizebox{\columnwidth}{!}{%    
    \begin{tabular}{cclllll}
    \hline
    \multirow{2}{*}{Dataset} & \multicolumn{2}{c}{Model}            & \multicolumn{2}{c}{next mark} & next time & CIFs  \\
                             & \multicolumn{1}{l}{module} & loss    & AUPRC         & AUROC         & RMSE      & NLL   \\ \hline
    \multirow{6}{*}{p12}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
    \multirow{6}{*}{p19}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
    \end{tabular}%
    }
\end{table}

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
% \usepackage{graphicx}
\begin{table}[h!]

    \centering
    \caption{Transfer learnings}
    \label{tab:table2}
    \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|cll|ll|ll|}
    \hline
    \multicolumn{3}{|c|}{model}                                                          & \multicolumn{2}{c|}{p12}           & \multicolumn{2}{c|}{p19}           \\ \hline
    \multicolumn{1}{|l|}{module}                & \multicolumn{1}{l|}{loss}    & freezed & \multicolumn{1}{l|}{AUPRC} & AUROC & \multicolumn{1}{l|}{AUPRC} & AUROC \\ \hline
    \multicolumn{1}{|c|}{\multirow{3}{*}{TE}}   & \multicolumn{1}{l|}{MC}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{ML}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{SHPMARK} &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \hline
    \multicolumn{1}{|c|}{\multirow{3}{*}{TEDA}} & \multicolumn{1}{l|}{MC}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{ML}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{SHPMARK} &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \hline
    \end{tabular}%
    }
    \end{table}

\subsection{Next event prediction}

using CIF does not seem to improve the next mark and time prediction.
this is new insight as previous studies has claimed this loss to be useful.

\subsection{Learned representation}

using CIF loss function leads to a better representation for sepsis.

\subsection{Likelihood estimation}

Although CIF does not imporve mark prediction, it has led to better representation of patient for downstream task such as sepsis prediction.

In addition, we can interprete some of learned CIF patterns.

explain the effect of time concatenation in SO dataset

\subsection{Model interpretability}

one advantage of proposed method is use of attention mechanisms in both event and state encoder.

tsne of learned representation. 4 modes:

\begin{itemize}
    \item (DA,TE)->(Mark, CIF)
\end{itemize}

attention of DA for sepsis prediction

attention matrix of events for SO dataset

\section{Conclusion}
\label{sec:Conclusion}

\section{Introduction}
\label{sec:introduction}
\IEEEPARstart{T}{his} document is a template for \LaTeX. If you are 
reading a paper or PDF version of this document, please download the 
electronic file, trans\_jour.tex, from the IEEE Web site at \underline
{http://www.ieee.org/authortools/trans\_jour.tex} so you can use it to prepare your manuscript. If 
you would prefer to use LaTeX, download IEEE's LaTeX style and sample files 
from the same Web page. You can also explore using the Overleaf editor at 
\underline
{https://www.overleaf.com/blog/278-how-to-use-overleaf-with-}\discretionary{}{}{}\underline
{ieee-collabratec-your-quick-guide-to-getting-started\#.}\discretionary{}{}{}\underline{xsVp6tpPkrKM9}

If your paper is intended for a conference, please contact your conference 
editor concerning acceptable word processor formats for your particular 
conference.

IEEE will do the final formatting of your paper. If your paper is intended 
for a conference, please observe the conference page limits.

\subsection{Abbreviations and Acronyms}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have already been defined in the abstract. Abbreviations 
such as IEEE, SI, ac, and dc do not have to be defined. Abbreviations that 
incorporate periods should not have spaces: write ``C.N.R.S.,'' not ``C. N. 
R. S.'' Do not use abbreviations in the title unless they are unavoidable 
(for example, ``IEEE'' in the title of this article).

\subsection{Other Recommendations}
Use one space after periods and colons. Hyphenate complex modifiers: 
``zero-field-cooled magnetization.'' Avoid dangling participles, such as, 
``Using \eqref{eq}, the potential was calculated.'' [It is not clear who or what 
used \eqref{eq}.] Write instead, ``The potential was calculated by using \eqref{eq},'' or 
``Using \eqref{eq}, we calculated the potential.''

Use a zero before decimal points: ``0.25,'' not ``.25.'' Use 
``cm$^{3}$,'' not ``cc.'' Indicate sample dimensions as ``0.1 cm 
$\times $ 0.2 cm,'' not ``0.1 $\times $ 0.2 cm$^{2}$.'' The 
abbreviation for ``seconds'' is ``s,'' not ``sec.'' Use 
``Wb/m$^{2}$'' or ``webers per square meter,'' not 
``webers/m$^{2}$.'' When expressing a range of values, write ``7 to 
9'' or ``7--9,'' not ``7$\sim $9.''

A parenthetical statement at the end of a sentence is punctuated outside of 
the closing parenthesis (like this). (A parenthetical sentence is punctuated 
within the parentheses.) In American English, periods and commas are within 
quotation marks, like ``this period.'' Other punctuation is ``outside''! 
Avoid contractions; for example, write ``do not'' instead of ``don't.'' The 
serial comma is preferred: ``A, B, and C'' instead of ``A, B and C.''

If you wish, you may write in the first person singular or plural and use 
the active voice (``I observed that $\ldots$'' or ``We observed that $\ldots$'' 
instead of ``It was observed that $\ldots$''). Remember to check spelling. If 
your native language is not English, please get a native English-speaking 
colleague to carefully proofread your paper.

Try not to use too many typefaces in the same article. You're writing
scholarly papers, not ransom notes. Also please remember that MathJax
can't handle really weird typefaces.

\subsection{Equations}
Number equations consecutively with equation numbers in parentheses flush 
with the right margin, as in \eqref{eq}. To make your equations more 
compact, you may use the solidus (~/~), the exp function, or appropriate 
exponents. Use parentheses to avoid ambiguities in denominators. Punctuate 
equations when they are part of a sentence, as in
\begin{equation}E=mc^2.\label{eq}\end{equation}

Be sure that the symbols in your equation have been defined before the 
equation appears or immediately following. Italicize symbols ($T$ might refer 
to temperature, but T is the unit tesla). Refer to ``\eqref{eq},'' not ``Eq. \eqref{eq}'' 
or ``equation \eqref{eq},'' except at the beginning of a sentence: ``Equation \eqref{eq} 
is $\ldots$ .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

If you are submitting your paper to a colorized journal, you can use
the following two lines at the start of the article to ensure its
appearance resembles the final copy:

\smallskip\noindent
\begin{small}
\begin{tabular}{l}
\verb+\+\texttt{documentclass[journal,twoside,web]\{ieeecolor\}}\\
\verb+\+\texttt{usepackage\{\textit{Journal\_Name}\}}
\end{tabular}
\end{small}

\section{Units}
Use either SI (MKS) or CGS as primary units. (SI units are strongly 
encouraged.) English units may be used as secondary units (in parentheses). 
This applies to papers in data storage. For example, write ``15 
Gb/cm$^{2}$ (100 Gb/in$^{2})$.'' An exception is when 
English units are used as identifiers in trade, such as ``3\textonehalf-in 
disk drive.'' Avoid combining SI and CGS units, such as current in amperes 
and magnetic field in oersteds. This often leads to confusion because 
equations do not balance dimensionally. If you must use mixed units, clearly 
state the units for each quantity in an equation.

The SI unit for magnetic field strength $H$ is A/m. However, if you wish to use 
units of T, either refer to magnetic flux density $B$ or magnetic field 
strength symbolized as $\mu _{0}H$. Use the center dot to separate 
compound units, e.g., ``A$\cdot $m$^{2}$.''

\section{Some Common Mistakes}
The word ``data'' is plural, not singular. The subscript for the 
permeability of vacuum $\mu _{0}$ is zero, not a lowercase letter 
``o.'' The term for residual magnetization is ``remanence''; the adjective 
is ``remanent''; do not write ``remnance'' or ``remnant.'' Use the word 
``micrometer'' instead of ``micron.'' A graph within a graph is an 
``inset,'' not an ``insert.'' The word ``alternatively'' is preferred to the 
word ``alternately'' (unless you really mean something that alternates). Use 
the word ``whereas'' instead of ``while'' (unless you are referring to 
simultaneous events). Do not use the word ``essentially'' to mean 
``approximately'' or ``effectively.'' Do not use the word ``issue'' as a 
euphemism for ``problem.'' When compositions are not specified, separate 
chemical symbols by en-dashes; for example, ``NiMn'' indicates the 
intermetallic compound Ni$_{0.5}$Mn$_{0.5}$ whereas 
``Ni--Mn'' indicates an alloy of some composition 
Ni$_{x}$Mn$_{1-x}$.

\begin{figure}[!t]
\centerline{\includegraphics[width=\columnwidth]{fig1.png}}
\caption{Magnetization as a function of applied field.
It is good practice to explain the significance of the figure in the caption.}
\label{fig1}
\end{figure}

Be aware of the different meanings of the homophones ``affect'' (usually a 
verb) and ``effect'' (usually a noun), ``complement'' and ``compliment,'' 
``discreet'' and ``discrete,'' ``principal'' (e.g., ``principal 
investigator'') and ``principle'' (e.g., ``principle of measurement''). Do 
not confuse ``imply'' and ``infer.''

Prefixes such as ``non,'' ``sub,'' ``micro,'' ``multi,'' and ``ultra'' are 
not independent words; they should be joined to the words they modify, 
usually without a hyphen. There is no period after the ``et'' in the Latin 
abbreviation ``\emph{et al.}'' (it is also italicized). The abbreviation ``i.e.,'' means 
``that is,'' and the abbreviation ``e.g.,'' means ``for example'' (these 
abbreviations are not italicized).

A general IEEE styleguide is available at \underline{http://www.ieee.org/authortools}.

\section{Guidelines for Graphics Preparation and Submission}
\label{sec:guidelines}

\subsection{Types of Graphics}
The following list outlines the different types of graphics published in 
IEEE journals. They are categorized based on their construction, and use of 
color/shades of gray:

\subsubsection{Color/Grayscale figures}
{Figures that are meant to appear in color, or shades of black/gray. Such 
figures may include photographs, illustrations, multicolor graphs, and 
flowcharts.}

\subsubsection{Line Art figures}
{Figures that are composed of only black lines and shapes. These figures 
should have no shades or half-tones of gray, only black and white.}

\subsubsection{Author photos}
{Head and shoulders shots of authors that appear at the end of our papers. }

\subsubsection{Tables}
{Data charts which are typically black and white, but sometimes include 
color.}

\begin{table}
\caption{Units for Magnetic Properties}
\label{table}
\setlength{\tabcolsep}{3pt}
\begin{tabular}{|p{25pt}|p{75pt}|p{115pt}|}
\hline
Symbol& 
Quantity& 
Conversion from Gaussian and \par CGS EMU to SI $^{\mathrm{a}}$ \\
\hline
$\Phi $& 
magnetic flux& 
1 Mx $\to  10^{-8}$ Wb $= 10^{-8}$ V$\cdot $s \\
$B$& 
magnetic flux density, \par magnetic induction& 
1 G $\to  10^{-4}$ T $= 10^{-4}$ Wb/m$^{2}$ \\
$H$& 
magnetic field strength& 
1 Oe $\to  10^{3}/(4\pi )$ A/m \\
$m$& 
magnetic moment& 
1 erg/G $=$ 1 emu \par $\to 10^{-3}$ A$\cdot $m$^{2} = 10^{-3}$ J/T \\
$M$& 
magnetization& 
1 erg/(G$\cdot $cm$^{3}) =$ 1 emu/cm$^{3}$ \par $\to 10^{3}$ A/m \\
4$\pi M$& 
magnetization& 
1 G $\to  10^{3}/(4\pi )$ A/m \\
$\sigma $& 
specific magnetization& 
1 erg/(G$\cdot $g) $=$ 1 emu/g $\to $ 1 A$\cdot $m$^{2}$/kg \\
$j$& 
magnetic dipole \par moment& 
1 erg/G $=$ 1 emu \par $\to 4\pi \times  10^{-10}$ Wb$\cdot $m \\
$J$& 
magnetic polarization& 
1 erg/(G$\cdot $cm$^{3}) =$ 1 emu/cm$^{3}$ \par $\to 4\pi \times  10^{-4}$ T \\
$\chi , \kappa $& 
susceptibility& 
1 $\to  4\pi $ \\
$\chi_{\rho }$& 
mass susceptibility& 
1 cm$^{3}$/g $\to  4\pi \times  10^{-3}$ m$^{3}$/kg \\
$\mu $& 
permeability& 
1 $\to  4\pi \times  10^{-7}$ H/m \par $= 4\pi \times  10^{-7}$ Wb/(A$\cdot $m) \\
$\mu_{r}$& 
relative permeability& 
$\mu \to \mu_{r}$ \\
$w, W$& 
energy density& 
1 erg/cm$^{3} \to  10^{-1}$ J/m$^{3}$ \\
$N, D$& 
demagnetizing factor& 
1 $\to  1/(4\pi )$ \\
\hline
\multicolumn{3}{p{251pt}}{Vertical lines are optional in tables. Statements that serve as captions for 
the entire table do not need footnote letters. }\\
\multicolumn{3}{p{251pt}}{$^{\mathrm{a}}$Gaussian units are the same as cg emu for magnetostatics; Mx 
$=$ maxwell, G $=$ gauss, Oe $=$ oersted; Wb $=$ weber, V $=$ volt, s $=$ 
second, T $=$ tesla, m $=$ meter, A $=$ ampere, J $=$ joule, kg $=$ 
kilogram, H $=$ henry.}
\end{tabular}
\label{tab1}
\end{table}

\subsection{Multipart figures}
Figures compiled of more than one sub-figure presented side-by-side, or 
stacked. If a multipart figure is made up of multiple figure
types (one part is lineart, and another is grayscale or color) the figure 
should meet the stricter guidelines.

\subsection{File Formats For Graphics}\label{formats}
Format and save your graphics using a suitable graphics processing program 
that will allow you to create the images as PostScript (PS), Encapsulated 
PostScript (.EPS), Tagged Image File Format (.TIFF), Portable Document 
Format (.PDF), Portable Network Graphics (.PNG), or Metapost (.MPS), sizes them, and adjusts 
the resolution settings. When 
submitting your final paper, your graphics should all be submitted 
individually in one of these formats along with the manuscript.

\subsection{Sizing of Graphics}
Most charts, graphs, and tables are one column wide (3.5 inches/88 
millimeters/21 picas) or page wide (7.16 inches/181 millimeters/43 
picas). The maximum depth a graphic can be is 8.5 inches (216 millimeters/54
picas). When choosing the depth of a graphic, please allow space for a 
caption. Figures can be sized between column and page widths if the author 
chooses, however it is recommended that figures are not sized less than 
column width unless when necessary.

There is currently one publication with column measurements that do not 
coincide with those listed above. Proceedings of the IEEE has a column 
measurement of 3.25 inches (82.5 millimeters/19.5 picas).

The final printed size of author photographs is exactly
1 inch wide by 1.25 inches tall (25.4 millimeters$\,\times\,$31.75 millimeters/6 
picas$\,\times\,$7.5 picas). Author photos printed in editorials measure 1.59 inches 
wide by 2 inches tall (40 millimeters$\,\times\,$50 millimeters/9.5 picas$\,\times\,$12 
picas).

\subsection{Resolution }
The proper resolution of your figures will depend on the type of figure it 
is as defined in the ``Types of Figures'' section. Author photographs, 
color, and grayscale figures should be at least 300dpi. Line art, including 
tables should be a minimum of 600dpi.

\subsection{Vector Art}
In order to preserve the figures' integrity across multiple computer 
platforms, we accept files in the following formats: .EPS/.PDF/.PS. All 
fonts must be embedded or text converted to outlines in order to achieve the 
best-quality results.

\subsection{Color Space}
The term color space refers to the entire sum of colors that can be 
represented within the said medium. For our purposes, the three main color 
spaces are Grayscale, RGB (red/green/blue) and CMYK 
(cyan/magenta/yellow/black). RGB is generally used with on-screen graphics, 
whereas CMYK is used for printing purposes.

All color figures should be generated in RGB or CMYK color space. Grayscale 
images should be submitted in Grayscale color space. Line art may be 
provided in grayscale OR bitmap colorspace. Note that ``bitmap colorspace'' 
and ``bitmap file format'' are not the same thing. When bitmap color space 
is selected, .TIF/.TIFF/.PNG are the recommended file formats.

\subsection{Accepted Fonts Within Figures}
When preparing your graphics IEEE suggests that you use of one of the 
following Open Type fonts: Times New Roman, Helvetica, Arial, Cambria, and 
Symbol. If you are supplying EPS, PS, or PDF files all fonts must be 
embedded. Some fonts may only be native to your operating system; without 
the fonts embedded, parts of the graphic may be distorted or missing.

A safe option when finalizing your figures is to strip out the fonts before 
you save the files, creating ``outline'' type. This converts fonts to 
artwork what will appear uniformly on any screen.

\subsection{Using Labels Within Figures}

\subsubsection{Figure Axis labels }
Figure axis labels are often a source of confusion. Use words rather than 
symbols. As an example, write the quantity ``Magnetization,'' or 
``Magnetization M,'' not just ``M.'' Put units in parentheses. Do not label 
axes only with units. As in Fig. 1, for example, write ``Magnetization 
(A/m)'' or ``Magnetization (A$\cdot$m$^{-1}$),'' not just ``A/m.'' Do not label axes with a ratio of quantities and 
units. For example, write ``Temperature (K),'' not ``Temperature/K.''

Multipliers can be especially confusing. Write ``Magnetization (kA/m)'' or 
``Magnetization (10$^{3}$ A/m).'' Do not write ``Magnetization 
(A/m)$\,\times\,$1000'' because the reader would not know whether the top 
axis label in Fig. 1 meant 16000 A/m or 0.016 A/m. Figure labels should be 
legible, approximately 8 to 10 point type.

\subsubsection{Subfigure Labels in Multipart Figures and Tables}
Multipart figures should be combined and labeled before final submission. 
Labels should appear centered below each subfigure in 8 point Times New 
Roman font in the format of (a) (b) (c).

\subsection{File Naming}
Figures (line artwork or photographs) should be named starting with the 
first 5 letters of the author's last name. The next characters in the 
filename should be the number that represents the sequential 
location of this image in your article. For example, in author 
``Anderson's'' paper, the first three figures would be named ander1.tif, 
ander2.tif, and ander3.ps.

Tables should contain only the body of the table (not the caption) and 
should be named similarly to figures, except that `.t' is inserted 
in-between the author's name and the table number. For example, author 
Anderson's first three tables would be named ander.t1.tif, ander.t2.ps, 
ander.t3.eps.

Author photographs should be named using the first five characters of the 
pictured author's last name. For example, four author photographs for a 
paper may be named: oppen.ps, moshc.tif, chen.eps, and duran.pdf.

If two authors or more have the same last name, their first initial(s) can 
be substituted for the fifth, fourth, third$\ldots$ letters of their surname 
until the degree where there is differentiation. For example, two authors 
Michael and Monica Oppenheimer's photos would be named oppmi.tif, and 
oppmo.eps.

\subsection{Referencing a Figure or Table Within Your Paper}
When referencing your figures and tables within your paper, use the 
abbreviation ``Fig.'' even at the beginning of a sentence. Do not abbreviate 
``Table.'' Tables should be numbered with Roman Numerals.

\subsection{Checking Your Figures: The IEEE Graphics Analyzer}
The IEEE Graphics Analyzer enables authors to pre-screen their graphics for 
compliance with IEEE Transactions and Journals standards before submission. 
The online tool, located at
\underline{http://graphicsqc.ieee.org/}, allows authors to 
upload their graphics in order to check that each file is the correct file 
format, resolution, size and colorspace; that no fonts are missing or 
corrupt; that figures are not compiled in layers or have transparency, and 
that they are named according to the IEEE Transactions and Journals naming 
convention. At the end of this automated process, authors are provided with 
a detailed report on each graphic within the web applet, as well as by 
email.

For more information on using the Graphics Analyzer or any other graphics 
related topic, contact the IEEE Graphics Help Desk by e-mail at 
graphics@ieee.org.

\subsection{Submitting Your Graphics}
Because IEEE will do the final formatting of your paper,
you do not need to position figures and tables at the top and bottom of each 
column. In fact, all figures, figure captions, and tables can be placed at 
the end of your paper. In addition to, or even in lieu of submitting figures 
within your final manuscript, figures should be submitted individually, 
separate from the manuscript in one of the file formats listed above in 
Section \ref{formats}. Place figure captions below the figures; place table titles 
above the tables. Please do not include captions as part of the figures, or 
put them in ``text boxes'' linked to the figures. Also, do not place borders 
around the outside of your figures.

\subsection{Color Processing/Printing in IEEE Journals}
All IEEE Transactions, Journals, and Letters allow an author to publish 
color figures on IEEE Xplore\textregistered\ at no charge, and automatically 
convert them to grayscale for print versions. In most journals, figures and 
tables may alternatively be printed in color if an author chooses to do so. 
Please note that this service comes at an extra expense to the author. If 
you intend to have print color graphics, include a note with your final 
paper indicating which figures or tables you would like to be handled that 
way, and stating that you are willing to pay the additional fee.

\section{Conclusion}
A conclusion section is not required. Although a conclusion may review the 
main points of the paper, do not replicate the abstract as the conclusion. A 
conclusion might elaborate on the importance of the work or suggest 
applications and extensions.

\appendices

Appendixes, if needed, appear before the acknowledgment.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in American English is 
without an ``e'' after the ``g.'' Use the singular heading even if you have 
many acknowledgments. Avoid expressions such as ``One of us (S.B.A.) would 
like to thank $\ldots$ .'' Instead, write ``F. A. Author thanks $\ldots$ .'' In most 
cases, sponsor and financial support acknowledgments are placed in the 
unnumbered footnote on the first page, not here.

\section*{References and Footnotes}

\subsection{References}
References need not be cited in text. When they are, they appear on the 
line, in square brackets, inside the punctuation. Multiple references are 
each numbered with separate brackets. When citing a section in a book, 
please give the relevant page numbers. In text, refer simply to the 
reference number. Do not use ``Ref.'' or ``reference'' except at the 
beginning of a sentence: ``Reference \cite{b3} shows $\ldots$ .'' Please do not use 
automatic endnotes in \emph{Word}, rather, type the reference list at the end of the 
paper using the ``References'' style.

Reference numbers are set flush left and form a column of their own, hanging 
out beyond the body of the reference. The reference numbers are on the line, 
enclosed in square brackets. In all references, the given name of the author 
or editor is abbreviated to the initial only and precedes the last name. Use 
them all; use \emph{et al.} only if names are not given. Use commas around Jr., 
Sr., and III in names. Abbreviate conference titles. When citing IEEE 
transactions, provide the issue number, page range, volume number, year, 
and/or month if available. When referencing a patent, provide the day and 
the month of issue, or application. References may not include all 
information; please obtain and include relevant information. Do not combine 
references. There must be only one reference with each number. If there is a 
URL included with the print reference, it can be included at the end of the 
reference.

Other than books, capitalize only the first word in a paper title, except 
for proper nouns and element symbols. For papers published in translation 
journals, please give the English citation first, followed by the original 
foreign-language citation See the end of this document for formats and 
examples of common references. For a complete discussion of references and 
their formats, see the IEEE style manual at
\underline{http://www.ieee.org/authortools}.

\subsection{Footnotes}
Number footnotes separately in superscript numbers.\footnote{It is recommended that footnotes be avoided (except for 
the unnumbered footnote with the receipt date on the first page). Instead, 
try to integrate the footnote information into the text.} Place the actual 
footnote at the bottom of the column in which it is cited; do not put 
footnotes in the reference list (endnotes). Use letters for table footnotes 
(see Table \ref{table}).

\section{Submitting Your Paper for Review}

\subsection{Final Stage}
When you submit your final version (after your paper has been accepted), 
print it in two-column format, including figures and tables. You must also 
send your final manuscript on a disk, via e-mail, or through a Web 
manuscript submission system as directed by the society contact. You may use 
\emph{Zip} for large files, or compress files using \emph{Compress, Pkzip, Stuffit,} or \emph{Gzip.}

Also, send a sheet of paper or PDF with complete contact information for all 
authors. Include full mailing addresses, telephone numbers, fax numbers, and 
e-mail addresses. This information will be used to send each author a 
complimentary copy of the journal in which the paper appears. In addition, 
designate one author as the ``corresponding author.'' This is the author to 
whom proofs of the paper will be sent. Proofs are sent to the corresponding 
author only.

\subsection{Review Stage Using ScholarOne\textregistered\ Manuscripts}
Contributions to the Transactions, Journals, and Letters may be submitted 
electronically on IEEE's on-line manuscript submission and peer-review 
system, ScholarOne\textregistered\ Manuscripts. You can get a listing of the 
publications that participate in ScholarOne at 
\underline{http://www.ieee.org/publications\_standards/publications/}\discretionary{}{}{}\underline{authors/authors\_submission.html}.
First check if you have an existing account. If there is none, please create 
a new account. After logging in, go to your Author Center and click ``Submit 
First Draft of a New Manuscript.''

Along with other information, you will be asked to select the subject from a 
pull-down list. Depending on the journal, there are various steps to the 
submission process; you must complete all steps for a complete submission. 
At the end of each step you must click ``Save and Continue''; just uploading 
the paper is not sufficient. After the last step, you should see a 
confirmation that the submission is complete. You should also receive an 
e-mail confirmation. For inquiries regarding the submission of your paper on 
ScholarOne Manuscripts, please contact oprs-support@ieee.org or call +1 732 
465 5861.

ScholarOne Manuscripts will accept files for review in various formats. 
Please check the guidelines of the specific journal for which you plan to 
submit.

You will be asked to file an electronic copyright form immediately upon 
completing the submission process (authors are responsible for obtaining any 
security clearances). Failure to submit the electronic copyright could 
result in publishing delays later. You will also have the opportunity to 
designate your article as ``open access'' if you agree to pay the IEEE open 
access fee.

\subsection{Final Stage Using ScholarOne Manuscripts}
Upon acceptance, you will receive an email with specific instructions 
regarding the submission of your final files. To avoid any delays in 
publication, please be sure to follow these instructions. Most journals 
require that final submissions be uploaded through ScholarOne Manuscripts, 
although some may still accept final submissions via email. Final 
submissions should include source files of your accepted manuscript, high 
quality graphic files, and a formatted pdf file. If you have any questions 
regarding the final submission process, please contact the administrative 
contact for the journal.

In addition to this, upload a file with complete contact information for all 
authors. Include full mailing addresses, telephone numbers, fax numbers, and 
e-mail addresses. Designate the author who submitted the manuscript on 
ScholarOne Manuscripts as the ``corresponding author.'' This is the only 
author to whom proofs of the paper will be sent.

\subsection{Copyright Form}
Authors must submit an electronic IEEE Copyright Form (eCF) upon submitting 
their final manuscript files. You can access the eCF system through your 
manuscript submission system or through the Author Gateway. You are 
responsible for obtaining any necessary approvals and/or security 
clearances. For additional information on intellectual property rights, 
visit the IEEE Intellectual Property Rights department web page at 
\underline{http://www.ieee.org/publications\_standards/publications/rights/}\discretionary{}{}{}\underline{index.html}.

\section{IEEE Publishing Policy}
The general IEEE policy requires that authors should only submit original 
work that has neither appeared elsewhere for publication, nor is under 
review for another refereed publication. The submitting author must disclose 
all prior publication(s) and current submissions when submitting a 
manuscript. Do not publish ``preliminary'' data or results. The submitting 
author is responsible for obtaining agreement of all coauthors and any 
consent required from employers or sponsors before submitting an article. 
The IEEE Transactions and Journals Department strongly discourages courtesy 
authorship; it is the obligation of the authors to cite only relevant prior 
work.

The IEEE Transactions and Journals Department does not publish conference 
records or proceedings, but can publish articles related to conferences that 
have undergone rigorous peer review. Minimally, two reviews are required for 
every article submitted for peer review.

\section{Publication Principles}
The two types of contents of that are published are; 1) peer-reviewed and 2) 
archival. The Transactions and Journals Department publishes scholarly 
articles of archival value as well as tutorial expositions and critical 
reviews of classical subjects and topics of current interest.

Authors should consider the following points:

\begin{enumerate}
\item Technical papers submitted for publication must advance the state of knowledge and must cite relevant prior work. 
\item The length of a submitted paper should be commensurate with the importance, or appropriate to the complexity, of the work. For example, an obvious extension of previously published work might not be appropriate for publication or might be adequately treated in just a few pages.
\item Authors must convince both peer reviewers and the editors of the scientific and technical merit of a paper; the standards of proof are higher when extraordinary or unexpected results are reported. 
\item Because replication is required for scientific progress, papers submitted for publication must provide sufficient information to allow readers to perform similar experiments or calculations and 
use the reported results. Although not everything need be disclosed, a paper 
must contain new, useable, and fully described information. For example, a 
specimen's chemical composition need not be reported if the main purpose of 
a paper is to introduce a new measurement technique. Authors should expect 
to be challenged by reviewers if the results are not supported by adequate 
data and critical details.
\item Papers that describe ongoing work or announce the latest technical achievement, which are suitable for presentation at a professional conference, may not be appropriate for publication.
\end{enumerate}

\section{Reference Examples}

\begin{itemize}

\item \emph{Basic format for books:}\\
J. K. Author, ``Title of chapter in the book,'' in \emph{Title of His Published Book, x}th ed. City of Publisher, (only U.S. State), Country: Abbrev. of Publisher, year, ch. $x$, sec. $x$, pp. \emph{xxx--xxx.}\\
See \cite{b1,b2}.

\item \emph{Basic format for periodicals:}\\
J. K. Author, ``Name of paper,'' \emph{Abbrev. Title of Periodical}, vol. \emph{x, no}. $x, $pp\emph{. xxx--xxx, }Abbrev. Month, year, DOI. 10.1109.\emph{XXX}.123456.\\
See \cite{b3}--\cite{b5}.

\item \emph{Basic format for reports:}\\
J. K. Author, ``Title of report,'' Abbrev. Name of Co., City of Co., Abbrev. State, Country, Rep. \emph{xxx}, year.\\
See \cite{b6,b7}.

\item \emph{Basic format for handbooks:}\\
\emph{Name of Manual/Handbook, x} ed., Abbrev. Name of Co., City of Co., Abbrev. State, Country, year, pp. \emph{xxx--xxx.}\\
See \cite{b8,b9}.

\item \emph{Basic format for books (when available online):}\\
J. K. Author, ``Title of chapter in the book,'' in \emph{Title of
Published Book}, $x$th ed. City of Publisher, State, Country: Abbrev.
of Publisher, year, ch. $x$, sec. $x$, pp. \emph{xxx--xxx}. [Online].
Available: \underline{http://www.web.com}\\
See \cite{b10}--\cite{b13}.

\item \emph{Basic format for journals (when available online):}\\
J. K. Author, ``Name of paper,'' \emph{Abbrev. Title of Periodical}, vol. $x$, no. $x$, pp. \emph{xxx--xxx}, Abbrev. Month, year. Accessed on: Month, Day, year, DOI: 10.1109.\emph{XXX}.123456, [Online].\\
See \cite{b14}--\cite{b16}.

\item \emph{Basic format for papers presented at conferences (when available online): }\\
J.K. Author. (year, month). Title. presented at abbrev. conference title. [Type of Medium]. Available: site/path/file\\
See \cite{b17}.

\item \emph{Basic format for reports and handbooks (when available online):}\\
J. K. Author. ``Title of report,'' Company. City, State, Country. Rep. no., (optional: vol./issue), Date. [Online] Available: site/path/file\\
See \cite{b18,b19}.

\item \emph{Basic format for computer programs and electronic documents (when available online): }\\
Legislative body. Number of Congress, Session. (year, month day). \emph{Number of bill or resolution}, \emph{Title}. [Type of medium]. Available: site/path/file\\
\textbf{\emph{NOTE: }ISO recommends that capitalization follow the accepted practice for the language or script in which the information is given.}\\
See \cite{b20}.

\item \emph{Basic format for patents (when available online):}\\
Name of the invention, by inventor's name. (year, month day). Patent Number [Type of medium]. Available: site/path/file\\
See \cite{b21}.

\item \emph{Basic format}\emph{for conference proceedings (published):}\\
J. K. Author, ``Title of paper,'' in \emph{Abbreviated Name of Conf.}, City of Conf., Abbrev. State (if given), Country, year, pp. \emph{xxxxxx.}\\
See \cite{b22}.

\item \emph{Example for papers presented at conferences (unpublished):}\\
See \cite{b23}.

\item \emph{Basic format for patents}$:$\\
J. K. Author, ``Title of patent,'' U.S. Patent \emph{x xxx xxx}, Abbrev. Month, day, year.\\
See \cite{b24}.

\item \emph{Basic format for theses (M.S.) and dissertations (Ph.D.):}
\begin{enumerate}
\item J. K. Author, ``Title of thesis,'' M.S. thesis, Abbrev. Dept., Abbrev. Univ., City of Univ., Abbrev. State, year.
\item J. K. Author, ``Title of dissertation,'' Ph.D. dissertation, Abbrev. Dept., Abbrev. Univ., City of Univ., Abbrev. State, year.
\end{enumerate}
See \cite{b25,b26}.

\item \emph{Basic format for the most common types of unpublished references:}
\begin{enumerate}
\item J. K. Author, private communication, Abbrev. Month, year.
\item J. K. Author, ``Title of paper,'' unpublished.
\item J. K. Author, ``Title of paper,'' to be published.
\end{enumerate}
See \cite{b27}--\cite{b29}.

\item \emph{Basic formats for standards:}
\begin{enumerate}
\item \emph{Title of Standard}, Standard number, date.
\item \emph{Title of Standard}, Standard number, Corporate author, location, date.
\end{enumerate}
See \cite{b30,b31}.

\item \emph{Article number in~reference examples:}\\
See \cite{b32,b33}.

\item \emph{Example when using et al.:}\\
See \cite{b34}.

\end{itemize}

\begin{thebibliography}{00}

\bibitem{b1} G. O. Young, ``Synthetic structure of industrial plastics,'' in \emph{Plastics,} 2\textsuperscript{nd} ed., vol. 3, J. Peters, Ed. New York, NY, USA: McGraw-Hill, 1964, pp. 15--64.

\bibitem{b2} W.-K. Chen, \emph{Linear Networks and Systems.} Belmont, CA, USA: Wadsworth, 1993, pp. 123--135.

\bibitem{b3} J. U. Duncombe, ``Infrared navigation---Part I: An assessment of feasibility,'' \emph{IEEE Trans. Electron Devices}, vol. ED-11, no. 1, pp. 34--39, Jan. 1959, 10.1109/TED.2016.2628402.

\bibitem{b4} E. P. Wigner, ``Theory of traveling-wave optical laser,'' \emph{Phys. Rev}., vol. 134, pp. A635--A646, Dec. 1965.

\bibitem{b5} E. H. Miller, ``A note on reflector arrays,'' \emph{IEEE Trans. Antennas Propagat}., to be published.

\bibitem{b6} E. E. Reber, R. L. Michell, and C. J. Carter, ``Oxygen absorption in the earth's atmosphere,'' Aerospace Corp., Los Angeles, CA, USA, Tech. Rep. TR-0200 (4230-46)-3, Nov. 1988.

\bibitem{b7} J. H. Davis and J. R. Cogdell, ``Calibration program for the 16-foot antenna,'' Elect. Eng. Res. Lab., Univ. Texas, Austin, TX, USA, Tech. Memo. NGL-006-69-3, Nov. 15, 1987.

\bibitem{b8} \emph{Transmission Systems for Communications}, 3\textsuperscript{rd} ed., Western Electric Co., Winston-Salem, NC, USA, 1985, pp. 44--60.

\bibitem{b9} \emph{Motorola Semiconductor Data Manual}, Motorola Semiconductor Products Inc., Phoenix, AZ, USA, 1989.

\bibitem{b10} G. O. Young, ``Synthetic structure of industrial
plastics,'' in Plastics, vol. 3, Polymers of Hexadromicon, J. Peters,
Ed., 2\textsuperscript{nd} ed. New York, NY, USA: McGraw-Hill, 1964, pp. 15-64.
[Online]. Available:
\underline{http://www.bookref.com}.

\bibitem{b11} \emph{The Founders' Constitution}, Philip B. Kurland
and Ralph Lerner, eds., Chicago, IL, USA: Univ. Chicago Press, 1987.
[Online]. Available: \underline{http://press-pubs.uchicago.edu/founders/}

\bibitem{b12} The Terahertz Wave eBook. ZOmega Terahertz Corp., 2014.
[Online]. Available:
\underline{http://dl.z-thz.com/eBook/zomega\_ebook\_pdf\_1206\_sr.pdf}. Accessed on: May 19, 2014.

\bibitem{b13} Philip B. Kurland and Ralph Lerner, eds., \emph{The
Founders' Constitution.} Chicago, IL, USA: Univ. of Chicago Press,
1987, Accessed on: Feb. 28, 2010, [Online] Available:
\underline{http://press-pubs.uchicago.edu/founders/}

\bibitem{b14} J. S. Turner, ``New directions in communications,'' \emph{IEEE J. Sel. Areas Commun}., vol. 13, no. 1, pp. 11-23, Jan. 1995.

\bibitem{b15} W. P. Risk, G. S. Kino, and H. J. Shaw, ``Fiber-optic frequency shifter using a surface acoustic wave incident at an oblique angle,'' \emph{Opt. Lett.}, vol. 11, no. 2, pp. 115--117, Feb. 1986.

\bibitem{b16} P. Kopyt \emph{et al., ``}Electric properties of graphene-based conductive layers from DC up to terahertz range,'' \emph{IEEE THz Sci. Technol.,} to be published. DOI: 10.1109/TTHZ.2016.2544142.

\bibitem{b17} PROCESS Corporation, Boston, MA, USA. Intranets:
Internet technologies deployed behind the firewall for corporate
productivity. Presented at INET96 Annual Meeting. [Online].
Available: \underline{http://home.process.com/Intranets/wp2.htp}

\bibitem{b18} R. J. Hijmans and J. van Etten, ``Raster: Geographic analysis and modeling with raster data,'' R Package Version 2.0-12, Jan. 12, 2012. [Online]. Available: \underline {http://CRAN.R-project.org/package=raster}

\bibitem{b19} Teralyzer. Lytera UG, Kirchhain, Germany [Online].
Available:
\underline{http://www.lytera.de/Terahertz\_THz\_Spectroscopy.php?id=home}, Accessed on: Jun. 5, 2014

\bibitem{b20} U.S. House. 102\textsuperscript{nd} Congress, 1\textsuperscript{st} Session. (1991, Jan. 11). \emph{H. Con. Res. 1, Sense of the Congress on Approval of}  \emph{Military Action}. [Online]. Available: LEXIS Library: GENFED File: BILLS

\bibitem{b21} Musical toothbrush with mirror, by L.M.R. Brooks. (1992, May 19). Patent D 326 189 [Online]. Available: NEXIS Library: LEXPAT File: DES

\bibitem{b22} D. B. Payne and J. R. Stern, ``Wavelength-switched pas- sively coupled single-mode optical network,'' in \emph{Proc. IOOC-ECOC,} Boston, MA, USA, 1985, pp. 585--590.

\bibitem{b23} D. Ebehard and E. Voges, ``Digital single sideband detection for interferometric sensors,'' presented at the \emph{2\textsuperscript{nd} Int. Conf. Optical Fiber Sensors,} Stuttgart, Germany, Jan. 2-5, 1984.

\bibitem{b24} G. Brandli and M. Dick, ``Alternating current fed power supply,'' U.S. Patent 4 084 217, Nov. 4, 1978.

\bibitem{b25} J. O. Williams, ``Narrow-band analyzer,'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, USA, 1993.

\bibitem{b26} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.

\bibitem{b27} A. Harrison, private communication, May 1995.

\bibitem{b28} B. Smith, ``An approach to graphs of linear forms,'' unpublished.

\bibitem{b29} A. Brahms, ``Representation error for real numbers in binary computer arithmetic,'' IEEE Computer Group Repository, Paper R-67-85.

\bibitem{b30} IEEE Criteria for Class IE Electric Systems, IEEE Standard 308, 1969.

\bibitem{b31} Letter Symbols for Quantities, ANSI Standard Y10.5-1968.

\bibitem{b32} R. Fardel, M. Nagel, F. Nuesch, T. Lippert, and A. Wokaun, ``Fabrication of organic light emitting diode pixels by laser-assisted forward transfer,'' \emph{Appl. Phys. Lett.}, vol. 91, no. 6, Aug. 2007, Art. no. 061103.~

\bibitem{b33} J. Zhang and N. Tansu, ``Optical gain and laser characteristics of InGaN quantum wells on ternary InGaN substrates,'' \emph{IEEE Photon. J.}, vol. 5, no. 2, Apr. 2013, Art. no. 2600111

\bibitem{b34} S. Azodolmolky~\emph{et al.}, Experimental demonstration of an impairment aware network planning and operation tool for transparent/translucent optical networks,''~\emph{J. Lightw. Technol.}, vol. 29, no. 4, pp. 439--448, Sep. 2011.

\end{thebibliography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a1.png}}]{First A. Author} (M'76--SM'81--F'87) and all authors may include 
biographies. Biographies are often not included in conference-related
papers. This author became a Member (M) of IEEE in 1976, a Senior
Member (SM) in 1981, and a Fellow (F) in 1987. The first paragraph may
contain a place and/or date of birth (list place, then date). Next,
the author's educational background is listed. The degrees should be
listed with type of degree in what field, which institution, city,
state, and country, and year the degree was earned. The author's major
field of study should be lower-cased.

The second paragraph uses the pronoun of the person (he or she) and not the 
author's last name. It lists military and work experience, including summer 
and fellowship jobs. Job titles are capitalized. The current job must have a 
location; previous positions may be listed 
without one. Information concerning previous publications may be included. 
Try not to list more than three books or published articles. The format for 
listing publishers of a book within the biography is: title of book 
(publisher name, year) similar to a reference. Current and previous research 
interests end the paragraph. The third paragraph begins with the author's 
title and last name (e.g., Dr.\ Smith, Prof.\ Jones, Mr.\ Kajor, Ms.\ Hunter). 
List any memberships in professional societies other than the IEEE. Finally, 
list any awards and work for IEEE committees and publications. If a 
photograph is provided, it should be of good quality, and 
professional-looking. Following are two examples of an author's biography.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a2.png}}]{Second B. Author} was born in Greenwich Village, New York, NY, USA in 
1977. He received the B.S. and M.S. degrees in aerospace engineering from 
the University of Virginia, Charlottesville, in 2001 and the Ph.D. degree in 
mechanical engineering from Drexel University, Philadelphia, PA, in 2008.

From 2001 to 2004, he was a Research Assistant with the Princeton Plasma 
Physics Laboratory. Since 2009, he has been an Assistant Professor with the 
Mechanical Engineering Department, Texas A{\&}M University, College Station. 
He is the author of three books, more than 150 articles, and more than 70 
inventions. His research interests include high-pressure and high-density 
nonthermal plasma discharge processes and applications, microscale plasma 
discharges, discharges in liquids, spectroscopic diagnostics, plasma 
propulsion, and innovation plasma applications. He is an Associate Editor of 
the journal \emph{Earth, Moon, Planets}, and holds two patents.

Dr. Author was a recipient of the International Association of Geomagnetism 
and Aeronomy Young Scientist Award for Excellence in 2008, and the IEEE 
Electromagnetic Compatibility Society Best Symposium Paper Award in 2011. 
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a3.png}}]{Third C. Author, Jr.} (M'87) received the B.S. degree in mechanical 
engineering from National Chung Cheng University, Chiayi, Taiwan, in 2004 
and the M.S. degree in mechanical engineering from National Tsing Hua 
University, Hsinchu, Taiwan, in 2006. He is currently pursuing the Ph.D. 
degree in mechanical engineering at Texas A{\&}M University, College 
Station, TX, USA.

From 2008 to 2009, he was a Research Assistant with the Institute of 
Physics, Academia Sinica, Tapei, Taiwan. His research interest includes the 
development of surface processing and biological/medical treatment 
techniques using nonthermal atmospheric pressure plasmas, fundamental study 
of plasma sources, and fabrication of micro- or nanostructured surfaces.

Mr. Author's awards and honors include the Frew Fellowship (Australian 
Academy of Science), the I. I. Rabi Prize (APS), the European Frequency and 
Time Forum Award, the Carl Zeiss Research Award, the William F. Meggers 
Award and the Adolph Lomb Medal (OSA).
\end{IEEEbiography}

\end{document}

### 2023-01-12 generic-color-brief.tex
first commit

--- 2023-01-12 Manuscript_v1.tex
@@
-\documentclass[journal,twoside,web]{ieeecolor}
+\documentclass[9pt,shortpaper,twoside,web]{ieeecolor}
 \usepackage{generic}
@@
 
-\usepackage{multirow}
+\usepackage{amsmath}
 
@@
 \markboth{\journalname, VOL. XX, NO. XX, XXXX 2017}
-{Author \MakeLowercase{\textit{et al.}}: Point-process based representation learning for Electronic Health Records}
+{Author \MakeLowercase{\textit{et al.}}: Neural temporal point process for
+modeling laboratory measurements in ICU}
 \begin{document}
-\title{Point-process based representation learning for Electronic Health Records}
-\author{Hojjat Karami, \IEEEmembership{Fellow, IEEE}, Anisoara Ionescu, and David Atienza, \IEEEmembership{Member, IEEE}
-\thanks{This paragraph of the first footnote will contain the date on 
-which you submitted your paper for review. It will also contain support 
-information, including sponsor and financial support acknowledgment. For 
-example, ``This work was supported in part by the U.S. Department of 
-Commerce under Grant BS123456.'' }
-\thanks{The next few paragraphs should contain 
-the authors' current affiliations, including current address and e-mail. For 
-example, F. A. Author is with the National Institute of Standards and 
-Technology, Boulder, CO 80305 USA (e-mail: author@boulder.nist.gov). }
-\thanks{S. B. Author, Jr., was with Rice University, Houston, TX 77005 USA. He is 
-now with the Department of Physics, Colorado State University, Fort Collins, 
-CO 80523 USA (e-mail: author@lamar.colostate.edu).}
-\thanks{T. C. Author is with 
-the Electrical Engineering Department, University of Colorado, Boulder, CO 
-80309 USA, on leave from the National Research Institute for Metals, 
-Tsukuba, Japan (e-mail: author@nrim.go.jp).}}
+\title{Neural temporal point process for modeling laboratory measurements in ICU}
+\author{H. Karami, \IEEEmembership{Fellow, IEEE}, Second B. Author, and Third C.
+Author, Jr., \IEEEmembership{Member, IEEE} 
+\thanks{This paragraph of the first
+footnote will contain the date on which you submitted your paper for review. It
+will also contain support information, including sponsor and financial support
+acknowledgment. For example, ``This work was supported in part by the U.S.
+Department of Commerce under Grant BS123456.'' } \thanks{The next few paragraphs
+should contain the authors' current affiliations, including current address and
+e-mail. For example, F. A. Author is with the National Institute of Standards
+and Technology, Boulder, CO 80305 USA (e-mail: author@boulder.nist.gov). }
+\thanks{S. B. Author, Jr., was with Rice University, Houston, TX 77005 USA. He
+is now with the Department of Physics, Colorado State University, Fort Collins,
+CO 80523 USA (e-mail: author@lamar.colostate.edu).} \thanks{T. C. Author is with
+the Electrical Engineering Department, University of Colorado, Boulder, CO 80309
+USA, on leave from the National Research Institute for Metals, Tsukuba, Japan
+(e-mail: author@nrim.go.jp).}}
 
@@
 \begin{abstract}
-    It is very well accepted that missingness in electronic health records (EHRs) are not at random which is regarded as informative missingness.
-The clinician’s decision on when to observe lab tests over time can be modeled using point processes.
-We propose a novel framework based on neural point process to analyze laboratory tests of ICU patients.
-This framework can take into account additional information for better characterization of conditional intensity function (CIF) as well as better accuracy in prediction of future timestamp and labels.
+    It is very well accepted that missingness in electronic health records
+    (EHRs) are not at random which is regarded as informative missingness. The
+    clinician’s decision on when to observe lab tests over time can be modeled
+    using point processes. We propose a novel framework based on neural point
+    process to analyze laboratory tests of ICU patients. This framework can take
+    into account additional information for better characterization of
+    conditional intensity function (CIF) as well as better accuracy in
+    prediction of future timestamp and labels.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Enter keywords or phrases in alphabetical 
-order, separated by commas. For a list of suggested keywords, send a blank 
-e-mail to keywords@ieee.org or visit the site.
+point process, Electronic Health Records (EHR),
 \end{IEEEkeywords}
@@
 \section{Introduction}
-\label{sec:intro}
-
-Writing an effective abstract is an indispensable part of any form of research, since it that can motivate the audience to read follow the rest of the text.
-In this task, I have analyzed five abstracts from the field of artificial intelligence (AI).
-The table in section 2 shows a summary of the most important aspects of the analyzed abstracts.
+\label{sec:introduction}
 
@@
 \label{sec:Background}
-
-\textbf{Temporal point process}
-
-\textbf{Neural temporal point process}
-
-\textbf{Handling irregular sampling}
 
@@
 \begin{align*}
-    f^{\prime}(\mathcal{S}_p) = g^{\prime} \left(  \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) \right)  \\
+    f^{\prime}(a\mathcal{S}_p) = g^{\prime} \left(  \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) \right)  \\
     K_p=[f^{\prime}(\mathcal{S}_p), s_p]^T W^K   
@@
 
-Furthermore, we define a query vector $w^q \in \mathbb{R}^d$, which allows the model to summarize different aspects of the dataset via
+Furthermore, we define a query vector $w^q \in \mathbb{R}^d$, which allow the model to summarize different aspects of the dataset via
 
@@
 \end{equation*}
-
-Without loss of generality, we can consider multiple heads by adding an additional dimension to keys and queries.\\
 
@@
 
-\subsection{Mark and Time Decoder}
-
-by concatenating encoded events $x_{enc}$ and additional information $f^{\prime}(\mathcal{S}_p)$, we can predict next marks and times as follows:
-
-\begin{equation}
-    \hat{e}_{j+1}=MLP([x_j, f^{\prime}(\mathcal{S}_j)])
-\end{equation}
-
-\begin{equation}
-    \hat{t}_{j+1}=MLP([x_j, f^{\prime}(\mathcal{S}_j)])
-\end{equation}
-
-\begin{equation}
-    \mathcal{L}_{mark} =
-\end{equation}
-\begin{equation}
-    \mathcal{L}_{time} = \sum_{j=1}^{L-1} \left(  (\hat{t}_{j+1}-t_j)-(t_{j+1}-t_j)  \right)^2
-\end{equation}
-
-\subsection{Event Decoder}
-Once we obtain a representation of a patient using embedded events and states, we can try to parameterize conditional intensity functions (CIFs) of the events.
-
-In neural point process literature, many approaches have been propose to decode either conditional or cumulative intensity function.
-
-\begin{equation}
-    \lambda_k(t|\mathcal{H}_t ) = f_k\left(\alpha_k \frac{t-t_j}{t_j}+\mathbf{w}_k^T \mathbf{x}_{enc}(t_j)+\mathbf{y}_k^T \mathbf{s}_{enc}(t_j) +b_k   \right)
-\end{equation}
-
-\begin{equation}
-    \lambda_k(t|\mathcal{H}_t ) = f_k\left(\alpha_k \frac{t-t_j}{t_j}+\mathbf{w}_k^T \mathbf{x}_{enc}(t_j)+b_k   \right)
-\end{equation}
-
 \section{Experiments}
@@
 
-\textbf{Physionet 2019 Sepsis Early Prediction Challenge (\emph{P19})}.
-This dataset contains clinical data of about 40k patients in ICU.
+\textbf{Physionet 2019 Sepsis Early Prediction Challenge }.
+This is a dataset contains clinical data of about 40k patients in ICU.
 Clinical data consist of demographics, vital signs and laboratory values as well as sepsis label in a one-hour time grid.
 Our objective is to predict the timestamp of next lab sampling events as well as measured variables (event marks) given the patient history.
-
-% To show the effectiveness of the proposed method, we further consider the following scenarios:
-% \begin{itemize}
-%     \item Enc(Event)+Dec(NextEvent): Here
-%     \item Enc(Event)+Dec(NextEvent+CIF): Here
-%     \item Enc(Event+state)+Dec(NextEvent): Here
-%     \item Enc(Event+state)+Dec(NextEvent+CIF): Here
-% \end{itemize}
-
-\textbf{MIMIC-IV (\emph{M4})}.
-We selected Medical Information Mart for Intensive Care (MIMIC) IV [1], which is a real-world clinical database comprising health data relating to over 40,000 patients admitted to ICU at the Beth Israel Deaconess Medical Center.
-
-\textbf{Synthea (\emph{Syn})}.
-We used the Synthea simulator (Walonoski et al., 2018) which generates patient-level EHRs using human expert curated Markov processes.
-Here, we reused the already processed version of this data by Edgauard.
-
-\textbf{Stackoverflow (\emph{SO})}.
-StackOverflow is a question-answering website.
-The website rewards users with badges to promote engagement in the community, and the same badge can be rewarded multiple times to the same user.
-We collect data in a two-year period, and we treat each user’s reward history as a sequence.
-Each event in the sequence signifies receipt of a particular medal.
-
-\subsection*{Scenarios}
-To investigate the effectiveness of the proposed method, we consider three input scenarios (TE, DA, TE+DA) as well as three loss functions (next event, CIF, next event+CIF) which would result in nine scenarios.
-
-The first series of experiments are conducted to investigate the advantage of encoding additional information for paramterization of intensity functions. We consider seven scenarios:
-Here, the baseline models
-
-To show the effectiveness of time concatenation we report.
+fsdfs.
 
 \subsection*{Baselines}
-we use NEURALTPP that is already developed pipeline by as they already considered a lot of comibinations.
-
-\subsection*{Metrics}
-We report the weighted AUPRC, AUROC of next predicted event as well as root mean sqaure error (RMSE) of next measurement interval.
-For evaluating the goodness of fit for the parameterized point process, we report normalized negative likelihood normalized by number of ocurred event (NLL/events).
-Furthermore, we can also evaluate the learned representation of each patient to predict the sepsis label in a binary classification task.
 
 \section{Results and Discussion}
+
 \label{sec:Results and Discussion}
-
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-% \usepackage{graphicx}
-\begin{table}[h!]
-    \centering
-    \caption{unsupervised learning}
-    \label{tab:table1}
-    \resizebox{\columnwidth}{!}{%    
-    \begin{tabular}{cclllll}
-    \hline
-    \multirow{2}{*}{Dataset} & \multicolumn{2}{c}{Model}            & \multicolumn{2}{c}{next mark} & next time & CIFs  \\
-                             & \multicolumn{1}{l}{module} & loss    & AUPRC         & AUROC         & RMSE      & NLL   \\ \hline
-    \multirow{6}{*}{p12}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
-    \multirow{6}{*}{p19}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
-    \end{tabular}%
-    }
-\end{table}
-
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-% \usepackage{graphicx}
-\begin{table}[h!]
-
-    \centering
-    \caption{Transfer learnings}
-    \label{tab:table2}
-    \resizebox{\columnwidth}{!}{%
-    \begin{tabular}{|cll|ll|ll|}
-    \hline
-    \multicolumn{3}{|c|}{model}                                                          & \multicolumn{2}{c|}{p12}           & \multicolumn{2}{c|}{p19}           \\ \hline
-    \multicolumn{1}{|l|}{module}                & \multicolumn{1}{l|}{loss}    & freezed & \multicolumn{1}{l|}{AUPRC} & AUROC & \multicolumn{1}{l|}{AUPRC} & AUROC \\ \hline
-    \multicolumn{1}{|c|}{\multirow{3}{*}{TE}}   & \multicolumn{1}{l|}{MC}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
-    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{ML}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
-    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{SHPMARK} &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \hline
-    \multicolumn{1}{|c|}{\multirow{3}{*}{TEDA}} & \multicolumn{1}{l|}{MC}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
-    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{ML}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
-    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{SHPMARK} &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \hline
-    \end{tabular}%
-    }
-    \end{table}
-
-\subsection{Next event prediction}
-
-using CIF does not seem to improve the next mark and time prediction.
-this is new insight as previous studies has claimed this loss to be useful.
-
-\subsection{Learned representation}
-
-using CIF loss function leads to a better representation for sepsis.
-
-\subsection{Likelihood estimation}
-
-Although CIF does not imporve mark prediction, it has led to better representation of patient for downstream task such as sepsis prediction.
-
-In addition, we can interprete some of learned CIF patterns.
-
-explain the effect of time concatenation in SO dataset
-
-\subsection{Model interpretability}
-
-one advantage of proposed method is use of attention mechanisms in both event and state encoder.
-
-tsne of learned representation. 4 modes:
-
-\begin{itemize}
-    \item (DA,TE)->(Mark, CIF)
-\end{itemize}
-
-attention of DA for sepsis prediction
-
-attention matrix of events for SO dataset
 
@@
 
-\section{Introduction}
-\label{sec:introduction}
-\IEEEPARstart{T}{his} document is a template for \LaTeX. If you are 
+\begin{equation}
+
+\frac{1}{n} \newline
+\frac{d^2 y}{d x^2} \bigotimes \cos \mathcal{L} \mathbb{L} \hat{\rightarrow \Rightarrow } 
+\Rightarrow \Longrightarrow 
+\end{equation}
+
+This document is a template for \LaTeX. If you are 
 reading a paper or PDF version of this document, please download the 
@@
 \subsection{Multipart figures}
-Figures compiled of more than one sub-figure presented side-by-side, or 
+Figures compiled of more than one sub-figure are presented side-by-side, or 
 stacked. If a multipart figure is made up of multiple figure
@@
 Contributions to the Transactions, Journals, and Letters may be submitted 
-electronically on IEEE's on-line manuscript submission and peer-review 
+electronically on IEEE's online manuscript submission and peer-review 
 system, ScholarOne\textregistered\ Manuscripts. You can get a listing of the 
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a1.png}}]{First A. Author} (M'76--SM'81--F'87) and all authors may include 
-biographies. Biographies are often not included in conference-related
-papers. This author became a Member (M) of IEEE in 1976, a Senior
-Member (SM) in 1981, and a Fellow (F) in 1987. The first paragraph may
-contain a place and/or date of birth (list place, then date). Next,
-the author's educational background is listed. The degrees should be
-listed with type of degree in what field, which institution, city,
-state, and country, and year the degree was earned. The author's major
-field of study should be lower-cased.
-
-The second paragraph uses the pronoun of the person (he or she) and not the 
-author's last name. It lists military and work experience, including summer 
-and fellowship jobs. Job titles are capitalized. The current job must have a 
-location; previous positions may be listed 
-without one. Information concerning previous publications may be included. 
-Try not to list more than three books or published articles. The format for 
-listing publishers of a book within the biography is: title of book 
-(publisher name, year) similar to a reference. Current and previous research 
-interests end the paragraph. The third paragraph begins with the author's 
-title and last name (e.g., Dr.\ Smith, Prof.\ Jones, Mr.\ Kajor, Ms.\ Hunter). 
-List any memberships in professional societies other than the IEEE. Finally, 
-list any awards and work for IEEE committees and publications. If a 
-photograph is provided, it should be of good quality, and 
-professional-looking. Following are two examples of an author's biography.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a2.png}}]{Second B. Author} was born in Greenwich Village, New York, NY, USA in 
-1977. He received the B.S. and M.S. degrees in aerospace engineering from 
-the University of Virginia, Charlottesville, in 2001 and the Ph.D. degree in 
-mechanical engineering from Drexel University, Philadelphia, PA, in 2008.
-
-From 2001 to 2004, he was a Research Assistant with the Princeton Plasma 
-Physics Laboratory. Since 2009, he has been an Assistant Professor with the 
-Mechanical Engineering Department, Texas A{\&}M University, College Station. 
-He is the author of three books, more than 150 articles, and more than 70 
-inventions. His research interests include high-pressure and high-density 
-nonthermal plasma discharge processes and applications, microscale plasma 
-discharges, discharges in liquids, spectroscopic diagnostics, plasma 
-propulsion, and innovation plasma applications. He is an Associate Editor of 
-the journal \emph{Earth, Moon, Planets}, and holds two patents.
-
-Dr. Author was a recipient of the International Association of Geomagnetism 
-and Aeronomy Young Scientist Award for Excellence in 2008, and the IEEE 
-Electromagnetic Compatibility Society Best Symposium Paper Award in 2011. 
-\end{IEEEbiography}
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a3.png}}]{Third C. Author, Jr.} (M'87) received the B.S. degree in mechanical 
-engineering from National Chung Cheng University, Chiayi, Taiwan, in 2004 
-and the M.S. degree in mechanical engineering from National Tsing Hua 
-University, Hsinchu, Taiwan, in 2006. He is currently pursuing the Ph.D. 
-degree in mechanical engineering at Texas A{\&}M University, College 
-Station, TX, USA.
-
-From 2008 to 2009, he was a Research Assistant with the Institute of 
-Physics, Academia Sinica, Tapei, Taiwan. His research interest includes the 
-development of surface processing and biological/medical treatment 
-techniques using nonthermal atmospheric pressure plasmas, fundamental study 
-of plasma sources, and fabrication of micro- or nanostructured surfaces.
-
-Mr. Author's awards and honors include the Frew Fellowship (Australian 
-Academy of Science), the I. I. Rabi Prize (APS), the European Frequency and 
-Time Forum Award, the Carl Zeiss Research Award, the William F. Meggers 
-Award and the Adolph Lomb Medal (OSA).
-\end{IEEEbiography}
-
 \end{document}

### 2023-01-26 Manuscript_v1.tex
commit
com

@@
 
-\textbf{Synthea (\emph{Syn})}.
+\textbf{Synthea(\emph{Syn})}.
 We used the Synthea simulator (Walonoski et al., 2018) which generates patient-level EHRs using human expert curated Markov processes.
@@
 
+In this section, we present our results regarding the advantage of state and event encoding.
+
+\subsection{Negative Likelihood with state encoding}
+
+Table 1 shows the result for estimation of negative likelihood in different datasets and scnearios. It is obvious that state encoding has led to lower NLL.
+
+\subsection{Time concatenation}
+
+\subsection{Downstream task with event encoding}
+
+Another key element of our work is to show the effectiveness of point process modeling for a down-stream task.
+In Table, we have reported the performance metrics for the mortality prediction task.
+We have compared our results with several sota's DL models that are compatible with irregular time series.
+
+\subsection{Learned representions}
+Fig 1 visualizes the tsne plot for the two scenarios.
+
+\subsection{Model interpretability}
+
+one advantage of proposed method is use of attention mechanisms in both event and state encoder.
+Fig 1 shows the attention mechanism
 % Please add the following required packages to your document preamble:
@@
 
-\subsection{Next event prediction}
-
-using CIF does not seem to improve the next mark and time prediction.
-this is new insight as previous studies has claimed this loss to be useful.
-
-\subsection{Learned representation}
-
-using CIF loss function leads to a better representation for sepsis.
-
 \subsection{Likelihood estimation}
 
-Although CIF does not imporve mark prediction, it has led to better representation of patient for downstream task such as sepsis prediction.
-
-In addition, we can interprete some of learned CIF patterns.
+Although CIF does not improve mark prediction, it has led to better representation of patient for downstream task such as sepsis prediction.
+
+In addition, we can interpret some of learned CIF patterns.
 
 explain the effect of time concatenation in SO dataset
-
-\subsection{Model interpretability}
-
-one advantage of proposed method is use of attention mechanisms in both event and state encoder.
 

### 2023-01-30 Manuscript_v1.tex
commit

@@
 
+\subsection{Effect of minor improvements}
+
+effect of time concatenation
+
+compare single+mark with mc or ml
+
+\begin{table}[h!]
+\centering
+\caption{Effect}
+\label{tab:my-table}
+\begin{tabular}{lllll}
+        &        & shp+mark &     & baseline \\ \cline{3-5} 
+Dataset & Metric & concat   & sum &          \\ \hline
+SO      & NLL    &          &     &          \\
+        & AUROC  &          &     &          \\
+        & time   &          &     &          \\ \hline
+Synthea & NLL    &          &     &          \\
+        & AUROC  &          &     &          \\
+        & time   &          &     &          \\ \hline
+ReTweet & NLL    &          &     &          \\
+        & AUROC  &          &     &          \\
+        & time   &          &     &          \\ \hline
+\end{tabular}
+\end{table}
+
 \subsection{Negative Likelihood with state encoding}
@@
 
-\subsection{Time concatenation}
+\begin{table}[h!]
+\centering
+\caption{NLL estimation}
+\label{tab:my-table}
+\begin{tabular}{lllll}
+\hline
+        &         &    & Model  &          \\ \cline{3-5} 
+Dataset & setting & TE & TE+DAM & TE+noise \\ \hline
+P12     & sc      &    &        &          \\
+        & mc1     &    &        &          \\
+        & mc2     &    &        &          \\ \hline
+p19     & sc      &    &        &          \\
+        & mc1     &    &        &          \\
+        & mc2     &    &        &          \\ \hline
+\end{tabular}
+\end{table}
+
+can u provide one example patient?
 
@@
 
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table*}
+\centering
+\caption{Down stream task}
+\label{tab:my-table}
+\begin{tabular}{cccccccccccc}
+\hline
+                     &                      &        &  & F1   &        &  & AUPRC &        &  & AUROC &        \\ \cline{5-6} \cline{8-9} \cline{11-12} 
+Dataset              & Setting              & Center &  & DAM  & TE+DAM &  & DAM   & TE+DAM &  & DAM   & TE+DAM \\ \hline
+\multirow{8}{*}{P12} & \multirow{3}{*}{sc}  & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
+                     & \multirow{3}{*}{mc1} & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
+                     & mc2                  & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     & seft                 & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \hline
+\multirow{8}{*}{P19} & \multirow{3}{*}{sc}  & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
+                     & \multirow{3}{*}{mc1} & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
+                     & mc2                  & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
+                     & seft                 & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \hline
+\end{tabular}
+\end{table*}
+
 \subsection{Learned representions}
@@
 Fig 1 shows the attention mechanism
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-% \usepackage{graphicx}
-\begin{table}[h!]
-    \centering
-    \caption{unsupervised learning}
-    \label{tab:table1}
-    \resizebox{\columnwidth}{!}{%    
-    \begin{tabular}{cclllll}
-    \hline
-    \multirow{2}{*}{Dataset} & \multicolumn{2}{c}{Model}            & \multicolumn{2}{c}{next mark} & next time & CIFs  \\
-                             & \multicolumn{1}{l}{module} & loss    & AUPRC         & AUROC         & RMSE      & NLL   \\ \hline
-    \multirow{6}{*}{p12}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
-    \multirow{6}{*}{p19}     & \multirow{3}{*}{TE}        & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             & \multirow{3}{*}{TEDA}      & MC      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & ML      & 23.98         & 23.98         & 23.98     & 23.98 \\
-                             &                            & SHPMARK & 23.98         & 23.98         & 23.98     & 23.98 \\ \hline
-    \end{tabular}%
-    }
-\end{table}
-
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-% \usepackage{graphicx}
-\begin{table}[h!]
-
-    \centering
-    \caption{Transfer learnings}
-    \label{tab:table2}
-    \resizebox{\columnwidth}{!}{%
-    \begin{tabular}{|cll|ll|ll|}
-    \hline
-    \multicolumn{3}{|c|}{model}                                                          & \multicolumn{2}{c|}{p12}           & \multicolumn{2}{c|}{p19}           \\ \hline
-    \multicolumn{1}{|l|}{module}                & \multicolumn{1}{l|}{loss}    & freezed & \multicolumn{1}{l|}{AUPRC} & AUROC & \multicolumn{1}{l|}{AUPRC} & AUROC \\ \hline
-    \multicolumn{1}{|c|}{\multirow{3}{*}{TE}}   & \multicolumn{1}{l|}{MC}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
-    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{ML}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
-    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{SHPMARK} &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \hline
-    \multicolumn{1}{|c|}{\multirow{3}{*}{TEDA}} & \multicolumn{1}{l|}{MC}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
-    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{ML}      &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \cline{2-7} 
-    \multicolumn{1}{|c|}{}                      & \multicolumn{1}{l|}{SHPMARK} &         & \multicolumn{1}{l|}{23.98} & 23.98 & \multicolumn{1}{l|}{23.98} & 23.98 \\ \hline
-    \end{tabular}%
-    }
-    \end{table}
 

### 2023-02-02 Manuscript_v1.tex
commit

@@
 
+\usepackage{booktabs}
+\usepackage{xcolor}
+\usepackage{colortbl}
+\usepackage{rotating}
 \usepackage{multirow}
@@
 
+\begin{figure*}[!t]
+\centerline{\includegraphics{images/model_arch.png}}
+\caption{Magnetization as a function of applied field.
+It is good practice to explain the significance of the figure in the caption.}
+\label{fig1}
+\end{figure*}
+
+The entire data consists of $N$ samples  $\mathcal{D}=\{\mathcal{O}_{i}=(\mathcal{E}_{i},\mathcal{S}_{i}  )\}_{i=1}^N$ where $\mathcal{E}_{i}$ and $\mathcal{S}_{i}$ represents event and state sequences, respectively.
+
+Each event sequnce consist of $L$ tuples $\mathcal{E}_{i}=\{ (t_j, e_j)
+\}_{j=1}^{L_i}$ where $t_j\in \mathbb{R}  $ is the event timestamp and $ e_j $ is the event mark.
+
 Event sequence data consists of N sequences $\{\mathcal{S}_{i}\}_{i=1}^N$, where
@@
 
+General schematic of TEDAM is depicted in FIG which consists of two separate module: Transformer Event Encoder (TEE) and Deep Attention Module (DAM).
+TEE encodes timestamp and mark of events using a transformer architecture, while DAM encodes all available information using one-dimensional attention mechanism.
+Finally, the learned representations are concatenated to be optimized for different loss functions.
+
 \subsection{Event Encoder}
-We use a similar transformer architecture [thp] for encoding events. The
-advantage of the attention mechanism is that it discards recurrent architecture.
+We use a similar transformer architecture [thp] for encoding events with minor modifications. A transformer architecture is capable of .
+
+let . that can be embedded using a trainable embedding matrix.
+
+In the first step, we embed all event marks $E_{emb}=E \times U$ where $E_i \in [L,K] $ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{K \times d_{emb} }  $ is the trainable embedding matrix.
+In the second step, timestamps should be encoded and added to the event embedding, however, we propose to concatenate time encodings that can lead to better characterization of conditional intensity functions.
+Finally, the input of the transformer encoder will be $x_{emb}=[y(k_j), z(t_j)]$.
+
+Here, we use the standard transformer encoder similar to [vaswani] with masking matrix to prevent the model from looking into the future.
+we obtain the encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $ where $h_j$ containes the all available information before occurence of $j-th$ event.
+
+\subsection{bin of Event Encoder}
+
 it is necessary to include temporal information. Similar to the original
@@
 
+% Table generated by Excel2LaTeX from sheet 'time concat'
 \begin{table}[h!]
     \centering
-\caption{Effect}
-\label{tab:my-table}
-\begin{tabular}{lllll}
-        &        & shp+mark &     & baseline \\ \cline{3-5} 
-Dataset & Metric & concat   & sum &          \\ \hline
-SO      & NLL    &          &     &          \\
-        & AUROC  &          &     &          \\
-        & time   &          &     &          \\ \hline
-Synthea & NLL    &          &     &          \\
-        & AUROC  &          &     &          \\
-        & time   &          &     &          \\ \hline
-ReTweet & NLL    &          &     &          \\
-        & AUROC  &          &     &          \\
-        & time   &          &     &          \\ \hline
-\end{tabular}
-\end{table}
+    \caption{Add caption}
+      \begin{tabular}{ccccc}
+      \toprule
+            &       & \multicolumn{2}{c}{shp+mark} &  \\
+  \cmidrule{3-4}    Dataset & Metric & concat & sum   & baseline \\
+      \midrule
+      SO    & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \midrule
+      Synthea & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \midrule
+      ReTweet & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \bottomrule
+      \end{tabular}%
+    \label{tab:addlabel}%
+  \end{table}%
 
@@
 
+% Table generated by Excel2LaTeX from sheet 'state enc'
 \begin{table}[h!]
     \centering
-\caption{NLL estimation}
-\label{tab:my-table}
-\begin{tabular}{lllll}
-\hline
-        &         &    & Model  &          \\ \cline{3-5} 
-Dataset & setting & TE & TE+DAM & TE+noise \\ \hline
-P12     & sc      &    &        &          \\
-        & mc1     &    &        &          \\
-        & mc2     &    &        &          \\ \hline
-p19     & sc      &    &        &          \\
-        & mc1     &    &        &          \\
-        & mc2     &    &        &          \\ \hline
-\end{tabular}
-\end{table}
+    \caption{Add caption}
+      \begin{tabular}{ccccc}
+      \toprule
+            &       & \multicolumn{3}{c}{Model} \\
+  \cmidrule{3-5}    Dataset & setting & TE    & TE+DAM & TE+noise \\
+      \midrule
+      \multirow{3}[2]{*}{P12} & sc    & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & mc1   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & mc2   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \midrule
+      \multirow{3}[2]{*}{p19} & sc    & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & mc1   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & mc2   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \bottomrule
+      \end{tabular}%
+    \label{tab:addlabel}%
+  \end{table}%
 
@@
 
-% Please add the following required packages to your document preamble:
-% \usepackage{multirow}
-\begin{table*}
+% Table generated by Excel2LaTeX from sheet 'Sheet3'
+\begin{table*}[htbp]
     \centering
-\caption{Down stream task}
-\label{tab:my-table}
-\begin{tabular}{cccccccccccc}
-\hline
-                     &                      &        &  & F1   &        &  & AUPRC &        &  & AUROC &        \\ \cline{5-6} \cline{8-9} \cline{11-12} 
-Dataset              & Setting              & Center &  & DAM  & TE+DAM &  & DAM   & TE+DAM &  & DAM   & TE+DAM \\ \hline
-\multirow{8}{*}{P12} & \multirow{3}{*}{sc}  & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
-                     & \multirow{3}{*}{mc1} & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
-                     & mc2                  & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     & seft                 & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \hline
-\multirow{8}{*}{P19} & \multirow{3}{*}{sc}  & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
-                     & \multirow{3}{*}{mc1} & 1      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     &                      & 2      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     &                      & 3      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \cline{2-12} 
-                     & mc2                  & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\
-                     & seft                 & -      &  & 0.55 & 0.55   &  & 0.55  & 0.55   &  & 0.55  & 0.55   \\ \hline
-\end{tabular}
-\end{table*}
+    \caption{Add caption}
+      \begin{tabular}{ccccccccccc}
+      \toprule
+            &       &       & \multicolumn{2}{c}{F1} &       & \multicolumn{2}{c}{AUPRC} &       & \multicolumn{2}{c}{AUROC} \\
+  \cmidrule{4-5}\cmidrule{7-8}\cmidrule{10-11}    Dataset & Setting & Center & DAM   & TE+DAM &       & DAM   & TE+DAM &       & DAM   & TE+DAM \\
+      \midrule
+      \multirow{8}[6]{*}{P12} & \multirow{3}[2]{*}{sc} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc1} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+  \cmidrule{2-11}          & mc2   & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            & seft  & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+      \midrule
+      \multirow{8}[6]{*}{P19} & \multirow{3}[2]{*}{sc} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc1} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+  \cmidrule{2-11}          & mc2   & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+            & seft  & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+      \bottomrule
+      \end{tabular}%
+    \label{tab:addlabel}%
+  \end{table*}%
 

### 2023-02-08 Manuscript_v1.tex
update

@@
 \label{sec:Related Works}
+
+Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where each sample is represented as an event sequence  $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is number of event, $t_j$ is event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label).
+The history of events is denoted as $\mathcal{H}_t=\{(t_j,e_j):t_j<t \} $
+
+In the multi-class setting where each mark can be exactly one of M marks, the point process is characterized by $M$ conidtional intensity functions (CIFs):
+
+\begin{equation}\label{eq:11}
+   \lambda_m^{*}(t)=\lim \frac{P(  \text{event of type m in } [t,t+\Delta t) |   \mathcal{H}_t  )}{\Delta t} 
+\end{equation}
+
+The default negative log-likelihood (NLL) of a point process for a single event sequnce $\mathcal{S}_{i}$ is defined as:
+
+\begin{multline}\label{eq:11}
+    -\log p(\mathcal{S}_{i})  =  - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log \lambda_m^{*}(t_j) \\   
+    + \sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
+\end{multline}
+
+If we assume that marks and events are independent,
+
+\begin{multline}\label{eq:11}
+    -\log p(\mathcal{S}_{i})  = \\
+      - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log p^{*}(e_j=m) \lambda^{*}(t_j) \\   
+    + \sum_{m = 1}^{M} p^{*}(e_j=m) \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
+\end{multline}
+
+[ntpp] has proposed to extend for mult
+
+marked cased
+
+relation to autoencoder
 

### 2023-02-09 Manuscript_v1.tex
update

@@
 
-\section{Related Works}
-\label{sec:Related Works}
-
+\subsubsection{Temporal point process}
 Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where each sample is represented as an event sequence  $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is number of event, $t_j$ is event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label).
@@
 
-[ntpp] has proposed to extend for mult
-
-marked cased
-
-relation to autoencoder
+[ntpp] has proposed the following CIF to extend for multi-label case:
+
+\begin{multline}\label{eq:11}
+    -\log p(\mathcal{S}_{i})  =  - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log \lambda_m^{*}(t_j) \\   
+    + \sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)\\
+    - \sum_{j = 1}^{L}\sum_{m = 1}^{M} (1-1(e_j=m)) \log \left(1-\lambda_m^{*}(t_j) \exp \left( - \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right) \right)
+\end{multline}
+
+It should be noted that the real advantage of point process is modeling non-event likelihoods in the form of integrals.
+If we neglect the integrals, we would achieve the cross-entropy and binary cross entropy loss in the multi-class and multi-label settings respectively for the prediction of next mark given history of events.
+
+\subsubsection{Neural temporal point process}
+
+Encoder-decoder architectures have proven to be effective in many applications.
+The main idea of a neural point process is to first encode the history of events until $t_j$ using different neural network architectures $h_j=Enc(\mathcal{H}_j)$.
+Then it tries to estimate $\lambda^*_m(t)$ or its integral for $ t_i < t \leq t_{i+1}$.
+
+\subsubsection{Deep learning for irregular sampling}
+An irregularly sampled data can be denoted as $\mathcal{D}=\{\mathcal{U}_i\}_{i=1}^{N}$ where $N$ is number of samples and each sample is represented as series of tuple $\mathcal{U}_i={(t_p,k_p,v_p)}$ where $t_p, k_p, v_p$ represents the time, modality and value of $p$-th datapoint respectively.
+
+\section{Related Works}
+\label{sec:Related Works}
 
@@
 \end{figure*}
+
+ The key advantage of our proposed model is to combine a  transformer-based event encoder with point process loss with a deep attention module that can handle an irregularly sampled time series for a any downstream prediciton task.
 

### 2023-02-09 Manuscript_v1.tex
update

@@
 
+\usepackage{biblatex} %Imports biblatex package
+% \addbibresource{sample.bib} %Import the bibliography file
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 information, including sponsor and financial support acknowledgment. For 
-example, ``This work was supported in part by the U.S. Department of 
-Commerce under Grant BS123456.'' }
-\thanks{The next few paragraphs should contain 
-the authors' current affiliations, including current address and e-mail. For 
-example, F. A. Author is with the National Institute of Standards and 
-Technology, Boulder, CO 80305 USA (e-mail: author@boulder.nist.gov). }
+example, ``This work was supported by DigiPredict Grant BS123456.'' }
+\thanks{H. Karami is with the EPFL, Lausanne, Switzerland (e-mail: hojjat.karami@epfl.ch). }
 \thanks{S. B. Author, Jr., was with Rice University, Houston, TX 77005 USA. He is 
@@
 \begin{abstract}
-    It is very well accepted that missingness in electronic health records (EHRs) are not at random which is regarded as informative missingness.
-The clinician’s decision on when to observe lab tests over time can be modeled using point processes.
-We propose a novel framework based on neural point process to analyze laboratory tests of ICU patients.
-This framework can take into account additional information for better characterization of conditional intensity function (CIF) as well as better accuracy in prediction of future timestamp and labels.
+
+    Irregular @hornSetFunctionsTime2020 sampling of time series in electronic health records (EHRs) is challenging for model development.
+In addition, the pattern of missingness for certain clinical variables are not at random as it is determined by clinicians decision and the state of patient.
+Point process is a mathematical framework for handling event sequence data which is also consistent with the irregular sampling.
+We propose \emph{TEEDAM}, a deep neural network that can learn patient representation from irregulary sample time series as well as informative missingness pattern of certain laboratory variables.
+We performed various experiments to show the effectiveness of event and state encoding for characterization of conditional intensity functions as well as downstream prediction task.
+Results show that in some cases learning from patterns may improve the performance of prediction task.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Enter keywords or phrases in alphabetical 
-order, separated by commas. For a list of suggested keywords, send a blank 
-e-mail to keywords@ieee.org or visit the site.
+    electronic health records (EHRs), Point Process, irregular sampling, informative missingness
 \end{IEEEkeywords}
@@
 
-Writing an effective abstract is an indispensable part of any form of research, since it that can motivate the audience to read follow the rest of the text.
-In this task, I have analyzed five abstracts from the field of artificial intelligence (AI).
-The table in section 2 shows a summary of the most important aspects of the analyzed abstracts.
+healthcare systems are one of main area for deploying machine learning models.
+Thanks to recent advancements in data collection technologies, policies and computing power data are recorded in a hosptial.
+This data consists of multiple modalities frome tabular and time series to clinical notes and images.
+We focus on tabular and time series.
+
+talk about the irregular sampling in EHRs and causes.///
+In and ideal setup, variables are measured on a regular basis, however, in a hospital due to medical costs, patient associated risks, sampling device.
+
+conseq of irr sampling///
+
+potential adv of irr sampl///
+
+imputation methods///
+
+inherently compatible methods///
+
+pp framework///
+
+Our aims??// In this work, we propose a DNN for joint modeling of NTPP and a DAM.
+Our first aim is to investigate the effectiveness of state encoding for CIF characterization.
+Second, we are interested in seeing the utility of event encoding a downstream task
+
+contribs?///
 
@@
 
- The key advantage of our proposed model is to combine a  transformer-based event encoder with point process loss with a deep attention module that can handle an irregularly sampled time series for a any downstream prediciton task.
-
-The entire data consists of $N$ samples  $\mathcal{D}=\{\mathcal{O}_{i}=(\mathcal{E}_{i},\mathcal{S}_{i}  )\}_{i=1}^N$ where $\mathcal{E}_{i}$ and $\mathcal{S}_{i}$ represents event and state sequences, respectively.
-
-Each event sequnce consist of $L$ tuples $\mathcal{E}_{i}=\{ (t_j, e_j)
-\}_{j=1}^{L_i}$ where $t_j\in \mathbb{R}  $ is the event timestamp and $ e_j $ is the event mark.
-
-Event sequence data consists of N sequences $\{\mathcal{S}_{i}\}_{i=1}^N$, where
- each sequence $\mathcal{S}_{i}$ is a series of $L_i$ events $\{ (t_j, e_j)
- \}_{j=1}^{L_i}$. Here, $e_j$ represents events that could be independent our mutually 
- exclusive occurring at $t_i$.
- \newline
- In addition to the event data, we might have additional information
-  $\{\mathcal{D}_{i}\}_{i=1}^N$. Suppose that each state is represented as $\{
-  (t_k, v_k, m_k) \}_{k=1}^{P_i}$, consisting of a time value $t_k\in
-  \mathbb{R}^{+} $, an observed value  $z_k\in \mathbb{R} $ and a modality
-  indicator  $m_k\in \{1, ..., M\} $.
-
-General schematic of TEDAM is depicted in FIG which consists of two separate module: Transformer Event Encoder (TEE) and Deep Attention Module (DAM).
-TEE encodes timestamp and mark of events using a transformer architecture, while DAM encodes all available information using one-dimensional attention mechanism.
-Finally, the learned representations are concatenated to be optimized for different loss functions.
+ The key advantage of our proposed model is to combine a  transformer-based event encoder (TEE) with point process loss with a deep attention module (DAM) that can handle an irregularly sampled time series for a any downstream prediciton task.
+In this situation the data is represented as $\mathcal{D}=\{\mathcal{S}_i,\mathcal{U}_i\}_{i=1}^{N}$.
+General schematic of TEDAM is depicted in FIG.
 
 \subsection{Event Encoder}
-We use a similar transformer architecture [thp] for encoding events with minor modifications. A transformer architecture is capable of .
+We use a similar transformer architecture [thp] for encoding events with minor modifications for event embedding.
 
-let . that can be embedded using a trainable embedding matrix.
+In the first step, we embed all event marks $E_{emb}=E \times W_emb$ where $E_i \in \mathbb{R}^{L \times M}$ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{M \times d_{emb} }  $ is the trainable embedding matrix.
 
-In the first step, we embed all event marks $E_{emb}=E \times U$ where $E_i \in [L,K] $ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{K \times d_{emb} }  $ is the trainable embedding matrix.
 In the second step, timestamps should be encoded and added to the event embedding, however, we propose to concatenate time encodings that can lead to better characterization of conditional intensity functions.
-Finally, the input of the transformer encoder will be $x_{emb}=[y(k_j), z(t_j)]$.
+Finally, the input of the transformer encoder will be $X_{emb}=[E_{emb}, T_{emb}] \in \mathbb{R}^{L \times (d_emb+t_emb)}$.
 
 Here, we use the standard transformer encoder similar to [vaswani] with masking matrix to prevent the model from looking into the future.
-we obtain the encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $ where $h_j$ containes the all available information before occurence of $j-th$ event.
+we obtain the encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $ where $h_j$ containes the all available information before occurence of $j$-th event.
 
-\subsection{bin of Event Encoder}
+\subsection{State Encoder}
 
-it is necessary to include temporal information. Similar to the original
-positional encoding [vaswani], we use a temporal encoding procedure:
+Similar to [setF], we use an attention-based aggregation approach for encoding all additional information.
+Each side information  $ (t_k, v_k, m_k) $ can be represented by $s_k=(z(t_k), v_k, m_k)$.
+we define attention $a(\mathcal{S}_k,s_k )$
+
+We define $\mathcal{S}_p$ to be the set of the first $p$ available information.
+The goal is to calculate $a(\mathcal{S}_p,s_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{S}_p$.
+This is achieved by computing an embedding of the set elements using a smaller set functions $f^{\prime}$, and projecting the concatenation of the set representation and the individual set element into d-dimensional space:
+
+The goal is to calculate $a(\mathcal{S}_p,s_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{S}_p$.
+
+This is achieved by embedding set elements $\mathcal{S}_p$ using a simpler set function $f^{\prime}$ such as a cumulative mean.
+
+\begin{equation}
+    f^{\prime}(\mathcal{U}_p) = g^{\prime} \left(  \frac{1}{|p|} \sum_{u_k \in \mathcal{U}_p}  h^{\prime}(u_k;\theta^{\prime}) ;\rho^{\prime}\right) 
+\end{equation}
+
+Then we can compute key values using key matrix $W^k \in \mathbb{R}^{(d_g+d_s) \times d_{prod}}$
+
+\begin{equation}
+    K_p=[f^{\prime}(\mathcal{U}_p), u_p]^T W^K   
+\end{equation}
+
+using a query vector $w^q \in \mathbb{R}^{d_{prod}}$
+
+\begin{equation*}
+    a(\mathcal{U}_p,u_p)=  softmax(\frac{K_p.w^q}{\sqrt{d}  })
+\end{equation*}
+
+Finally, we compute a weighted aggregation of set elements:
+
+\begin{equation*}
+    z_p = f(\mathcal{U}_p) =     
+    g\left(
+    \sum_{s_k \in \mathcal{U}_p}  a(\mathcal{U}_p,s_k)h(s_k;\theta);\rho 
+    \right) 
+\end{equation*}
+
+we regard $z_p \in \mathbb{R}^{d_{\rho}}$ as the representation of state data until arrival of $p$-th datapoint.
+
+Finally, we need to combine event embeddings $ H_{L \times d_e}$ and state embeddings $Z_{P \times d_{\rho}}$, however, the length of each does not match.
+As a result, we consider the reduced version of state matrix as below:
+
+\begin{equation*}
+    Z^{red}_i=Z_p where p=argmax(z \leq i)
+\end{equation*}
+
+Without loss of generality, we can consider multiple heads by adding an additional dimension to keys and queries.\\
+
+All formulas are:
+
+\begin{equation*}
+    \begin{cases}
+
+f^{\prime}(\mathcal{S}_p) = g^{\prime} \left(   
+ \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) 
+\right) \\
+
+K_p=[f^{\prime}(\mathcal{S}_p), s_p]^T W^K \\
+e_p=\frac{K_p.w^q}{\sqrt{d}  }\\
+
+a(\mathcal{S}_p,s_k)=\frac{\exp (e_p)}{\sum_{k\leq p} \exp (e_k)  } \\
+
+f(\mathcal{S}_p) = \sum_{s_k \in \mathcal{S}_p}  a(\mathcal{S}_p,s_k)h_{\theta}(s_k) \\
+
+z_p=g_{\psi}\left(         f(\mathcal{S}_p)         \right) \\
+
+\end{cases} 
+\end{equation*}
 
@@
 \end{equation} 
-
-Here, $d_t \in \mathbb{N} $ is the dimensionality of encoded timestamp and $z
-\in \mathbb{R}^{d_t} $ is the embeding vector of timestamp.
-
-Each event mark $e_j$ is projected to a sparse binary vector representation. We
-add an embedding layer to achieve a more compact and efficient representation
-$emb$. Here, $w$ and $b$ are weights and biases of the embedding layer which can be
-learned during network training.
-
-While previous works suggested adding temporal encoding to the embedded events,
-we propose to concatenate these two vectors. The effectiveness of this
-concatenation is further investigated in the results.
-
-\begin{equation} x_{emb}=[y(k_j), z(t_j)] \end{equation}
-
-Now that the $x_{emb}$ is ready for encoding, it is encoded through a standard
-transformer encoder with multiple layers and attention heads.
-
-\begin{equation}
-    x_{enc}=TE(x_{emb})    
-\end{equation}
-
-\subsection{State Encoder}
-Here, we propose a method to incorporate additional information for a
-better representation. In healthcare, much data is available from different
-modalities such as vital signs and laboratory values.
-
-Similar to [setF], we use an attention-based aggregation approach for encoding all additional information.
-Each side information  $ (t_k, v_k, m_k) $ can be represented by $s_k=(z(t_k), v_k, m_k)$.
-we define attention $a(\mathcal{S}_k,s_k )$
-
-We define $\mathcal{S}_p$ to be the set of the first $p$ available information.
-The goal is to calculate $a(\mathcal{S}_p,s_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{S}_p$.
-This is achieved by computing an embedding of the set elements using a smaller set functions $f^{\prime}$, and projecting the concatenation of the set representation and the individual set element into d-dimensional space:
-
-\begin{align*}
-    f^{\prime}(\mathcal{S}_p) = g^{\prime} \left(  \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) \right)  \\
-    K_p=[f^{\prime}(\mathcal{S}_p), s_p]^T W^K   
-\end{align*}
-
-Furthermore, we define a query vector $w^q \in \mathbb{R}^d$, which allows the model to summarize different aspects of the dataset via
-
-\begin{equation*}
-    e_p=\frac{K_p.w^q}{\sqrt{d}  }
-\end{equation*}
-
-Now, the desired attention can be computed as follows:
-
-\begin{equation*}
-    a(\mathcal{S}_p,s_k)=\frac{\exp (e_p)}{\sum_{k\leq p} \exp (e_k)  } 
-\end{equation*}
-
-Finally, we compute a weighted aggregation of set elements:
-
-\begin{equation*}
-    f(\mathcal{S}_p) =     
-    g_{\psi}\left(
-    \sum_{s_k \in \mathcal{S}_p}  a(\mathcal{S}_p,s_k)h_{\theta}(s_k) 
-    \right) 
-\end{equation*}
-
-Without loss of generality, we can consider multiple heads by adding an additional dimension to keys and queries.\\
-
-All formulas are:
-
-\begin{equation*}
-    \begin{cases} 
-f^{\prime}(\mathcal{S}_p) = g^{\prime} \left(   
- \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) 
-\right) \\
-K_p=[f^{\prime}(\mathcal{S}_p), s_p]^T W^K \\
-e_p=\frac{K_p.w^q}{\sqrt{d}  }\\
-a(\mathcal{S}_p,s_k)=\frac{\exp (e_p)}{\sum_{k\leq p} \exp (e_k)  } \\
-f(\mathcal{S}_p) = \sum_{s_k \in \mathcal{S}_p}  a(\mathcal{S}_p,s_k)h_{\theta}(s_k) \\
-    z_p=g_{\psi}\left(         f(\mathcal{S}_p)         \right) \\
-    \end{cases} 
-\end{equation*}
-
-\subsection{Mark and Time Decoder}
-
-by concatenating encoded events $x_{enc}$ and additional information $f^{\prime}(\mathcal{S}_p)$, we can predict next marks and times as follows:
-
-\begin{equation}
-    \hat{e}_{j+1}=MLP([x_j, f^{\prime}(\mathcal{S}_j)])
-\end{equation}
-
-\begin{equation}
-    \hat{t}_{j+1}=MLP([x_j, f^{\prime}(\mathcal{S}_j)])
-\end{equation}
-
-\begin{equation}
-    \mathcal{L}_{mark} =
-\end{equation}
-\begin{equation}
-    \mathcal{L}_{time} = \sum_{j=1}^{L-1} \left(  (\hat{t}_{j+1}-t_j)-(t_{j+1}-t_j)  \right)^2
-\end{equation}
-
-\subsection{Event Decoder}
-Once we obtain a representation of a patient using embedded events and states, we can try to parameterize conditional intensity functions (CIFs) of the events.
-
-In neural point process literature, many approaches have been propose to decode either conditional or cumulative intensity function.
-
 \begin{equation}
@@
 
+\subsection{Event Decoder}
+Once we obtain a representation of a patient using embedded events and states, we can try to parameterize conditional intensity functions (CIFs) of the events.
+
+In neural point process literature, many approaches have been propose to decode either conditional or cumulative intensity function.
+We will use a decoder similar to [sahp] as it can model both exciting and inhibiting effects for modeling CIFs.
+
+% \begin{gather*} 
+%    \mu_{m,i+1}=gelu(h_{i+1}W_{m,\mu}), 
+% \end{gather*}
+
+\begin{gather*} 
+    \mu_{m,i+1}=gelu(h_{i+1}W_{m,\mu}  +  z_{i+1}W_{m,\mu}), \\ 
+    \eta_{m,i+1}=gelu(h_{i+1}W_{m,\eta}  +  z_{i+1}W_{m,\eta}), \\
+    \gamma_{m,i+1}=gelu(h_{i+1}W_{m,\gamma}  +  z_{i+1}W_{m,\gamma}),
+    \end{gather*}
+
+Finally, we can express the intensity function as follows:
+
+\begin{multline*}
+    \lambda_m(t)=softplus(\mu_{m,i+1}+\\
+    (\eta_{m,i+1}- \mu_{m,i+1}) \exp(-\gamma_{m,i+1}(t-t_i))    ),
+\end{multline*}
+
+for $t \in (t_i, t_{i+1}]$,where the $softplus$ is used to constrain thei intensity function to be positive.
+
+\subsection{Loss Function}
+
+We define a multi-objective loss function 
+$ \mathcal{L}   =   \mathcal{L}_{CIF}   +   \mathcal{L}_{mark}  +   \mathcal{L}_{state}$.
+
 \section{Experiments}
@@
 
+We perform various experiments to show the effectiveness of each component in out model.
 \subsection*{Datasets}
+
+To show the utility of time concatenation and marked loss, we consider three datasets:
+
+\textbf{Synthea(\emph{Syn})}.
+We used the Synthea simulator (Walonoski et al., 2018) which generates patient-level EHRs using human expert curated Markov processes.
+Here, we reused the already processed version of this data by [ntpp].
+
+\textbf{ReTweets (\emph{RT})}.
+The Retweets dataset contains sequences of tweets, where each sequence contains an origin tweet (i.e., some user initiates a tweet), and some follow-up tweets.
+We record the time and the user tag of each tweet.
+Further, users are grouped into three categories based on the number of their followers: “small”, “medium”, and “large”
+
+\textbf{Stackoverflow (\emph{SO})}.
+is a question-answering website.
+The website rewards users with badges to promote engagement in the community, and the same badge can be rewarded multiple times to the same user.
+We collect data in a two-year period, and we treat each user’s reward history as a sequence.
+Each event in the sequence signifies receipt of a particular medal.
+
+Furthermore, we consider two EHRs provided by physionet challenge to investigate the advantage of irregular sample and point process modeling in the same time.
+
+\textbf{Physionet 2012 Mortality Prediction Challenge (\emph{P19})}.
+The 2012 Physionet challenge dataset (Goldberger et al., 2000),contains 12, 000 ICU stays each of which lasts at least 48 h.
+For each stay, a set of general descriptors (such as gender or age) are collected at admission time.
+Depending on the course of the stay and patient status, up to 37 time series variables were measured (e.g.
+blood pressure, lactate, and respiration rate).
+While some modalities might be measured in regular time intervals (e.g.
+hourly or daily), some are only collected when required; moreover, not all variables are available for each stay.
 
@@
 
-% To show the effectiveness of the proposed method, we further consider the following scenarios:
-% \begin{itemize}
-%     \item Enc(Event)+Dec(NextEvent): Here
-%     \item Enc(Event)+Dec(NextEvent+CIF): Here
-%     \item Enc(Event+state)+Dec(NextEvent): Here
-%     \item Enc(Event+state)+Dec(NextEvent+CIF): Here
-% \end{itemize}
+\subsection*{Scenarios}
 
-\textbf{MIMIC-IV (\emph{M4})}.
-We selected Medical Information Mart for Intensive Care (MIMIC) IV [1], which is a real-world clinical database comprising health data relating to over 40,000 patients admitted to ICU at the Beth Israel Deaconess Medical Center.
+To show the utility of time concatenation and marked-shp we report the metrics for SO, RT and SYN and compare it with 3 baselines: SAHP, THP, and GRU-CP.
 
-\textbf{Synthea(\emph{Syn})}.
-We used the Synthea simulator (Walonoski et al., 2018) which generates patient-level EHRs using human expert curated Markov processes.
-Here, we reused the already processed version of this data by Edgauard.
+To show the utility of additional information for CIF modeling, we report NLL/events for P12 and P19 in the following conditions: TE, TE+DAM, TE+noise
 
-\textbf{Stackoverflow (\emph{SO})}.
-StackOverflow is a question-answering website.
-The website rewards users with badges to promote engagement in the community, and the same badge can be rewarded multiple times to the same user.
-We collect data in a two-year period, and we treat each user’s reward history as a sequence.
-Each event in the sequence signifies receipt of a particular medal.
-
-\subsection*{Scenarios}
-To investigate the effectiveness of the proposed method, we consider three input scenarios (TE, DA, TE+DA) as well as three loss functions (next event, CIF, next event+CIF) which would result in nine scenarios.
-
-The first series of experiments are conducted to investigate the advantage of encoding additional information for paramterization of intensity functions. We consider seven scenarios:
-Here, the baseline models
-
-To show the effectiveness of time concatenation we report.
+Finally, we investigate whether point process characterization can be useful in a downstream task or not.
+As each dataset consists of different hospitals and the pattern of irregular sampling might differ from hosptial to hospital, we assume different settings: single-center, multi-center and external evaluation.
 
 \subsection*{Baselines}
-we use NEURALTPP that is already developed pipeline by as they already considered a lot of comibinations.
+
+we compare our model against existing models: THP, SAHP, GRU-CP.
 
@@
 
+\subsection*{Training Details}
+To be completed.
 \section{Results and Discussion}
@@
 % Table generated by Excel2LaTeX from sheet 'time concat'
-\begin{table}[h!]
+\begin{table*}[htbp]
     \centering
     \caption{Add caption}
-      \begin{tabular}{ccccc}
+      \begin{tabular}{cccccc}
       \toprule
-            &       & \multicolumn{2}{c}{shp+mark} &  \\
-  \cmidrule{3-4}    Dataset & Metric & concat & sum   & baseline \\
+            &       & \multicolumn{2}{c}{TEEDAM} &       &  \\
+  \cmidrule{3-4}    Dataset & Metric & concat & sum   & SAHP  & GRU-CP \\
       \midrule
-      SO    & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \multirow{2}[2]{*}{SO} & LL/\#events &       &       &       &  \\
+            & AUROC &       &       &       &  \\
       \midrule
-      Synthea & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \multirow{2}[2]{*}{Synthea} & LL/\#events &       &       &       &  \\
+            & AUROC &       &       &       & 0.85(.014) \\
       \midrule
-      ReTweet & NLL   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & AUROC & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & time  & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \multirow{2}[2]{*}{ReTweet} & LL/\#events & 1.83 (0.09) & 1.5 (0.18) &       &  \\
+            & AUROC & 0.68 (0.03) & 0.67 (0.05) &       & 0.611(0.001) \\
       \bottomrule
@@
     \label{tab:addlabel}%
-  \end{table}%
+  \end{table*}%
 
@@
 \label{sec:Conclusion}
-
-\section{Introduction}
-\label{sec:introduction}
-\IEEEPARstart{T}{his} document is a template for \LaTeX. If you are 
-reading a paper or PDF version of this document, please download the 
-electronic file, trans\_jour.tex, from the IEEE Web site at \underline
-{http://www.ieee.org/authortools/trans\_jour.tex} so you can use it to prepare your manuscript. If 
-you would prefer to use LaTeX, download IEEE's LaTeX style and sample files 
-from the same Web page. You can also explore using the Overleaf editor at 
-\underline
-{https://www.overleaf.com/blog/278-how-to-use-overleaf-with-}\discretionary{}{}{}\underline
-{ieee-collabratec-your-quick-guide-to-getting-started\#.}\discretionary{}{}{}\underline{xsVp6tpPkrKM9}
-
-If your paper is intended for a conference, please contact your conference 
-editor concerning acceptable word processor formats for your particular 
-conference.
-
-IEEE will do the final formatting of your paper. If your paper is intended 
-for a conference, please observe the conference page limits.
-
-\subsection{Abbreviations and Acronyms}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have already been defined in the abstract. Abbreviations 
-such as IEEE, SI, ac, and dc do not have to be defined. Abbreviations that 
-incorporate periods should not have spaces: write ``C.N.R.S.,'' not ``C. N. 
-R. S.'' Do not use abbreviations in the title unless they are unavoidable 
-(for example, ``IEEE'' in the title of this article).
-
-\subsection{Other Recommendations}
-Use one space after periods and colons. Hyphenate complex modifiers: 
-``zero-field-cooled magnetization.'' Avoid dangling participles, such as, 
-``Using \eqref{eq}, the potential was calculated.'' [It is not clear who or what 
-used \eqref{eq}.] Write instead, ``The potential was calculated by using \eqref{eq},'' or 
-``Using \eqref{eq}, we calculated the potential.''
-
-Use a zero before decimal points: ``0.25,'' not ``.25.'' Use 
-``cm$^{3}$,'' not ``cc.'' Indicate sample dimensions as ``0.1 cm 
-$\times $ 0.2 cm,'' not ``0.1 $\times $ 0.2 cm$^{2}$.'' The 
-abbreviation for ``seconds'' is ``s,'' not ``sec.'' Use 
-``Wb/m$^{2}$'' or ``webers per square meter,'' not 
-``webers/m$^{2}$.'' When expressing a range of values, write ``7 to 
-9'' or ``7--9,'' not ``7$\sim $9.''
-
-A parenthetical statement at the end of a sentence is punctuated outside of 
-the closing parenthesis (like this). (A parenthetical sentence is punctuated 
-within the parentheses.) In American English, periods and commas are within 
-quotation marks, like ``this period.'' Other punctuation is ``outside''! 
-Avoid contractions; for example, write ``do not'' instead of ``don't.'' The 
-serial comma is preferred: ``A, B, and C'' instead of ``A, B and C.''
-
-If you wish, you may write in the first person singular or plural and use 
-the active voice (``I observed that $\ldots$'' or ``We observed that $\ldots$'' 
-instead of ``It was observed that $\ldots$''). Remember to check spelling. If 
-your native language is not English, please get a native English-speaking 
-colleague to carefully proofread your paper.
-
-Try not to use too many typefaces in the same article. You're writing
-scholarly papers, not ransom notes. Also please remember that MathJax
-can't handle really weird typefaces.
-
-\subsection{Equations}
-Number equations consecutively with equation numbers in parentheses flush 
-with the right margin, as in \eqref{eq}. To make your equations more 
-compact, you may use the solidus (~/~), the exp function, or appropriate 
-exponents. Use parentheses to avoid ambiguities in denominators. Punctuate 
-equations when they are part of a sentence, as in
-\begin{equation}E=mc^2.\label{eq}\end{equation}
-
-Be sure that the symbols in your equation have been defined before the 
-equation appears or immediately following. Italicize symbols ($T$ might refer 
-to temperature, but T is the unit tesla). Refer to ``\eqref{eq},'' not ``Eq. \eqref{eq}'' 
-or ``equation \eqref{eq},'' except at the beginning of a sentence: ``Equation \eqref{eq} 
-is $\ldots$ .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-If you are submitting your paper to a colorized journal, you can use
-the following two lines at the start of the article to ensure its
-appearance resembles the final copy:
-
-\smallskip\noindent
-\begin{small}
-\begin{tabular}{l}
-\verb+\+\texttt{documentclass[journal,twoside,web]\{ieeecolor\}}\\
-\verb+\+\texttt{usepackage\{\textit{Journal\_Name}\}}
-\end{tabular}
-\end{small}
-
-\section{Units}
-Use either SI (MKS) or CGS as primary units. (SI units are strongly 
-encouraged.) English units may be used as secondary units (in parentheses). 
-This applies to papers in data storage. For example, write ``15 
-Gb/cm$^{2}$ (100 Gb/in$^{2})$.'' An exception is when 
-English units are used as identifiers in trade, such as ``3\textonehalf-in 
-disk drive.'' Avoid combining SI and CGS units, such as current in amperes 
-and magnetic field in oersteds. This often leads to confusion because 
-equations do not balance dimensionally. If you must use mixed units, clearly 
-state the units for each quantity in an equation.
-
-The SI unit for magnetic field strength $H$ is A/m. However, if you wish to use 
-units of T, either refer to magnetic flux density $B$ or magnetic field 
-strength symbolized as $\mu _{0}H$. Use the center dot to separate 
-compound units, e.g., ``A$\cdot $m$^{2}$.''
-
-\section{Some Common Mistakes}
-The word ``data'' is plural, not singular. The subscript for the 
-permeability of vacuum $\mu _{0}$ is zero, not a lowercase letter 
-``o.'' The term for residual magnetization is ``remanence''; the adjective 
-is ``remanent''; do not write ``remnance'' or ``remnant.'' Use the word 
-``micrometer'' instead of ``micron.'' A graph within a graph is an 
-``inset,'' not an ``insert.'' The word ``alternatively'' is preferred to the 
-word ``alternately'' (unless you really mean something that alternates). Use 
-the word ``whereas'' instead of ``while'' (unless you are referring to 
-simultaneous events). Do not use the word ``essentially'' to mean 
-``approximately'' or ``effectively.'' Do not use the word ``issue'' as a 
-euphemism for ``problem.'' When compositions are not specified, separate 
-chemical symbols by en-dashes; for example, ``NiMn'' indicates the 
-intermetallic compound Ni$_{0.5}$Mn$_{0.5}$ whereas 
-``Ni--Mn'' indicates an alloy of some composition 
-Ni$_{x}$Mn$_{1-x}$.
-
-\begin{figure}[!t]
-\centerline{\includegraphics[width=\columnwidth]{fig1.png}}
-\caption{Magnetization as a function of applied field.
-It is good practice to explain the significance of the figure in the caption.}
-\label{fig1}
-\end{figure}
-
-Be aware of the different meanings of the homophones ``affect'' (usually a 
-verb) and ``effect'' (usually a noun), ``complement'' and ``compliment,'' 
-``discreet'' and ``discrete,'' ``principal'' (e.g., ``principal 
-investigator'') and ``principle'' (e.g., ``principle of measurement''). Do 
-not confuse ``imply'' and ``infer.''
-
-Prefixes such as ``non,'' ``sub,'' ``micro,'' ``multi,'' and ``ultra'' are 
-not independent words; they should be joined to the words they modify, 
-usually without a hyphen. There is no period after the ``et'' in the Latin 
-abbreviation ``\emph{et al.}'' (it is also italicized). The abbreviation ``i.e.,'' means 
-``that is,'' and the abbreviation ``e.g.,'' means ``for example'' (these 
-abbreviations are not italicized).
-
-A general IEEE styleguide is available at \underline{http://www.ieee.org/authortools}.
-
-\section{Guidelines for Graphics Preparation and Submission}
-\label{sec:guidelines}
-
-\subsection{Types of Graphics}
-The following list outlines the different types of graphics published in 
-IEEE journals. They are categorized based on their construction, and use of 
-color/shades of gray:
-
-\subsubsection{Color/Grayscale figures}
-{Figures that are meant to appear in color, or shades of black/gray. Such 
-figures may include photographs, illustrations, multicolor graphs, and 
-flowcharts.}
-
-\subsubsection{Line Art figures}
-{Figures that are composed of only black lines and shapes. These figures 
-should have no shades or half-tones of gray, only black and white.}
-
-\subsubsection{Author photos}
-{Head and shoulders shots of authors that appear at the end of our papers. }
-
-\subsubsection{Tables}
-{Data charts which are typically black and white, but sometimes include 
-color.}
-
-\begin{table}
-\caption{Units for Magnetic Properties}
-\label{table}
-\setlength{\tabcolsep}{3pt}
-\begin{tabular}{|p{25pt}|p{75pt}|p{115pt}|}
-\hline
-Symbol& 
-Quantity& 
-Conversion from Gaussian and \par CGS EMU to SI $^{\mathrm{a}}$ \\
-\hline
-$\Phi $& 
-magnetic flux& 
-1 Mx $\to  10^{-8}$ Wb $= 10^{-8}$ V$\cdot $s \\
-$B$& 
-magnetic flux density, \par magnetic induction& 
-1 G $\to  10^{-4}$ T $= 10^{-4}$ Wb/m$^{2}$ \\
-$H$& 
-magnetic field strength& 
-1 Oe $\to  10^{3}/(4\pi )$ A/m \\
-$m$& 
-magnetic moment& 
-1 erg/G $=$ 1 emu \par $\to 10^{-3}$ A$\cdot $m$^{2} = 10^{-3}$ J/T \\
-$M$& 
-magnetization& 
-1 erg/(G$\cdot $cm$^{3}) =$ 1 emu/cm$^{3}$ \par $\to 10^{3}$ A/m \\
-4$\pi M$& 
-magnetization& 
-1 G $\to  10^{3}/(4\pi )$ A/m \\
-$\sigma $& 
-specific magnetization& 
-1 erg/(G$\cdot $g) $=$ 1 emu/g $\to $ 1 A$\cdot $m$^{2}$/kg \\
-$j$& 
-magnetic dipole \par moment& 
-1 erg/G $=$ 1 emu \par $\to 4\pi \times  10^{-10}$ Wb$\cdot $m \\
-$J$& 
-magnetic polarization& 
-1 erg/(G$\cdot $cm$^{3}) =$ 1 emu/cm$^{3}$ \par $\to 4\pi \times  10^{-4}$ T \\
-$\chi , \kappa $& 
-susceptibility& 
-1 $\to  4\pi $ \\
-$\chi_{\rho }$& 
-mass susceptibility& 
-1 cm$^{3}$/g $\to  4\pi \times  10^{-3}$ m$^{3}$/kg \\
-$\mu $& 
-permeability& 
-1 $\to  4\pi \times  10^{-7}$ H/m \par $= 4\pi \times  10^{-7}$ Wb/(A$\cdot $m) \\
-$\mu_{r}$& 
-relative permeability& 
-$\mu \to \mu_{r}$ \\
-$w, W$& 
-energy density& 
-1 erg/cm$^{3} \to  10^{-1}$ J/m$^{3}$ \\
-$N, D$& 
-demagnetizing factor& 
-1 $\to  1/(4\pi )$ \\
-\hline
-\multicolumn{3}{p{251pt}}{Vertical lines are optional in tables. Statements that serve as captions for 
-the entire table do not need footnote letters. }\\
-\multicolumn{3}{p{251pt}}{$^{\mathrm{a}}$Gaussian units are the same as cg emu for magnetostatics; Mx 
-$=$ maxwell, G $=$ gauss, Oe $=$ oersted; Wb $=$ weber, V $=$ volt, s $=$ 
-second, T $=$ tesla, m $=$ meter, A $=$ ampere, J $=$ joule, kg $=$ 
-kilogram, H $=$ henry.}
-\end{tabular}
-\label{tab1}
-\end{table}
-
-\subsection{Multipart figures}
-Figures compiled of more than one sub-figure presented side-by-side, or 
-stacked. If a multipart figure is made up of multiple figure
-types (one part is lineart, and another is grayscale or color) the figure 
-should meet the stricter guidelines.
-
-\subsection{File Formats For Graphics}\label{formats}
-Format and save your graphics using a suitable graphics processing program 
-that will allow you to create the images as PostScript (PS), Encapsulated 
-PostScript (.EPS), Tagged Image File Format (.TIFF), Portable Document 
-Format (.PDF), Portable Network Graphics (.PNG), or Metapost (.MPS), sizes them, and adjusts 
-the resolution settings. When 
-submitting your final paper, your graphics should all be submitted 
-individually in one of these formats along with the manuscript.
-
-\subsection{Sizing of Graphics}
-Most charts, graphs, and tables are one column wide (3.5 inches/88 
-millimeters/21 picas) or page wide (7.16 inches/181 millimeters/43 
-picas). The maximum depth a graphic can be is 8.5 inches (216 millimeters/54
-picas). When choosing the depth of a graphic, please allow space for a 
-caption. Figures can be sized between column and page widths if the author 
-chooses, however it is recommended that figures are not sized less than 
-column width unless when necessary.
-
-There is currently one publication with column measurements that do not 
-coincide with those listed above. Proceedings of the IEEE has a column 
-measurement of 3.25 inches (82.5 millimeters/19.5 picas).
-
-The final printed size of author photographs is exactly
-1 inch wide by 1.25 inches tall (25.4 millimeters$\,\times\,$31.75 millimeters/6 
-picas$\,\times\,$7.5 picas). Author photos printed in editorials measure 1.59 inches 
-wide by 2 inches tall (40 millimeters$\,\times\,$50 millimeters/9.5 picas$\,\times\,$12 
-picas).
-
-\subsection{Resolution }
-The proper resolution of your figures will depend on the type of figure it 
-is as defined in the ``Types of Figures'' section. Author photographs, 
-color, and grayscale figures should be at least 300dpi. Line art, including 
-tables should be a minimum of 600dpi.
-
-\subsection{Vector Art}
-In order to preserve the figures' integrity across multiple computer 
-platforms, we accept files in the following formats: .EPS/.PDF/.PS. All 
-fonts must be embedded or text converted to outlines in order to achieve the 
-best-quality results.
-
-\subsection{Color Space}
-The term color space refers to the entire sum of colors that can be 
-represented within the said medium. For our purposes, the three main color 
-spaces are Grayscale, RGB (red/green/blue) and CMYK 
-(cyan/magenta/yellow/black). RGB is generally used with on-screen graphics, 
-whereas CMYK is used for printing purposes.
-
-All color figures should be generated in RGB or CMYK color space. Grayscale 
-images should be submitted in Grayscale color space. Line art may be 
-provided in grayscale OR bitmap colorspace. Note that ``bitmap colorspace'' 
-and ``bitmap file format'' are not the same thing. When bitmap color space 
-is selected, .TIF/.TIFF/.PNG are the recommended file formats.
-
-\subsection{Accepted Fonts Within Figures}
-When preparing your graphics IEEE suggests that you use of one of the 
-following Open Type fonts: Times New Roman, Helvetica, Arial, Cambria, and 
-Symbol. If you are supplying EPS, PS, or PDF files all fonts must be 
-embedded. Some fonts may only be native to your operating system; without 
-the fonts embedded, parts of the graphic may be distorted or missing.
-
-A safe option when finalizing your figures is to strip out the fonts before 
-you save the files, creating ``outline'' type. This converts fonts to 
-artwork what will appear uniformly on any screen.
-
-\subsection{Using Labels Within Figures}
-
-\subsubsection{Figure Axis labels }
-Figure axis labels are often a source of confusion. Use words rather than 
-symbols. As an example, write the quantity ``Magnetization,'' or 
-``Magnetization M,'' not just ``M.'' Put units in parentheses. Do not label 
-axes only with units. As in Fig. 1, for example, write ``Magnetization 
-(A/m)'' or ``Magnetization (A$\cdot$m$^{-1}$),'' not just ``A/m.'' Do not label axes with a ratio of quantities and 
-units. For example, write ``Temperature (K),'' not ``Temperature/K.''
-
-Multipliers can be especially confusing. Write ``Magnetization (kA/m)'' or 
-``Magnetization (10$^{3}$ A/m).'' Do not write ``Magnetization 
-(A/m)$\,\times\,$1000'' because the reader would not know whether the top 
-axis label in Fig. 1 meant 16000 A/m or 0.016 A/m. Figure labels should be 
-legible, approximately 8 to 10 point type.
-
-\subsubsection{Subfigure Labels in Multipart Figures and Tables}
-Multipart figures should be combined and labeled before final submission. 
-Labels should appear centered below each subfigure in 8 point Times New 
-Roman font in the format of (a) (b) (c).
-
-\subsection{File Naming}
-Figures (line artwork or photographs) should be named starting with the 
-first 5 letters of the author's last name. The next characters in the 
-filename should be the number that represents the sequential 
-location of this image in your article. For example, in author 
-``Anderson's'' paper, the first three figures would be named ander1.tif, 
-ander2.tif, and ander3.ps.
-
-Tables should contain only the body of the table (not the caption) and 
-should be named similarly to figures, except that `.t' is inserted 
-in-between the author's name and the table number. For example, author 
-Anderson's first three tables would be named ander.t1.tif, ander.t2.ps, 
-ander.t3.eps.
-
-Author photographs should be named using the first five characters of the 
-pictured author's last name. For example, four author photographs for a 
-paper may be named: oppen.ps, moshc.tif, chen.eps, and duran.pdf.
-
-If two authors or more have the same last name, their first initial(s) can 
-be substituted for the fifth, fourth, third$\ldots$ letters of their surname 
-until the degree where there is differentiation. For example, two authors 
-Michael and Monica Oppenheimer's photos would be named oppmi.tif, and 
-oppmo.eps.
-
-\subsection{Referencing a Figure or Table Within Your Paper}
-When referencing your figures and tables within your paper, use the 
-abbreviation ``Fig.'' even at the beginning of a sentence. Do not abbreviate 
-``Table.'' Tables should be numbered with Roman Numerals.
-
-\subsection{Checking Your Figures: The IEEE Graphics Analyzer}
-The IEEE Graphics Analyzer enables authors to pre-screen their graphics for 
-compliance with IEEE Transactions and Journals standards before submission. 
-The online tool, located at
-\underline{http://graphicsqc.ieee.org/}, allows authors to 
-upload their graphics in order to check that each file is the correct file 
-format, resolution, size and colorspace; that no fonts are missing or 
-corrupt; that figures are not compiled in layers or have transparency, and 
-that they are named according to the IEEE Transactions and Journals naming 
-convention. At the end of this automated process, authors are provided with 
-a detailed report on each graphic within the web applet, as well as by 
-email.
-
-For more information on using the Graphics Analyzer or any other graphics 
-related topic, contact the IEEE Graphics Help Desk by e-mail at 
-graphics@ieee.org.
-
-\subsection{Submitting Your Graphics}
-Because IEEE will do the final formatting of your paper,
-you do not need to position figures and tables at the top and bottom of each 
-column. In fact, all figures, figure captions, and tables can be placed at 
-the end of your paper. In addition to, or even in lieu of submitting figures 
-within your final manuscript, figures should be submitted individually, 
-separate from the manuscript in one of the file formats listed above in 
-Section \ref{formats}. Place figure captions below the figures; place table titles 
-above the tables. Please do not include captions as part of the figures, or 
-put them in ``text boxes'' linked to the figures. Also, do not place borders 
-around the outside of your figures.
-
-\subsection{Color Processing/Printing in IEEE Journals}
-All IEEE Transactions, Journals, and Letters allow an author to publish 
-color figures on IEEE Xplore\textregistered\ at no charge, and automatically 
-convert them to grayscale for print versions. In most journals, figures and 
-tables may alternatively be printed in color if an author chooses to do so. 
-Please note that this service comes at an extra expense to the author. If 
-you intend to have print color graphics, include a note with your final 
-paper indicating which figures or tables you would like to be handled that 
-way, and stating that you are willing to pay the additional fee.
-
-\section{Conclusion}
-A conclusion section is not required. Although a conclusion may review the 
-main points of the paper, do not replicate the abstract as the conclusion. A 
-conclusion might elaborate on the importance of the work or suggest 
-applications and extensions.
-
-\appendices
-
-Appendixes, if needed, appear before the acknowledgment.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in American English is 
-without an ``e'' after the ``g.'' Use the singular heading even if you have 
-many acknowledgments. Avoid expressions such as ``One of us (S.B.A.) would 
-like to thank $\ldots$ .'' Instead, write ``F. A. Author thanks $\ldots$ .'' In most 
-cases, sponsor and financial support acknowledgments are placed in the 
-unnumbered footnote on the first page, not here.
-
-\section*{References and Footnotes}
-
-\subsection{References}
-References need not be cited in text. When they are, they appear on the 
-line, in square brackets, inside the punctuation. Multiple references are 
-each numbered with separate brackets. When citing a section in a book, 
-please give the relevant page numbers. In text, refer simply to the 
-reference number. Do not use ``Ref.'' or ``reference'' except at the 
-beginning of a sentence: ``Reference \cite{b3} shows $\ldots$ .'' Please do not use 
-automatic endnotes in \emph{Word}, rather, type the reference list at the end of the 
-paper using the ``References'' style.
-
-Reference numbers are set flush left and form a column of their own, hanging 
-out beyond the body of the reference. The reference numbers are on the line, 
-enclosed in square brackets. In all references, the given name of the author 
-or editor is abbreviated to the initial only and precedes the last name. Use 
-them all; use \emph{et al.} only if names are not given. Use commas around Jr., 
-Sr., and III in names. Abbreviate conference titles. When citing IEEE 
-transactions, provide the issue number, page range, volume number, year, 
-and/or month if available. When referencing a patent, provide the day and 
-the month of issue, or application. References may not include all 
-information; please obtain and include relevant information. Do not combine 
-references. There must be only one reference with each number. If there is a 
-URL included with the print reference, it can be included at the end of the 
-reference.
-
-Other than books, capitalize only the first word in a paper title, except 
-for proper nouns and element symbols. For papers published in translation 
-journals, please give the English citation first, followed by the original 
-foreign-language citation See the end of this document for formats and 
-examples of common references. For a complete discussion of references and 
-their formats, see the IEEE style manual at
-\underline{http://www.ieee.org/authortools}.
-
-\subsection{Footnotes}
-Number footnotes separately in superscript numbers.\footnote{It is recommended that footnotes be avoided (except for 
-the unnumbered footnote with the receipt date on the first page). Instead, 
-try to integrate the footnote information into the text.} Place the actual 
-footnote at the bottom of the column in which it is cited; do not put 
-footnotes in the reference list (endnotes). Use letters for table footnotes 
-(see Table \ref{table}).
-
-\section{Submitting Your Paper for Review}
-
-\subsection{Final Stage}
-When you submit your final version (after your paper has been accepted), 
-print it in two-column format, including figures and tables. You must also 
-send your final manuscript on a disk, via e-mail, or through a Web 
-manuscript submission system as directed by the society contact. You may use 
-\emph{Zip} for large files, or compress files using \emph{Compress, Pkzip, Stuffit,} or \emph{Gzip.}
-
-Also, send a sheet of paper or PDF with complete contact information for all 
-authors. Include full mailing addresses, telephone numbers, fax numbers, and 
-e-mail addresses. This information will be used to send each author a 
-complimentary copy of the journal in which the paper appears. In addition, 
-designate one author as the ``corresponding author.'' This is the author to 
-whom proofs of the paper will be sent. Proofs are sent to the corresponding 
-author only.
-
-\subsection{Review Stage Using ScholarOne\textregistered\ Manuscripts}
-Contributions to the Transactions, Journals, and Letters may be submitted 
-electronically on IEEE's on-line manuscript submission and peer-review 
-system, ScholarOne\textregistered\ Manuscripts. You can get a listing of the 
-publications that participate in ScholarOne at 
-\underline{http://www.ieee.org/publications\_standards/publications/}\discretionary{}{}{}\underline{authors/authors\_submission.html}.
-First check if you have an existing account. If there is none, please create 
-a new account. After logging in, go to your Author Center and click ``Submit 
-First Draft of a New Manuscript.''
-
-Along with other information, you will be asked to select the subject from a 
-pull-down list. Depending on the journal, there are various steps to the 
-submission process; you must complete all steps for a complete submission. 
-At the end of each step you must click ``Save and Continue''; just uploading 
-the paper is not sufficient. After the last step, you should see a 
-confirmation that the submission is complete. You should also receive an 
-e-mail confirmation. For inquiries regarding the submission of your paper on 
-ScholarOne Manuscripts, please contact oprs-support@ieee.org or call +1 732 
-465 5861.
-
-ScholarOne Manuscripts will accept files for review in various formats. 
-Please check the guidelines of the specific journal for which you plan to 
-submit.
-
-You will be asked to file an electronic copyright form immediately upon 
-completing the submission process (authors are responsible for obtaining any 
-security clearances). Failure to submit the electronic copyright could 
-result in publishing delays later. You will also have the opportunity to 
-designate your article as ``open access'' if you agree to pay the IEEE open 
-access fee.
-
-\subsection{Final Stage Using ScholarOne Manuscripts}
-Upon acceptance, you will receive an email with specific instructions 
-regarding the submission of your final files. To avoid any delays in 
-publication, please be sure to follow these instructions. Most journals 
-require that final submissions be uploaded through ScholarOne Manuscripts, 
-although some may still accept final submissions via email. Final 
-submissions should include source files of your accepted manuscript, high 
-quality graphic files, and a formatted pdf file. If you have any questions 
-regarding the final submission process, please contact the administrative 
-contact for the journal.
-
-In addition to this, upload a file with complete contact information for all 
-authors. Include full mailing addresses, telephone numbers, fax numbers, and 
-e-mail addresses. Designate the author who submitted the manuscript on 
-ScholarOne Manuscripts as the ``corresponding author.'' This is the only 
-author to whom proofs of the paper will be sent.
-
-\subsection{Copyright Form}
-Authors must submit an electronic IEEE Copyright Form (eCF) upon submitting 
-their final manuscript files. You can access the eCF system through your 
-manuscript submission system or through the Author Gateway. You are 
-responsible for obtaining any necessary approvals and/or security 
-clearances. For additional information on intellectual property rights, 
-visit the IEEE Intellectual Property Rights department web page at 
-\underline{http://www.ieee.org/publications\_standards/publications/rights/}\discretionary{}{}{}\underline{index.html}.
-
-\section{IEEE Publishing Policy}
-The general IEEE policy requires that authors should only submit original 
-work that has neither appeared elsewhere for publication, nor is under 
-review for another refereed publication. The submitting author must disclose 
-all prior publication(s) and current submissions when submitting a 
-manuscript. Do not publish ``preliminary'' data or results. The submitting 
-author is responsible for obtaining agreement of all coauthors and any 
-consent required from employers or sponsors before submitting an article. 
-The IEEE Transactions and Journals Department strongly discourages courtesy 
-authorship; it is the obligation of the authors to cite only relevant prior 
-work.
-
-The IEEE Transactions and Journals Department does not publish conference 
-records or proceedings, but can publish articles related to conferences that 
-have undergone rigorous peer review. Minimally, two reviews are required for 
-every article submitted for peer review.
-
-\section{Publication Principles}
-The two types of contents of that are published are; 1) peer-reviewed and 2) 
-archival. The Transactions and Journals Department publishes scholarly 
-articles of archival value as well as tutorial expositions and critical 
-reviews of classical subjects and topics of current interest.
-
-Authors should consider the following points:
-
-\begin{enumerate}
-\item Technical papers submitted for publication must advance the state of knowledge and must cite relevant prior work. 
-\item The length of a submitted paper should be commensurate with the importance, or appropriate to the complexity, of the work.
-For example, an obvious extension of previously published work might not be appropriate for publication or might be adequately treated in just a few pages.
-\item Authors must convince both peer reviewers and the editors of the scientific and technical merit of a paper; the standards of proof are higher when extraordinary or unexpected results are reported.
-\item Because replication is required for scientific progress, papers submitted for publication must provide sufficient information to allow readers to perform similar experiments or calculations and 
-use the reported results. Although not everything need be disclosed, a paper 
-must contain new, useable, and fully described information. For example, a 
-specimen's chemical composition need not be reported if the main purpose of 
-a paper is to introduce a new measurement technique. Authors should expect 
-to be challenged by reviewers if the results are not supported by adequate 
-data and critical details.
-\item Papers that describe ongoing work or announce the latest technical achievement, which are suitable for presentation at a professional conference, may not be appropriate for publication.
-\end{enumerate}
-
-\section{Reference Examples}
-
-\begin{itemize}
-
-\item \emph{Basic format for books:}\\
-J.
-K.
-Author, ``Title of chapter in the book,'' in \emph{Title of His Published Book, x}th ed.
-City of Publisher, (only U.S.
-State), Country: Abbrev.
-of Publisher, year, ch.
-$x$, sec.
-$x$, pp.
-\emph{xxx--xxx.}\\
-See \cite{b1,b2}.
-
-\item \emph{Basic format for periodicals:}\\
-J. K. Author, ``Name of paper,'' \emph{Abbrev. Title of Periodical}, vol. \emph{x, no}. $x, $pp\emph{. xxx--xxx, }Abbrev. Month, year, DOI. 10.1109.\emph{XXX}.123456.\\
-See \cite{b3}--\cite{b5}.
-
-\item \emph{Basic format for reports:}\\
-J. K. Author, ``Title of report,'' Abbrev. Name of Co., City of Co., Abbrev. State, Country, Rep. \emph{xxx}, year.\\
-See \cite{b6,b7}.
-
-\item \emph{Basic format for handbooks:}\\
-\emph{Name of Manual/Handbook, x} ed., Abbrev. Name of Co., City of Co., Abbrev. State, Country, year, pp. \emph{xxx--xxx.}\\
-See \cite{b8,b9}.
-
-\item \emph{Basic format for books (when available online):}\\
-J. K. Author, ``Title of chapter in the book,'' in \emph{Title of
-Published Book}, $x$th ed. City of Publisher, State, Country: Abbrev.
-of Publisher, year, ch. $x$, sec. $x$, pp. \emph{xxx--xxx}. [Online].
-Available: \underline{http://www.web.com}\\
-See \cite{b10}--\cite{b13}.
-
-\item \emph{Basic format for journals (when available online):}\\
-J.
-K.
-Author, ``Name of paper,'' \emph{Abbrev.
-Title of Periodical}, vol.
-$x$, no.
-$x$, pp.
-\emph{xxx--xxx}, Abbrev.
-Month, year.
-Accessed on: Month, Day, year, DOI: 10.1109.\emph{XXX}.123456, [Online].\\
-See \cite{b14}--\cite{b16}.
-
-\item \emph{Basic format for papers presented at conferences (when available online): }\\
-J.K. Author. (year, month). Title. presented at abbrev. conference title. [Type of Medium]. Available: site/path/file\\
-See \cite{b17}.
-
-\item \emph{Basic format for reports and handbooks (when available online):}\\
-J. K. Author. ``Title of report,'' Company. City, State, Country. Rep. no., (optional: vol./issue), Date. [Online] Available: site/path/file\\
-See \cite{b18,b19}.
-
-\item \emph{Basic format for computer programs and electronic documents (when available online): }\\
-Legislative body. Number of Congress, Session. (year, month day). \emph{Number of bill or resolution}, \emph{Title}. [Type of medium]. Available: site/path/file\\
-\textbf{\emph{NOTE: }ISO recommends that capitalization follow the accepted practice for the language or script in which the information is given.}\\
-See \cite{b20}.
-
-\item \emph{Basic format for patents (when available online):}\\
-Name of the invention, by inventor's name. (year, month day). Patent Number [Type of medium]. Available: site/path/file\\
-See \cite{b21}.
-
-\item \emph{Basic format}\emph{for conference proceedings (published):}\\
-J. K. Author, ``Title of paper,'' in \emph{Abbreviated Name of Conf.}, City of Conf., Abbrev. State (if given), Country, year, pp. \emph{xxxxxx.}\\
-See \cite{b22}.
-
-\item \emph{Example for papers presented at conferences (unpublished):}\\
-See \cite{b23}.
-
-\item \emph{Basic format for patents}$:$\\
-J. K. Author, ``Title of patent,'' U.S. Patent \emph{x xxx xxx}, Abbrev. Month, day, year.\\
-See \cite{b24}.
-
-\item \emph{Basic format for theses (M.S.) and dissertations (Ph.D.):}
-\begin{enumerate}
-\item J. K. Author, ``Title of thesis,'' M.S. thesis, Abbrev. Dept., Abbrev. Univ., City of Univ., Abbrev. State, year.
-\item J. K. Author, ``Title of dissertation,'' Ph.D. dissertation, Abbrev. Dept., Abbrev. Univ., City of Univ., Abbrev. State, year.
-\end{enumerate}
-See \cite{b25,b26}.
-
-\item \emph{Basic format for the most common types of unpublished references:}
-\begin{enumerate}
-\item J. K. Author, private communication, Abbrev. Month, year.
-\item J. K. Author, ``Title of paper,'' unpublished.
-\item J. K. Author, ``Title of paper,'' to be published.
-\end{enumerate}
-See \cite{b27}--\cite{b29}.
-
-\item \emph{Basic formats for standards:}
-\begin{enumerate}
-\item \emph{Title of Standard}, Standard number, date.
-\item \emph{Title of Standard}, Standard number, Corporate author, location, date.
-\end{enumerate}
-See \cite{b30,b31}.
-
-\item \emph{Article number in~reference examples:}\\
-See \cite{b32,b33}.
-
-\item \emph{Example when using et al.:}\\
-See \cite{b34}.
-
-\end{itemize}
 

### 2023-02-16 Manuscript_v1.tex
update

@@
 \usepackage{generic}
-\usepackage{cite}
+% \usepackage{cite}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 \usepackage{biblatex} %Imports biblatex package
-% \addbibresource{sample.bib} %Import the bibliography file
+\addbibresource{zotero.bib} %Import the bibliography file
 
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+%     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
 \markboth{\journalname, VOL. XX, NO. XX, XXXX 2017}
@@
 
-    Irregular @hornSetFunctionsTime2020 sampling of time series in electronic health records (EHRs) is challenging for model development.
+    Irregular  sampling of time series in electronic health records (EHRs) is challenging for model development.
 In addition, the pattern of missingness for certain clinical variables are not at random as it is determined by clinicians decision and the state of patient.
@@
 \begin{IEEEkeywords}
-    electronic health records (EHRs), Point Process, irregular sampling, informative missingness
+    electronic health records (EHRs), Point Process, irregular sampling, informative missingness,
 \end{IEEEkeywords}
@@
 We focus on tabular and time series.
+\cite{zuoTransformerHawkesProcess2020a}
 
@@
 
-conseq of irr sampling///
+conseq of irr sampling/// Irregularly sampled time series in electronic health records can have serious consequences for data analysis and decision making.
+Irregular sampling occurs when the time intervals between data points are inconsistent, which can lead to biased or inaccurate results when trying to make inferences about the underlying process.
+This can also make it difficult to accurately compare the data over time, as well as to detect patterns or trends.
+Furthermore, irregular sampling can cause problems for machine learning algorithms, which may struggle to accurately capture the relationships between variables in the data.
+This can result in poor predictions or incorrect diagnoses, potentially leading to negative impacts on patient care.
+To mitigate these consequences, it is important to ensure that electronic health records are sampled at regular intervals, or that appropriate statistical techniques are applied to account for the irregularity.
 
@@
 
-imputation methods///
+imputation methods/// Imputation techniques are widely used in electronic health records (EHRs) to handle missing data.
+EHRs often contain incomplete or missing information due to various reasons such as measurement errors, system failures, or data entry issues.
+Imputation refers to the process of filling in the missing data with plausible values based on the available information.
+There are several imputation techniques available, each with their own strengths and weaknesses.
+For example, mean imputation replaces missing values with the mean of the available values in the same column, while regression imputation uses a regression model to estimate the missing values based on the relationship between the variables.
+Other popular imputation methods include multiple imputation, hot deck imputation, and k-nearest neighbors imputation.
+The choice of imputation technique depends on the type and amount of missing data, as well as the goals of the analysis.
+Regardless of the technique used, it is important to carefully consider the impact of imputed values on the analysis and to report the imputation method used in the results.
 
-inherently compatible methods///
+inherently compatible methods/// Deep learning models that are compatible with irregularly sampled time series are important in many real-world applications, such as stock market predictions, speech recognition, and medical diagnosis.
+These models need to be able to handle data that is not evenly spaced, as is often the case in time-sensitive applications.
+One popular deep learning architecture that can handle irregularly sampled time series is the Recurrent Neural Network (RNN).
+RNNs are well-suited for this task because they have the ability to process sequential data, taking into account not only the current input but also the previous inputs.
+Another architecture that can handle irregularly sampled time series is the Convolutional Neural Network (CNN), which can be used to extract features from the data and then pass the processed data to an RNN for further analysis.
+Additionally, Attention Mechanisms can be integrated into RNNs and CNNs to help the model focus on important features in the data.
+Overall, these deep learning models offer a powerful toolset for handling irregularly sampled time series and provide useful insights into this type of data.
 
-pp framework///
+pp framework/// Point processes are mathematical models used to describe the distribution of events in time or space.
+These events can be occurrences of earthquakes, nerve impulses, customer purchases, or anything else that can be counted or measured.
+Neural Point Processes are a type of machine learning models that learn the patterns and relationships between events using artificial neural networks.
+These models can be used for tasks such as predicting future events, estimating the rate of event occurrence, or identifying correlations between events.
+They offer a flexible and powerful way to analyze point process data, as they can handle complex dependencies between events and incorporate prior knowledge about the process.
 
@@
 
-\subsubsection{Temporal point process}
-Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where each sample is represented as an event sequence  $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is number of event, $t_j$ is event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label).
-The history of events is denoted as $\mathcal{H}_t=\{(t_j,e_j):t_j<t \} $
+\subsubsection*{Temporal point process}
 
-In the multi-class setting where each mark can be exactly one of M marks, the point process is characterized by $M$ conidtional intensity functions (CIFs):
+Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where each sample is represented as an event sequence  $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is the total number of events, $t_j$ is event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label).
+The history of events is denoted as $\mathcal{H}_t=\{(t_j,e_j):t_j<t \}$.
 
-\begin{equation}\label{eq:11}
+The core idea of the point process framework is the definition of conditional intensity functions (CIFs) which is the probability of the occurence of an event of type $m$ in an infinitesimal time window $[t,t+dt)$:
+
+\begin{equation} 
    \lambda_m^{*}(t)=\lim \frac{P(  \text{event of type m in } [t,t+\Delta t) |   \mathcal{H}_t  )}{\Delta t} 
@@
 
-The default negative log-likelihood (NLL) of a point process for a single event sequnce $\mathcal{S}_{i}$ is defined as:
+Multivariate Hawkes process is the traditional approach to characterize CIFs:
 
-\begin{multline}\label{eq:11}
-    -\log p(\mathcal{S}_{i})  =  - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log \lambda_m^{*}(t_j) \\   
-    + \sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
+\begin{equation} 
+    \lambda_m^{*}(t)= \mu_m + \sum_{(t^{\prime},e^{\prime})\in \mathcal{H}_t} \phi(t-t^{\prime})
+ \end{equation}
+
+\subsubsection*{Neural temporal point process}
+
+Encoder-decoder architectures have proven to be effective in many applications.
+The main idea of a neural temporal point process (NTPP) is to first encode the history of events until $t_j$ using a neural network architecture $h_j=Enc(\mathcal{H}_{j+1};\theta)$.
+Then it tries to estimate $\lambda^*_m(t)=Dec(h_j;\phi)$ for $ t \in (t_j,t_{j+1}]$.
+
+Initial works used recurrent neural networks either in a discrete way [16], a continuous way [39], or a multi-scale multi-channel architecture [24]; then self-attention was also used by [54, 57].
+[57] develop a Structured Transformer Hawkes (STH) model that studies the relationship between point processes.
+It assumes the modeling of one point process might contribute to the modeling of another.
+Each vertex in the STH’s graph is associated with a point process.
+Other efforts have been made to model the cumulative intensity function [42] and conditional probability density [46].
+One issue that has been ignored in these works is that these NPP models take all types of historic events to predict the future ones, neglecting that some event types may not contribute to the prediction of another type.
+This study aims to correct this defect by learning to omit those non-contributing event types when predicting the target type via NPPs.
+
+\subsubsection*{Parameter Estimation}
+
+Based on conditional intensity function, it is straightforward to derive conditional probability density function $p^{*}_{m}(t)$ in $(t_j, t_{j+1}]$:
+
+\begin{equation} 
+    p_m^{*}(t)=\lambda_m^{*}(t) \exp \left[-\sum_{m=1}^{M} \int_{t_j}^{t_{j+1}}\lambda_m^{*}(t^{\prime})dt^{\prime}\right]
+ \end{equation}
+
+ Intuitively, the integral term indicates that there is no event of any type in the interval.
+
+In the multi-class setting, the log-likelihood (LL) of a point process for a single event sequnce $\mathcal{S}_{i}$ is defined as:
+
+\begin{multline} 
+    \log p_{mc}(\mathcal{S}_{i})  =  \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log p_m^{*}(t_j) \\   
+    = \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log \lambda_m^{*}(t_j)\\
+    \sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
 \end{multline}
 
-If we assume that marks and events are independent,
+However, in many cases such as EHRs it is common to have co-occuring events. [ntpp] proposed to add:
 
-\begin{multline}\label{eq:11}
-    -\log p(\mathcal{S}_{i})  = \\
-      - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log p^{*}(e_j=m) \lambda^{*}(t_j) \\   
-    + \sum_{m = 1}^{M} p^{*}(e_j=m) \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
+% If we assume that marks and events are independent,
+
+% \begin{multline} 
+%     -\log p(\mathcal{S}_{i})  = \\
+%       - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log p^{*}(e_j=m) \lambda^{*}(t_j) \\   
+%     + \sum_{m = 1}^{M} p^{*}(e_j=m) \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
+% \end{multline}
+
+\begin{multline} 
+    \log p_{ml}(\mathcal{S}_{i}) = \log p_{mc}(\mathcal{S}_{i}) \\
+    + \sum_{m = 1}^{M} (1-1(e_j=m)) \log \left(1-p_m^{*}(t_j)  \right)
 \end{multline}
 
-[ntpp] has proposed the following CIF to extend for multi-label case:
-
-\begin{multline}\label{eq:11}
-    -\log p(\mathcal{S}_{i})  =  - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log \lambda_m^{*}(t_j) \\   
-    + \sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)\\
-    - \sum_{j = 1}^{L}\sum_{m = 1}^{M} (1-1(e_j=m)) \log \left(1-\lambda_m^{*}(t_j) \exp \left( - \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right) \right)
-\end{multline}
+% \begin{multline} 
+%     -\log p(\mathcal{S}_{i})  =  - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log \lambda_m^{*}(t_j) \\   
+%     + \sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)\\
+%     - \sum_{j = 1}^{L}\sum_{m = 1}^{M} (1-1(e_j=m)) \log \left(1-\lambda_m^{*}(t_j) \exp \left( - \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right) \right)
+% \end{multline}
 
@@
 
-\subsubsection{Neural temporal point process}
+Another approach is the marked case which assumes that marks and timestamps are conditionally independent given $\mathcal{H}_t$.
 
-Encoder-decoder architectures have proven to be effective in many applications.
-The main idea of a neural point process is to first encode the history of events until $t_j$ using different neural network architectures $h_j=Enc(\mathcal{H}_j)$.
-Then it tries to estimate $\lambda^*_m(t)$ or its integral for $ t_i < t \leq t_{i+1}$.
+\begin{multline} 
+    \log p_{marked}(\mathcal{S}_{i}) =    
+    \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log p^{*}(e_j=m) \\
+    +\sum_{j = 1}^{L}\lambda^{*}(t_j) - \int_{t_0}^{t_{L}}\lambda^{*}(t^{\prime})dt^{\prime}  
+\end{multline}
 
-\subsubsection{Deep learning for irregular sampling}
+\subsubsection*{Deep learning for irregular sampling}
 An irregularly sampled data can be denoted as $\mathcal{D}=\{\mathcal{U}_i\}_{i=1}^{N}$ where $N$ is number of samples and each sample is represented as series of tuple $\mathcal{U}_i={(t_p,k_p,v_p)}$ where $t_p, k_p, v_p$ represents the time, modality and value of $p$-th datapoint respectively.
@@
 
- The key advantage of our proposed model is to combine a  transformer-based event encoder (TEE) with point process loss with a deep attention module (DAM) that can handle an irregularly sampled time series for a any downstream prediciton task.
-In this situation the data is represented as $\mathcal{D}=\{\mathcal{S}_i,\mathcal{U}_i\}_{i=1}^{N}$.
+ The key advantage of our proposed model is to combine a  transformer-based event encoder (TEE) with a deep attention module (DAM) that can handle an irregularly sampled time series for a any downstream prediciton task.
+In this case the data is represented as $\mathcal{D}=\{\mathcal{S}_i,\mathcal{U}_i\}_{i=1}^{N}$.
 General schematic of TEDAM is depicted in FIG.
@@
 
-In the first step, we embed all event marks $E_{emb}=E \times W_emb$ where $E_i \in \mathbb{R}^{L \times M}$ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{M \times d_{emb} }  $ is the trainable embedding matrix.
+In the first step, we embed all event marks $E_{emb}=E \times W_{emb}$ where $E_i \in \mathbb{R}^{L \times M}$ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{M \times d_{emb} }  $ is the trainable embedding matrix.
 
 In the second step, timestamps should be encoded and added to the event embedding, however, we propose to concatenate time encodings that can lead to better characterization of conditional intensity functions.
-Finally, the input of the transformer encoder will be $X_{emb}=[E_{emb}, T_{emb}] \in \mathbb{R}^{L \times (d_emb+t_emb)}$.
+Finally, the input of the transformer encoder will be $X_{emb}=[E_{emb}, T_{emb}] \in \mathbb{R}^{L \times (d_{emb}+t_{emb})}$.
 
 Here, we use the standard transformer encoder similar to [vaswani] with masking matrix to prevent the model from looking into the future.
-we obtain the encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $ where $h_j$ containes the all available information before occurence of $j$-th event.
+we obtain the encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $ where $h_j$ containes the all available information until occurence of $j$-th event.
 
@@
 Each side information  $ (t_k, v_k, m_k) $ can be represented by $s_k=(z(t_k), v_k, m_k)$.
-we define attention $a(\mathcal{S}_k,s_k )$
+we define attention $a(\mathcal{U}_k,s_k )$
 
-We define $\mathcal{S}_p$ to be the set of the first $p$ available information.
-The goal is to calculate $a(\mathcal{S}_p,s_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{S}_p$.
+We define $\mathcal{U}_p$ to be the set of the first $p$ available information.
+The goal is to calculate $a(\mathcal{U}_p,s_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{U}_p$.
 This is achieved by computing an embedding of the set elements using a smaller set functions $f^{\prime}$, and projecting the concatenation of the set representation and the individual set element into d-dimensional space:
-
-The goal is to calculate $a(\mathcal{S}_p,s_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{S}_p$.
-
-This is achieved by embedding set elements $\mathcal{S}_p$ using a simpler set function $f^{\prime}$ such as a cumulative mean.
 
@@
 \begin{equation*}
-    Z^{red}_i=Z_p where p=argmax(z \leq i)
+    Z^{red}_i=Z_p \text{ where } p=argmax(z \leq i)
 \end{equation*}
@@
 
-All formulas are:
+\subsection{All formulas}
 
-\begin{equation*}
-    \begin{cases}
+% \begin{equation*}
+%     \begin{cases}
 
-f^{\prime}(\mathcal{S}_p) = g^{\prime} \left(   
- \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) 
-\right) \\
+% f^{\prime}(\mathcal{S}_p) = g^{\prime} \left(   
+%  \frac{1}{p} \sum_{s_k \in \mathcal{S}_p}  h^{\prime}_{\theta}(s_k) 
+% \right) \\
 
-K_p=[f^{\prime}(\mathcal{S}_p), s_p]^T W^K \\
-e_p=\frac{K_p.w^q}{\sqrt{d}  }\\
+% K_p=[f^{\prime}(\mathcal{S}_p), s_p]^T W^K \\
+% e_p=\frac{K_p.w^q}{\sqrt{d}  }\\
 
-a(\mathcal{S}_p,s_k)=\frac{\exp (e_p)}{\sum_{k\leq p} \exp (e_k)  } \\
+% a(\mathcal{S}_p,s_k)=\frac{\exp (e_p)}{\sum_{k\leq p} \exp (e_k)  } \\
 
-f(\mathcal{S}_p) = \sum_{s_k \in \mathcal{S}_p}  a(\mathcal{S}_p,s_k)h_{\theta}(s_k) \\
+% f(\mathcal{S}_p) = \sum_{s_k \in \mathcal{S}_p}  a(\mathcal{S}_p,s_k)h_{\theta}(s_k) \\
 
-z_p=g_{\psi}\left(         f(\mathcal{S}_p)         \right) \\
+% z_p=g_{\psi}\left(         f(\mathcal{S}_p)         \right) \\
 
-\end{cases} 
-\end{equation*}
+% \end{cases} 
+% \end{equation*}
 
-\begin{equation}\label{eq:1}
-    [z(t_j)]_{k} = 
-     \begin{cases} 
-        \sin\left(  \frac{t_j}{\mathcal{T}^{(k-1)/d_t} }  \right) & \text{if } k \text{ is odd}\\
-        \sin\left(  \frac{t_j}{\mathcal{T}^{k/d_t} }  \right) & \text{if } k \text{ is even}
+% \begin{equation}\label{eq:1}
+%     [z(t_j)]_{k} = 
+%      \begin{cases} 
+%         \sin\left(  \frac{t_j}{\mathcal{T}^{(k-1)/d_t} }  \right) & \text{if } k \text{ is odd}\\
+%         \sin\left(  \frac{t_j}{\mathcal{T}^{k/d_t} }  \right) & \text{if } k \text{ is even}
 
-        \end{cases} 
-\end{equation} 
-\begin{equation}
-    \lambda_k(t|\mathcal{H}_t ) = f_k\left(\alpha_k \frac{t-t_j}{t_j}+\mathbf{w}_k^T \mathbf{x}_{enc}(t_j)+\mathbf{y}_k^T \mathbf{s}_{enc}(t_j) +b_k   \right)
-\end{equation}
+%         \end{cases} 
+% \end{equation} 
+% \begin{equation}
+%     \lambda_k(t|\mathcal{H}_t ) = f_k\left(\alpha_k \frac{t-t_j}{t_j}+\mathbf{w}_k^T \mathbf{x}_{enc}(t_j)+\mathbf{y}_k^T \mathbf{s}_{enc}(t_j) +b_k   \right)
+% \end{equation}
 
-\begin{equation}
-    \lambda_k(t|\mathcal{H}_t ) = f_k\left(\alpha_k \frac{t-t_j}{t_j}+\mathbf{w}_k^T \mathbf{x}_{enc}(t_j)+b_k   \right)
-\end{equation}
+% \begin{equation}
+%     \lambda_k(t|\mathcal{H}_t ) = f_k\left(\alpha_k \frac{t-t_j}{t_j}+\mathbf{w}_k^T \mathbf{x}_{enc}(t_j)+b_k   \right)
+% \end{equation}
 
@@
 \begin{multline*}
-    \lambda_m(t)=softplus(\mu_{m,i+1}+\\
+    \lambda_m(t)=\text{softplus}(\mu_{m,i+1}+\\
     (\eta_{m,i+1}- \mu_{m,i+1}) \exp(-\gamma_{m,i+1}(t-t_i))    ),
@@
 
-\textbf{Physionet 2012 Mortality Prediction Challenge (\emph{P19})}.
+\textbf{Physionet 2012 Mortality Prediction Challenge (\emph{P12})}.
 The 2012 Physionet challenge dataset (Goldberger et al., 2000),contains 12, 000 ICU stays each of which lasts at least 48 h.
@@
 
-effect of time concatenation
+[TB1] shows the performance metrics of different datasets.
 
-compare single+mark with mc or ml
+In general, time concatenation has lead to better F1-score/AUROC as well as LL/events compared to adding time. This idea can be used for future research in NTPPs for better characterization of CIFs.
+
+Another interesting fact is that in some datasets (A, B) next mark prediction has competetive performance compared to the point process loss.
+However, in many cases, point process loss significantly improves the performance metrics which indicates that modeling non-event likelihood could be beneficial.
+In addition, we propose to use next-mark prediction case as a baseline.
+
+Another interesting results is that independent assumption in the marked loss may lead to worse performance compared to MC/ML loss.
 
@@
   \caption{Add caption}
-      \begin{tabular}{cccccc}
+    \begin{tabular}{cccccccccccccc}
     \toprule
-            &       & \multicolumn{2}{c}{TEEDAM} &       &  \\
-  \cmidrule{3-4}    Dataset & Metric & concat & sum   & SAHP  & GRU-CP \\
+          &       & \multicolumn{8}{c}{TEEDAM}                                    &       &       &       &  \\
+\cmidrule{3-10}          &       & \multicolumn{2}{c}{AE (next mark)} &       & \multicolumn{2}{c}{PP(single+mark)} &       & \multicolumn{2}{c}{PP (MC/ML)} &       &       &       &  \\
+\cmidrule{3-4}\cmidrule{6-7}\cmidrule{9-10}    Dataset & Metric & concat & sum   &       & concat & sum   &       & concat & sum   & Latent & SAHP  & THP   & GRU-CP \\
     \midrule
-      \multirow{2}[2]{*}{SO} & LL/\#events &       &       &       &  \\
-            & AUROC &       &       &       &  \\
+    \multirow{2}[2]{*}{SO} & LL/\#events & ND    & ND    &       & -0.56 & -0.57 &       &       &       & -1.54 & -1.86 & -1.84 & NR \\
+          & F1-score & 27.86 & 28.79 &       & 29.10 & 28.95 &       &       &       & 28.34(0.19) & 24.12 & 23.89 & 26 \\
     \midrule
-      \multirow{2}[2]{*}{Synthea} & LL/\#events &       &       &       &  \\
-            & AUROC &       &       &       & 0.85(.014) \\
+    \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & ND    & ND    &       &       &       &       &       &       & -3.89 & -4.56 & -4.57 & NR \\
+          & F1-score &       &       &       &       &       &       &       &       & 58.29 & 53.92 & 53.86 & NR \\
     \midrule
-      \multirow{2}[2]{*}{ReTweet} & LL/\#events & 1.83 (0.09) & 1.5 (0.18) &       &  \\
-            & AUROC & 0.68 (0.03) & 0.67 (0.05) &       & 0.611(0.001) \\
+    \multirow{2}[2]{*}{Synthea} & LL/\#events & ND    & ND    &       &       &       &       &       &       & ND    & ND    & ND    & NR \\
+          & AUROC &       &       &       &       &       &       &       &       & ND    & ND    & ND    & 0.85(.014) \\
+    \midrule
+    \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & ND    & ND    &       & 1.86  & 1.4   &       &       &       & ND    & ND    & ND    & NR \\
+          & AUROC & 63.3  & 61.01 &       & 68.7  & 66.2  &       & 73.87 &       & ND    & ND    & ND    & 0.611(0.001) \\
     \bottomrule
@@
 
-Table 1 shows the result for estimation of negative likelihood in different datasets and scnearios. It is obvious that state encoding has led to lower NLL.
+[TB2] shows the result for estimation of negative likelihood for the two datasets.
+
+In general we can see that using state information results in lower NLL, however, it cannot neccessarily increase the AUROC for next time prediction.
+
+Another observation is that more data will lead to higher LL.
 
 % Table generated by Excel2LaTeX from sheet 'state enc'
-\begin{table}[h!]
+\begin{table}[htbp]
     \centering
@@
       \midrule
-      \multirow{3}[2]{*}{P12} & sc    & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & mc1   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & mc2   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \multirow{3}[2]{*}{P12} & sc    & -1.99 & -1.77 & -1.92 \\
+            & mc1   & -0.99 & -1.14 & -0.89 \\
+            & mc2   & -1.5  & -1.33 & -1.47 \\
       \midrule
-      \multirow{3}[2]{*}{p19} & sc    & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & mc1   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & mc2   & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+      \multirow{3}[2]{*}{P19} & sc    & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
+            & mc1   & -1.81 & -1.62 & -1.69 \\
+            & mc2   & -2.77 & -2.38 & -2.67 \\
       \bottomrule
@@
 
-\subsection{Downstream task with event encoding}
+\subsection{Downstream task}
 
-Another key element of our work is to show the effectiveness of point process modeling for a down-stream task.
-In Table, we have reported the performance metrics for the mortality prediction task.
-We have compared our results with several sota's DL models that are compatible with irregular time series.
+[TB3] indicates the result for Mortality/sepsis prediction in different settings and hospital centers.
+
+In some cases, incorporating event embeddings can lead to better performance.
 
@@
 
-\begin{thebibliography}{00}
+% \begin{thebibliography}{00}
 
-\bibitem{b1} G. O. Young, ``Synthetic structure of industrial plastics,'' in \emph{Plastics,} 2\textsuperscript{nd} ed., vol. 3, J. Peters, Ed. New York, NY, USA: McGraw-Hill, 1964, pp. 15--64.
+% \bibitem{b1} G. O. Young, ``Synthetic structure of industrial plastics,'' in \emph{Plastics,} 2\textsuperscript{nd} ed., vol. 3, J. Peters, Ed. New York, NY, USA: McGraw-Hill, 1964, pp. 15--64.
 
-\bibitem{b2} W.-K. Chen, \emph{Linear Networks and Systems.} Belmont, CA, USA: Wadsworth, 1993, pp. 123--135.
+% \bibitem{b2} W.-K. Chen, \emph{Linear Networks and Systems.} Belmont, CA, USA: Wadsworth, 1993, pp. 123--135.
 
-\bibitem{b3} J. U. Duncombe, ``Infrared navigation---Part I: An assessment of feasibility,'' \emph{IEEE Trans. Electron Devices}, vol. ED-11, no. 1, pp. 34--39, Jan. 1959, 10.1109/TED.2016.2628402.
+% \bibitem{b3} J. U. Duncombe, ``Infrared navigation---Part I: An assessment of feasibility,'' \emph{IEEE Trans. Electron Devices}, vol. ED-11, no. 1, pp. 34--39, Jan. 1959, 10.1109/TED.2016.2628402.
 
-\bibitem{b4} E. P. Wigner, ``Theory of traveling-wave optical laser,'' \emph{Phys. Rev}., vol. 134, pp. A635--A646, Dec. 1965.
+% \bibitem{b4} E. P. Wigner, ``Theory of traveling-wave optical laser,'' \emph{Phys. Rev}., vol. 134, pp. A635--A646, Dec. 1965.
 
-\bibitem{b5} E. H. Miller, ``A note on reflector arrays,'' \emph{IEEE Trans. Antennas Propagat}., to be published.
+% \bibitem{b5} E. H. Miller, ``A note on reflector arrays,'' \emph{IEEE Trans. Antennas Propagat}., to be published.
 
-\bibitem{b6} E. E. Reber, R. L. Michell, and C. J. Carter, ``Oxygen absorption in the earth's atmosphere,'' Aerospace Corp., Los Angeles, CA, USA, Tech. Rep. TR-0200 (4230-46)-3, Nov. 1988.
+% \bibitem{b6} E. E. Reber, R. L. Michell, and C. J. Carter, ``Oxygen absorption in the earth's atmosphere,'' Aerospace Corp., Los Angeles, CA, USA, Tech. Rep. TR-0200 (4230-46)-3, Nov. 1988.
 
-\bibitem{b7} J. H. Davis and J. R. Cogdell, ``Calibration program for the 16-foot antenna,'' Elect. Eng. Res. Lab., Univ. Texas, Austin, TX, USA, Tech. Memo. NGL-006-69-3, Nov. 15, 1987.
+% \bibitem{b7} J. H. Davis and J. R. Cogdell, ``Calibration program for the 16-foot antenna,'' Elect. Eng. Res. Lab., Univ. Texas, Austin, TX, USA, Tech. Memo. NGL-006-69-3, Nov. 15, 1987.
 
-\bibitem{b8} \emph{Transmission Systems for Communications}, 3\textsuperscript{rd} ed., Western Electric Co., Winston-Salem, NC, USA, 1985, pp. 44--60.
+% \bibitem{b8} \emph{Transmission Systems for Communications}, 3\textsuperscript{rd} ed., Western Electric Co., Winston-Salem, NC, USA, 1985, pp. 44--60.
 
-\bibitem{b9} \emph{Motorola Semiconductor Data Manual}, Motorola Semiconductor Products Inc., Phoenix, AZ, USA, 1989.
+% \bibitem{b9} \emph{Motorola Semiconductor Data Manual}, Motorola Semiconductor Products Inc., Phoenix, AZ, USA, 1989.
 
-\bibitem{b10} G. O. Young, ``Synthetic structure of industrial
-plastics,'' in Plastics, vol. 3, Polymers of Hexadromicon, J. Peters,
-Ed., 2\textsuperscript{nd} ed. New York, NY, USA: McGraw-Hill, 1964, pp. 15-64.
-[Online]. Available:
-\underline{http://www.bookref.com}.
+% \bibitem{b10} G. O. Young, ``Synthetic structure of industrial
+% plastics,'' in Plastics, vol. 3, Polymers of Hexadromicon, J. Peters,
+% Ed., 2\textsuperscript{nd} ed. New York, NY, USA: McGraw-Hill, 1964, pp. 15-64.
+% [Online]. Available:
+% \underline{http://www.bookref.com}.
 
-\bibitem{b11} \emph{The Founders' Constitution}, Philip B. Kurland
-and Ralph Lerner, eds., Chicago, IL, USA: Univ. Chicago Press, 1987.
-[Online]. Available: \underline{http://press-pubs.uchicago.edu/founders/}
+% \bibitem{b11} \emph{The Founders' Constitution}, Philip B. Kurland
+% and Ralph Lerner, eds., Chicago, IL, USA: Univ. Chicago Press, 1987.
+% [Online]. Available: \underline{http://press-pubs.uchicago.edu/founders/}
 
-\bibitem{b12} The Terahertz Wave eBook. ZOmega Terahertz Corp., 2014.
-[Online]. Available:
-\underline{http://dl.z-thz.com/eBook/zomega\_ebook\_pdf\_1206\_sr.pdf}. Accessed on: May 19, 2014.
+% \bibitem{b12} The Terahertz Wave eBook. ZOmega Terahertz Corp., 2014.
+% [Online]. Available:
+% \underline{http://dl.z-thz.com/eBook/zomega\_ebook\_pdf\_1206\_sr.pdf}. Accessed on: May 19, 2014.
 
-\bibitem{b13} Philip B. Kurland and Ralph Lerner, eds., \emph{The
-Founders' Constitution.} Chicago, IL, USA: Univ. of Chicago Press,
-1987, Accessed on: Feb. 28, 2010, [Online] Available:
-\underline{http://press-pubs.uchicago.edu/founders/}
+% \bibitem{b13} Philip B. Kurland and Ralph Lerner, eds., \emph{The
+% Founders' Constitution.} Chicago, IL, USA: Univ. of Chicago Press,
+% 1987, Accessed on: Feb. 28, 2010, [Online] Available:
+% \underline{http://press-pubs.uchicago.edu/founders/}
 
-\bibitem{b14} J. S. Turner, ``New directions in communications,'' \emph{IEEE J. Sel. Areas Commun}., vol. 13, no. 1, pp. 11-23, Jan. 1995.
+% \bibitem{b14} J. S. Turner, ``New directions in communications,'' \emph{IEEE J. Sel. Areas Commun}., vol. 13, no. 1, pp. 11-23, Jan. 1995.
 
-\bibitem{b15} W.
+% \bibitem{b15} W.
 P.
@@
 
-\bibitem{b16} P.
+% \bibitem{b16} P.
 Kopyt \emph{et al., ``}Electric properties of graphene-based conductive layers from DC up to terahertz range,'' \emph{IEEE THz Sci.
@@
 
-\bibitem{b17} PROCESS Corporation, Boston, MA, USA. Intranets:
-Internet technologies deployed behind the firewall for corporate
-productivity. Presented at INET96 Annual Meeting. [Online].
-Available: \underline{http://home.process.com/Intranets/wp2.htp}
+% \bibitem{b17} PROCESS Corporation, Boston, MA, USA. Intranets:
+% Internet technologies deployed behind the firewall for corporate
+% productivity. Presented at INET96 Annual Meeting. [Online].
+% Available: \underline{http://home.process.com/Intranets/wp2.htp}
 
-\bibitem{b18} R.
+% \bibitem{b18} R.
 J.
@@
 
-\bibitem{b19} Teralyzer. Lytera UG, Kirchhain, Germany [Online].
-Available:
-\underline{http://www.lytera.de/Terahertz\_THz\_Spectroscopy.php?id=home}, Accessed on: Jun. 5, 2014
+% \bibitem{b19} Teralyzer. Lytera UG, Kirchhain, Germany [Online].
+% Available:
+% \underline{http://www.lytera.de/Terahertz\_THz\_Spectroscopy.php?id=home}, Accessed on: Jun. 5, 2014
 
-\bibitem{b20} U.S.
+% \bibitem{b20} U.S.
 House.
@@
 
-\bibitem{b21} Musical toothbrush with mirror, by L.M.R. Brooks. (1992, May 19). Patent D 326 189 [Online]. Available: NEXIS Library: LEXPAT File: DES
+% \bibitem{b21} Musical toothbrush with mirror, by L.M.R. Brooks. (1992, May 19). Patent D 326 189 [Online]. Available: NEXIS Library: LEXPAT File: DES
 
-\bibitem{b22} D. B. Payne and J. R. Stern, ``Wavelength-switched pas- sively coupled single-mode optical network,'' in \emph{Proc. IOOC-ECOC,} Boston, MA, USA, 1985, pp. 585--590.
+% \bibitem{b22} D. B. Payne and J. R. Stern, ``Wavelength-switched pas- sively coupled single-mode optical network,'' in \emph{Proc. IOOC-ECOC,} Boston, MA, USA, 1985, pp. 585--590.
 
-\bibitem{b23} D.
+% \bibitem{b23} D.
 Ebehard and E.
@@
 
-\bibitem{b24} G. Brandli and M. Dick, ``Alternating current fed power supply,'' U.S. Patent 4 084 217, Nov. 4, 1978.
+% \bibitem{b24} G. Brandli and M. Dick, ``Alternating current fed power supply,'' U.S. Patent 4 084 217, Nov. 4, 1978.
 
-\bibitem{b25} J. O. Williams, ``Narrow-band analyzer,'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, USA, 1993.
+% \bibitem{b25} J. O. Williams, ``Narrow-band analyzer,'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, USA, 1993.
 
-\bibitem{b26} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
+% \bibitem{b26} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
 
-\bibitem{b27} A. Harrison, private communication, May 1995.
+% \bibitem{b27} A. Harrison, private communication, May 1995.
 
-\bibitem{b28} B. Smith, ``An approach to graphs of linear forms,'' unpublished.
+% \bibitem{b28} B. Smith, ``An approach to graphs of linear forms,'' unpublished.
 
-\bibitem{b29} A. Brahms, ``Representation error for real numbers in binary computer arithmetic,'' IEEE Computer Group Repository, Paper R-67-85.
+% \bibitem{b29} A. Brahms, ``Representation error for real numbers in binary computer arithmetic,'' IEEE Computer Group Repository, Paper R-67-85.
 
-\bibitem{b30} IEEE Criteria for Class IE Electric Systems, IEEE Standard 308, 1969.
+% \bibitem{b30} IEEE Criteria for Class IE Electric Systems, IEEE Standard 308, 1969.
 
-\bibitem{b31} Letter Symbols for Quantities, ANSI Standard Y10.5-1968.
+% \bibitem{b31} Letter Symbols for Quantities, ANSI Standard Y10.5-1968.
 
-\bibitem{b32} R.
+% \bibitem{b32} R.
 Fardel, M.
@@
 
-\bibitem{b33} J. Zhang and N. Tansu, ``Optical gain and laser characteristics of InGaN quantum wells on ternary InGaN substrates,'' \emph{IEEE Photon. J.}, vol. 5, no. 2, Apr. 2013, Art. no. 2600111
+% \bibitem{b33} J. Zhang and N. Tansu, ``Optical gain and laser characteristics of InGaN quantum wells on ternary InGaN substrates,'' \emph{IEEE Photon. J.}, vol. 5, no. 2, Apr. 2013, Art. no. 2600111
 
-\bibitem{b34} S.
+% \bibitem{b34} S.
 Azodolmolky~\emph{et al.}, Experimental demonstration of an impairment aware network planning and operation tool for transparent/translucent optical networks,''~\emph{J.
@@
 
-\end{thebibliography}
+% \end{thebibliography}
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a1.png}}]{First A. Author} (M'76--SM'81--F'87) and all authors may include 
-biographies. Biographies are often not included in conference-related
-papers. This author became a Member (M) of IEEE in 1976, a Senior
-Member (SM) in 1981, and a Fellow (F) in 1987. The first paragraph may
-contain a place and/or date of birth (list place, then date). Next,
-the author's educational background is listed. The degrees should be
-listed with type of degree in what field, which institution, city,
-state, and country, and year the degree was earned. The author's major
-field of study should be lower-cased.
+\printbibliography
 
-The second paragraph uses the pronoun of the person (he or she) and not the 
-author's last name. It lists military and work experience, including summer 
-and fellowship jobs. Job titles are capitalized. The current job must have a 
-location; previous positions may be listed 
-without one. Information concerning previous publications may be included. 
-Try not to list more than three books or published articles. The format for 
-listing publishers of a book within the biography is: title of book 
-(publisher name, year) similar to a reference. Current and previous research 
-interests end the paragraph. The third paragraph begins with the author's 
-title and last name (e.g., Dr.\ Smith, Prof.\ Jones, Mr.\ Kajor, Ms.\ Hunter). 
-List any memberships in professional societies other than the IEEE. Finally, 
-list any awards and work for IEEE committees and publications. If a 
-photograph is provided, it should be of good quality, and 
-professional-looking. Following are two examples of an author's biography.
-\end{IEEEbiography}
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a1.png}}]{First A. Author} (M'76--SM'81--F'87) and all authors may include 
+% biographies. Biographies are often not included in conference-related
+% papers. This author became a Member (M) of IEEE in 1976, a Senior
+% Member (SM) in 1981, and a Fellow (F) in 1987. The first paragraph may
+% contain a place and/or date of birth (list place, then date). Next,
+% the author's educational background is listed. The degrees should be
+% listed with type of degree in what field, which institution, city,
+% state, and country, and year the degree was earned. The author's major
+% field of study should be lower-cased.
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a2.png}}]{Second B. Author} was born in Greenwich Village, New York, NY, USA in 
-1977. He received the B.S. and M.S. degrees in aerospace engineering from 
-the University of Virginia, Charlottesville, in 2001 and the Ph.D. degree in 
-mechanical engineering from Drexel University, Philadelphia, PA, in 2008.
+% The second paragraph uses the pronoun of the person (he or she) and not the 
+% author's last name. It lists military and work experience, including summer 
+% and fellowship jobs. Job titles are capitalized. The current job must have a 
+% location; previous positions may be listed 
+% without one. Information concerning previous publications may be included. 
+% Try not to list more than three books or published articles. The format for 
+% listing publishers of a book within the biography is: title of book 
+% (publisher name, year) similar to a reference. Current and previous research 
+% interests end the paragraph. The third paragraph begins with the author's 
+% title and last name (e.g., Dr.\ Smith, Prof.\ Jones, Mr.\ Kajor, Ms.\ Hunter). 
+% List any memberships in professional societies other than the IEEE. Finally, 
+% list any awards and work for IEEE committees and publications. If a 
+% photograph is provided, it should be of good quality, and 
+% professional-looking. Following are two examples of an author's biography.
+% \end{IEEEbiography}
 
-From 2001 to 2004, he was a Research Assistant with the Princeton Plasma 
-Physics Laboratory. Since 2009, he has been an Assistant Professor with the 
-Mechanical Engineering Department, Texas A{\&}M University, College Station. 
-He is the author of three books, more than 150 articles, and more than 70 
-inventions. His research interests include high-pressure and high-density 
-nonthermal plasma discharge processes and applications, microscale plasma 
-discharges, discharges in liquids, spectroscopic diagnostics, plasma 
-propulsion, and innovation plasma applications. He is an Associate Editor of 
-the journal \emph{Earth, Moon, Planets}, and holds two patents.
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a2.png}}]{Second B. Author} was born in Greenwich Village, New York, NY, USA in 
+% 1977. He received the B.S. and M.S. degrees in aerospace engineering from 
+% the University of Virginia, Charlottesville, in 2001 and the Ph.D. degree in 
+% mechanical engineering from Drexel University, Philadelphia, PA, in 2008.
 
-Dr. Author was a recipient of the International Association of Geomagnetism 
-and Aeronomy Young Scientist Award for Excellence in 2008, and the IEEE 
-Electromagnetic Compatibility Society Best Symposium Paper Award in 2011. 
-\end{IEEEbiography}
+% From 2001 to 2004, he was a Research Assistant with the Princeton Plasma 
+% Physics Laboratory. Since 2009, he has been an Assistant Professor with the 
+% Mechanical Engineering Department, Texas A{\&}M University, College Station. 
+% He is the author of three books, more than 150 articles, and more than 70 
+% inventions. His research interests include high-pressure and high-density 
+% nonthermal plasma discharge processes and applications, microscale plasma 
+% discharges, discharges in liquids, spectroscopic diagnostics, plasma 
+% propulsion, and innovation plasma applications. He is an Associate Editor of 
+% the journal \emph{Earth, Moon, Planets}, and holds two patents.
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a3.png}}]{Third C. Author, Jr.} (M'87) received the B.S. degree in mechanical 
-engineering from National Chung Cheng University, Chiayi, Taiwan, in 2004 
-and the M.S. degree in mechanical engineering from National Tsing Hua 
-University, Hsinchu, Taiwan, in 2006. He is currently pursuing the Ph.D. 
-degree in mechanical engineering at Texas A{\&}M University, College 
-Station, TX, USA.
+% Dr. Author was a recipient of the International Association of Geomagnetism 
+% and Aeronomy Young Scientist Award for Excellence in 2008, and the IEEE 
+% Electromagnetic Compatibility Society Best Symposium Paper Award in 2011. 
+% \end{IEEEbiography}
 
-From 2008 to 2009, he was a Research Assistant with the Institute of 
-Physics, Academia Sinica, Tapei, Taiwan. His research interest includes the 
-development of surface processing and biological/medical treatment 
-techniques using nonthermal atmospheric pressure plasmas, fundamental study 
-of plasma sources, and fabrication of micro- or nanostructured surfaces.
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{a3.png}}]{Third C. Author, Jr.} (M'87) received the B.S. degree in mechanical 
+% engineering from National Chung Cheng University, Chiayi, Taiwan, in 2004 
+% and the M.S. degree in mechanical engineering from National Tsing Hua 
+% University, Hsinchu, Taiwan, in 2006. He is currently pursuing the Ph.D. 
+% degree in mechanical engineering at Texas A{\&}M University, College 
+% Station, TX, USA.
 
-Mr. Author's awards and honors include the Frew Fellowship (Australian 
-Academy of Science), the I. I. Rabi Prize (APS), the European Frequency and 
-Time Forum Award, the Carl Zeiss Research Award, the William F. Meggers 
-Award and the Adolph Lomb Medal (OSA).
-\end{IEEEbiography}
+% From 2008 to 2009, he was a Research Assistant with the Institute of 
+% Physics, Academia Sinica, Tapei, Taiwan. His research interest includes the 
+% development of surface processing and biological/medical treatment 
+% techniques using nonthermal atmospheric pressure plasmas, fundamental study 
+% of plasma sources, and fabrication of micro- or nanostructured surfaces.
+
+% Mr. Author's awards and honors include the Frew Fellowship (Australian 
+% Academy of Science), the I. I. Rabi Prize (APS), the European Frequency and 
+% Time Forum Award, the Carl Zeiss Research Award, the William F. Meggers 
+% Award and the Adolph Lomb Medal (OSA).
+% \end{IEEEbiography}
 

### 2023-03-02 Manuscript_v1.tex
update

@@
 \usepackage{amsmath,amssymb,amsfonts}
+\usepackage{bbm}
 \usepackage{algorithmic}
@@
 \usepackage{textcomp}
+
+\DeclareMathOperator*{\argmax}{argmax} % thin space, limits underneath in displays
+
+\usepackage{svg}
 
@@
 
-healthcare systems are one of main area for deploying machine learning models.
-Thanks to recent advancements in data collection technologies, policies and computing power data are recorded in a hosptial.
-This data consists of multiple modalities frome tabular and time series to clinical notes and images.
-We focus on tabular and time series.
-\cite{zuoTransformerHawkesProcess2020a}
+[ML in healthcare]Machine learning has the potential to revolutionize healthcare by leveraging vast amounts of data available in electronic health records (EHRs) to develop more accurate clinical decision support systems.
+EHRs store patient health information, such as medical history, medications, lab results, and diagnostic images, which can be used as input for machine learning algorithms to identify patterns and associations that could inform more precise diagnoses, better treatment plans, and earlier interventions.
+Clinical decision support systems that use machine learning can provide real-time, evidence-based recommendations to healthcare providers, reducing errors and improving patient outcomes.
 
-talk about the irregular sampling in EHRs and causes.///
-In and ideal setup, variables are measured on a regular basis, however, in a hospital due to medical costs, patient associated risks, sampling device.
+[irregular sampling challenge] One of data challenges for machine learning (ML) when using electronic health records (EHRs) is irregular sampling.
+EHR data is often collected at different times and frequencies, depending on a patient's healthcare needs and visit schedules, which can result in uneven and irregularly sampled time series data.
 
-conseq of irr sampling/// Irregularly sampled time series in electronic health records can have serious consequences for data analysis and decision making.
-Irregular sampling occurs when the time intervals between data points are inconsistent, which can lead to biased or inaccurate results when trying to make inferences about the underlying process.
-This can also make it difficult to accurately compare the data over time, as well as to detect patterns or trends.
-Furthermore, irregular sampling can cause problems for machine learning algorithms, which may struggle to accurately capture the relationships between variables in the data.
-This can result in poor predictions or incorrect diagnoses, potentially leading to negative impacts on patient care.
-To mitigate these consequences, it is important to ensure that electronic health records are sampled at regular intervals, or that appropriate statistical techniques are applied to account for the irregularity.
+missingness in tabular and time series.
 
-potential adv of irr sampl///
+[sources of missingness] Sources of missingness in EHRs must be carefully understood.
+For example, lab measurements are typically ordered as part of a diagnostic work-up, meaning that the presence of a datapoint conveys information about the patient's state
+% \cite*{ghassemiReviewChallengesOpportunities2020}.
+
+[conseqs of IRR] A typical EHR consist of tabular and time series data.
+Depending on the desired outcome, we can formulate the problem in static or longitudinal setting.
+There are various consequences asscociated with irregularly sampled data.
+Many machine learning models are not inherently compatible with
+
+imputation for static setting
+
+imputation for time series.
+
+inherently compatible methods.
 
 imputation methods/// Imputation techniques are widely used in electronic health records (EHRs) to handle missing data.
-EHRs often contain incomplete or missing information due to various reasons such as measurement errors, system failures, or data entry issues.
 Imputation refers to the process of filling in the missing data with plausible values based on the available information.
@@
 
-inherently compatible methods/// Deep learning models that are compatible with irregularly sampled time series are important in many real-world applications, such as stock market predictions, speech recognition, and medical diagnosis.
+There exist several ML frameworks that are inherently compatible with the missig or irregularly-sampled data such as Gaussian process and recurrent neural networks.
+Deep learning models that are compatible with irregularly sampled time series are important in many real-world applications, such as stock market predictions, speech recognition, and medical diagnosis.
 These models need to be able to handle data that is not evenly spaced, as is often the case in time-sensitive applications.
@@
 Overall, these deep learning models offer a powerful toolset for handling irregularly sampled time series and provide useful insights into this type of data.
+These models does not explicitly take into account the fact wether the absence of some datapoints could convey some unique information.
 
-pp framework/// Point processes are mathematical models used to describe the distribution of events in time or space.
+Point processes are mathematical models used to describe the distribution of events in time or space.
 These events can be occurrences of earthquakes, nerve impulses, customer purchases, or anything else that can be counted or measured.
-Neural Point Processes are a type of machine learning models that learn the patterns and relationships between events using artificial neural networks.
+In healthcare, lab measurements can be regarded as sequence of events that are ordered by clinicians.
+At the core of point process is the defintion of conditional intensity function and the corresponding log likelihood which simultanously models the occurence of events using the history of past events.[GAP]
+More recently, neural Point Processes have been developed for better characteriztion of CIFs by leveraing the power of deep neural networks.
 These models can be used for tasks such as predicting future events, estimating the rate of event occurrence, or identifying correlations between events.
 They offer a flexible and powerful way to analyze point process data, as they can handle complex dependencies between events and incorporate prior knowledge about the process.
+Traditioaly, a point process encodes timestample and type of events, however, it is common in healtcare to have additional sources of information that could be useful for CIF characterization.
+For example the absolute value of patients vital signs might be predictive for next event prediction.[GAP]
 
-Our aims??// In this work, we propose a DNN for joint modeling of NTPP and a DAM.
-Our first aim is to investigate the effectiveness of state encoding for CIF characterization.
-Second, we are interested in seeing the utility of event encoding a downstream task
+our model TEEDAM is a deep learning model for EHR that can simultanously 
+encodes all of events and values and optimizes for a point process objective function as well as a desired outcome.
+We have performed extensive experiments on existing event sequence data to show the effectivness of our improved transformer event encoder against the existing methods.
+In addition, we show that encoding states can reduce point process NLL in two EHRs dataset (p12,p19) and in some cases it can lead to better performance for next event prediction.
+Finally, we provide explanations regarding the interpretablity of out model thanks to the utilized attention mechanisms.
 
-contribs?///
+the contributions of this paper are:
+\begin{itemize}
+    \item cont1
+    \item cont2
+\end{itemize}
+
+explain structure of paper
 
@@
 
-\subsubsection*{Temporal point process}
+\subsection{Temporal point process}
 
-Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where each sample is represented as an event sequence  $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is the total number of events, $t_j$ is event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label).
-The history of events is denoted as $\mathcal{H}_t=\{(t_j,e_j):t_j<t \}$.
+A temporal point process [13, 14] is a stochastic process that models a sequence of events in a continuous time domain.
+Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where $N$ is total number of samples and each sample $\mathcal{S}_{i}$ is represented as a sequence of events $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is the total number of occured events, $t_j$ is the event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label).
+Furthermore, the history of events at time $t$ is denoted as $\mathcal{H}_t=\{(t_j,e_j):t_j<t \}$.
 
@@
 
-Multivariate Hawkes process is the traditional approach to characterize CIFs:
+Here, $*$ denotes conditioning on the history of events.
+Multivariate Hawkes process is the traditional approach to characterize CIFs by assumming a fixed form of intensity to account for the additive influence of an history event:
 
@@
 
-\subsubsection*{Neural temporal point process}
+ where $\mu \ge 0 $, aka base intensity, is an exogenous component that is independent of history, while $\phi(t)>0$, excitation function, is an endogenous component depending on the history that shows the mutual influences.
+The excitation function can ge characterized using exponentials, linear combination of $M$ basis functions \cite*{xuLearningGrangerCausality2016}.
+
+\subsection{Neural temporal point process}
 
@@
 The main idea of a neural temporal point process (NTPP) is to first encode the history of events until $t_j$ using a neural network architecture $h_j=Enc(\mathcal{H}_{j+1};\theta)$.
-Then it tries to estimate $\lambda^*_m(t)=Dec(h_j;\phi)$ for $ t \in (t_j,t_{j+1}]$.
+Then it tries to estimate CIFs using a different decoder architectures $\lambda^*_m(t)=Dec(h_j;\phi)$ for $ t \in (t_j,t_{j+1}]$.
 
-Initial works used recurrent neural networks either in a discrete way [16], a continuous way [39], or a multi-scale multi-channel architecture [24]; then self-attention was also used by [54, 57].
-[57] develop a Structured Transformer Hawkes (STH) model that studies the relationship between point processes.
-It assumes the modeling of one point process might contribute to the modeling of another.
-Each vertex in the STH’s graph is associated with a point process.
-Other efforts have been made to model the cumulative intensity function [42] and conditional probability density [46].
-One issue that has been ignored in these works is that these NPP models take all types of historic events to predict the future ones, neglecting that some event types may not contribute to the prediction of another type.
-This study aims to correct this defect by learning to omit those non-contributing event types when predicting the target type via NPPs.
+Initial works has used recurrent encoders such as RNN, GRU or LSTM [refs] in which the hidden state gets updated after arrival of new event as $h_{j+1}=Update(h_j,(t_j,e_j))$.
+The main advantage is that they allow us to compute history embeddings in $O(L)$ time, however, they are prone to neglect long-term inter-event influences.
+On the other hand, set aggregation encoders directly encodes all past events into a history embedding.
+adopting attention mechanism is one way that can capture long-term influences and can be trained in parallel as well.
+For example, \cite*{zuoTransformerHawkesProcess2020a} proposed transformer hawkes process (THP) that adopts a transformer architecture for event encoding.
 
-\subsubsection*{Parameter Estimation}
+The learned event embeddings can be later used for estimating conditional intensity functions, comulative conditional intensity functions or event probability density functions.
+
+\subsection{Parameter Estimation}
 
@@
 
- Intuitively, the integral term indicates that there is no event of any type in the interval.
+ The paramters of point process can be learnt by Maximum Likeihood Estimation (MLE) framework.
+However, more advanced methods such as adversarial learning, reinforcement learning have also been proposed.
 
@@
 
-\begin{multline} 
-    \log p_{mc}(\mathcal{S}_{i})  =  \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log p_m^{*}(t_j) \\   
-    = \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log \lambda_m^{*}(t_j)\\
-    \sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
+\begin{multline} \label{eq:CIF-mc}
+    \log p_{mc}(\mathcal{S}_{i})  =  \sum_{j = 1}^{L}\sum_{m = 1}^{M} \mathbbm{1}(e_j=m)   \log p_m^{*}(t_j) \\   
+    = \sum_{j = 1}^{L}\sum_{m = 1}^{M} \mathbbm{1}(e_j=m)   \log \lambda_m^{*}(t_j)\\
+    -\sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
 \end{multline}
 
-However, in many cases such as EHRs it is common to have co-occuring events. [ntpp] proposed to add:
+Here, $\mathbbm{1}$ is the indicator function.
+The log-likelihood can be understood using the following two facts.
+First, the quantity $\lambda_k^{*}(t)dt$ corresponds to the probability of observing an event of type $k$ in the infinitesimal interval $[t_j,t_j+dt]$ conditioned on the past events $\mathbb{H}_{t_j}$.
+Second, we can compute the probability of not observing any eevents of type $k$ in the rest of the interval $[0,T]$ as $\exp \left(-\int_{0}^{T}  \lambda_m^{*}(s) ds\right)$.
 
-% If we assume that marks and events are independent,
+However, in many cases such as EHRs it is common to have co-occuring events. \cite*{enguehardNeuralTemporalPoint2020} proposed to use a binary cross entropy function:
 
-% \begin{multline} 
-%     -\log p(\mathcal{S}_{i})  = \\
-%       - \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log p^{*}(e_j=m) \lambda^{*}(t_j) \\   
-%     + \sum_{m = 1}^{M} p^{*}(e_j=m) \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
-% \end{multline}
-
-\begin{multline} 
+\begin{multline} \label{eq:CIF-ml}
     \log p_{ml}(\mathcal{S}_{i}) = \log p_{mc}(\mathcal{S}_{i}) \\
-    + \sum_{m = 1}^{M} (1-1(e_j=m)) \log \left(1-p_m^{*}(t_j)  \right)
+    + \sum_{m = 1}^{M} (1-\mathbbm{1}(e_j=m)) \log \left(1-p_m^{*}(t_j)  \right)
 \end{multline}
@@
 
-\begin{multline} 
+\begin{multline} \label{eq:CIF-marked}
     \log p_{marked}(\mathcal{S}_{i}) =    
@@
 
-\subsubsection*{Deep learning for irregular sampling}
-An irregularly sampled data can be denoted as $\mathcal{D}=\{\mathcal{U}_i\}_{i=1}^{N}$ where $N$ is number of samples and each sample is represented as series of tuple $\mathcal{U}_i={(t_p,k_p,v_p)}$ where $t_p, k_p, v_p$ represents the time, modality and value of $p$-th datapoint respectively.
+This marked case is basically an autoencoder for next mark prediction with a single dimension point process for timestamps only.
+
+\subsection{Deep learning for irregularly sampled data}
+An irregularly sampled data can be denoted as $\mathcal{D}=\{\mathcal{U}_i\}_{i=1}^{N}$ where $N$ is number of samples and each sample is represented as sequence of tuples $\mathcal{U}_i=\{(t_p,k_p,v_p)\}_{p=1}^{P}$ where $P$ is the total number of observations and $t_p, k_p, v_p$ represents the time, modality and value of $p$-th observation respectively.
+
+Recurrent neural networks can naturally deal with sequential data. These models can be made compatible with irregularly sampled data by adding a strategy to consider time.
 
@@
 \label{sec:Related Works}
+
+We focus on Neural point process based on attention mechanism that can potentially addresses the problems of slow serial computing and loss of long-term information.
+In addition, attention weights bring interpretability and can show peer influences between events.
+Self-attentive Hawkes Process (SAHP) \cite*{zhangSelfAttentiveHawkesProcess2020} proposes a multihead attention network as the history encoder.
+In addition, they use a sofplus function that can capture both excitaion and inhibition effects.
+Similarly, Transformer Hawkes Process (THP) \cite*{zuoTransformerHawkesProcess2020a} adopts the transformer architecture \cite*{vaswaniAttentionAllYou2017} with time encodings for event embeddings.
+They have introduced softplus function that can only capture mutual excitations between events.
+In a interesting study, reseraches studied different combination of encoders (self attention and GRU) and decoders and simulated on various datasets for comparison \cite*{enguehardNeuralTemporalPoint2020}.
+They demonsterated that attention-based TPPs appear to transmit pertinent EHR information and perform favourably compared to existing models.
+One gap in the current literature is that they don't have any mean for encoding additional information that can be useful for characterization of CIFs.
+
+Recurrent neural networks have been modified to consider irregulary sampled time series.
+For example, GRU-D \cite*{cheRecurrentNeuralNetworks2018} adapts GRU to consider missingness pattern in the form of feature masks and time intervals to achieve better prediction results.
+RETAIN \cite*{choiRETAINInterpretablePredictive2017} is based on a two-level neural attention model that is specifically designed for EHRs data.
+SeFT \cite*{hornSetFunctionsTime2020} is based on recent advancements in differentiable set function learning, extremely parallelizable with a beneficial memory footprint, thus scaling well to large datasets of long time series and online monitoring scenarios.
+Their use of aggregation function is similar to Transformers that computes the embeddings of set elements independently, leading to lower runtime and memory complexity of O(n).
+Although, these models are nearly end-to-end that eliminates the need for imputation pipeline, it is still unclear how much they are affected by the missigness pattern in EHRs.
+In addition, they do not explicitly model the missingness pattern in the data.
 
@@
 
+The proposed model, TEEDAM, consists of two modules for encoding a dataset with both event sequence data and irregularly sampled time series. The schematic of the model is depicted in \ref*{fig1}.
+
  The key advantage of our proposed model is to combine a transformer-based event encoder (TEE) with a deep attention module (DAM) that can handle an irregularly sampled time series for a any downstream prediciton task.
-In this case the data is represented as $\mathcal{D}=\{\mathcal{S}_i,\mathcal{U}_i\}_{i=1}^{N}$.
-General schematic of TEDAM is depicted in FIG.
+The data is represented as $\mathcal{D}=\{(\mathcal{S}_i,\mathcal{U}_i)\}_{i=1}^{N}$, where $\mathcal{S}_i$ and $\mathcal{U}_i$ are event sequence data and irregular time series for the $i$-th sample.
 
 \subsection{Event Encoder}
-We use a similar transformer architecture [thp] for encoding events with minor modifications for event embedding.
+We use a transformer event encoder similar to THP \cite*{zuoTransformerHawkesProcess2020a} with minor modifications for event embedding.
 
 In the first step, we embed all event marks $E_{emb}=E \times W_{emb}$ where $E_i \in \mathbb{R}^{L \times M}$ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{M \times d_{emb} }  $ is the trainable embedding matrix.
+In addition, we encode vector of timestamps $t=[t_1,t_2,...,t_L] $ to $Z=[z(t_1),z(t_2),...,z(t_L)]\in \mathbb{R}^{L \times d_{time}}$ using following transformation formula:
 
-In the second step, timestamps should be encoded and added to the event embedding, however, we propose to concatenate time encodings that can lead to better characterization of conditional intensity functions.
+\begin{equation}\label{eq:time-encoding}
+    [z(t_j)]_{k} = 
+     \begin{cases} 
+        \cos\left(  \frac{t_j}{\mathcal{T}^{(k-1)/d_t} }  \right) & \text{if } k \text{ is odd}\\
+        \sin\left(  \frac{t_j}{\mathcal{T}^{k/d_t} }  \right) & \text{if } k \text{ is even}
+
+        \end{cases} 
+\end{equation} 
+Here, $\mathcal{T}$ is representing maximum time scale.
+This is very similar to positional encodings in transformers where the index is substitued by the time stamp.
+In contrary to THP and original positional encoding that considers $d_{emb}=d_{time}$ and adds up time encoding to the event embedding, we propose to concatenate this two vectors:
+
+\begin{equation}
+    X_{ev}=[W_{emb}, Z] \in \mathbb{R}^{L \times (d_{emb}+t_{emb})}
+\end{equation}
+
+% In the second step, timestamps should be encoded and added to the event embedding, however, we propose to concatenate time encodings that can lead to better characterization of conditional intensity functions.
 Finally, the input of the transformer encoder will be $X_{emb}=[E_{emb}, T_{emb}] \in \mathbb{R}^{L \times (d_{emb}+t_{emb})}$.
 
-Here, we use the standard transformer encoder similar to [vaswani] with masking matrix to prevent the model from looking into the future.
-we obtain the encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $ where $h_j$ containes the all available information until occurence of $j$-th event.
+Finally, we use the standard transformer encoder to encode embedded events $X_{emb}$ tothe encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $.
+It is important to use approporiate mask matrix to prevent information leakage from future to the past.
+In this case, $h_j$ containes the all available information until occurence of $j$-th event.
 
@@
 
-Similar to [setF], we use an attention-based aggregation approach for encoding all additional information.
-Each side information  $ (t_k, v_k, m_k) $ can be represented by $s_k=(z(t_k), v_k, m_k)$.
-we define attention $a(\mathcal{U}_k,s_k )$
+We use the a deep attention module \cite*{hornSetFunctionsTime2020} for encoding all additional information including irregularly sampled time series.
+Each observation of a sample can be represented by $u_k=(z(t_k), v_k, m_k)$ where $z(t_k)$ is the same time encoding similar to equation \ref{eq:time-encoding}.% we define attention $a(\mathcal{U}_k,s_k )$
 
-We define $\mathcal{U}_p$ to be the set of the first $p$ available information.
-The goal is to calculate $a(\mathcal{U}_p,s_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{U}_p$.
+We define $\mathcal{U}_p$ to be the set of the first $p$ observations.
+The goal is to calculate the attention weight $a(\mathcal{U}_p,u_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{U}_p$.
 This is achieved by computing an embedding of the set elements using a smaller set functions $f^{\prime}$, and projecting the concatenation of the set representation and the individual set element into d-dimensional space:
 
-\begin{equation}
-    f^{\prime}(\mathcal{U}_p) = g^{\prime} \left(  \frac{1}{|p|} \sum_{u_k \in \mathcal{U}_p}  h^{\prime}(u_k;\theta^{\prime}) ;\rho^{\prime}\right) 
+\begin{equation}\label{eq:fprime}
+    y^{\prime}_{p}=f^{\prime}(\mathcal{U}_p) = g^{\prime} \left(  \frac{1}{|p|} \sum_{u_k \in \mathcal{U}_p}  h^{\prime}(u_k;\theta^{\prime}) ;\rho^{\prime}\right) 
 \end{equation}
 
-Then we can compute key values using key matrix $W^k \in \mathbb{R}^{(d_g+d_s) \times d_{prod}}$
+Here, we compute the mean of first $p$ observations after passing first $p$ observations through a multilayer perceptron (MLP) nueral network ($h^{\prime}(u_k;\theta^{\prime})$).
+Finally, a second transformation using $g^{\prime}$ is performed to obtain embeddings $f^{\prime}(\mathcal{U}_p) \in \mathbb{R}^{d_{g^{\prime}}}$.
+
+Then we can compute key values ($ K_p$) using key matrix $W^k \in \mathbb{R}^{(d_{g^{\prime}}+d_s) \times d_{prod}}$:
 
@@
 
-using a query vector $w^q \in \mathbb{R}^{d_{prod}}$
+Then using a query vector $w^q \in \mathbb{R}^{d_{prod}}$, we compute the desired attnetion weight in this way:
+
+\begin{equation}
+    a(\mathcal{U}_p,u_p)=  softmax(\frac{K_p.w^q}{\sqrt{d}  })
+\end{equation}
+
+Finally, we compute a weighted aggregation of set elements using attention weights similar to equation \ref*{eq:fprime}:
 
 \begin{equation*}
-    a(\mathcal{U}_p,u_p)=  softmax(\frac{K_p.w^q}{\sqrt{d}  })
-\end{equation*}
-
-Finally, we compute a weighted aggregation of set elements:
-
-\begin{equation*}
-    z_p = f(\mathcal{U}_p) =     
+    y_p=f(\mathcal{U}_p) =     
     g\left(
-    \sum_{s_k \in \mathcal{U}_p}  a(\mathcal{U}_p,s_k)h(s_k;\theta);\rho 
+    \sum_{u_k \in \mathcal{U}_p}  a(\mathcal{U}_p,u_k)h(u_k;\theta);\rho 
     \right) 
@@
 
-we regard $z_p \in \mathbb{R}^{d_{\rho}}$ as the representation of state data until arrival of $p$-th datapoint.
+We regard $y_p \in \mathbb{R}^{d_g}$ as the representation of first $p$ observations.
 
-Finally, we need to combine event embeddings $ H_{L \times d_e}$ and state embeddings $Z_{P \times d_{\rho}}$, however, the length of each does not match.
+The matrix $Y=[y_1,y_2,...,y_P] \in \mathbb{R}^{P \times d_g}$ keeps the entire representions of the data.
+To use state information for CIFs characterization, we down-sample $Y$ to $Y^{\prime}=[y^{\prime}_1,y^{\prime}_2,...,y^{\prime}_L]  \in \mathbb{R}^{L \times d_g}$ where $ y^{\prime}_j=y_p \text{ where } p=\argmax_{t_p \leq t_j} p$.
+
+Without loss of generality, we can consider multiple heads by adding an additional dimension to keys and queries.
+
+% We need to combine event embeddings $ H_{L \times d_e}$ and state embeddings $Y_{P \times d_g}$, however, the length of each matrices does not match.
 As a result, we consider the reduced version of state matrix as below:
 
-\begin{equation*}
-    Z^{red}_i=Z_p \text{ where } p=argmax(z \leq i)
-\end{equation*}
-
-Without loss of generality, we can consider multiple heads by adding an additional dimension to keys and queries.\\
+% \begin{equation*}
+%     y^{\prime}_j=y_p \text{ where } p=\argmax_{t_p \leq t_j} p
+% \end{equation*}
 
@@
 
-% \begin{equation}\label{eq:1}
-%     [z(t_j)]_{k} = 
-%      \begin{cases} 
-%         \sin\left(  \frac{t_j}{\mathcal{T}^{(k-1)/d_t} }  \right) & \text{if } k \text{ is odd}\\
-%         \sin\left(  \frac{t_j}{\mathcal{T}^{k/d_t} }  \right) & \text{if } k \text{ is even}
-
-%         \end{cases} 
-% \end{equation} 
 % \begin{equation}
@@
 
-\begin{gather*} 
-    \mu_{m,i+1}=gelu(h_{i+1}W_{m,\mu}  +  z_{i+1}W_{m,\mu}), \\ 
-    \eta_{m,i+1}=gelu(h_{i+1}W_{m,\eta}  +  z_{i+1}W_{m,\eta}), \\
-    \gamma_{m,i+1}=gelu(h_{i+1}W_{m,\gamma}  +  z_{i+1}W_{m,\gamma}),
-    \end{gather*}
+\begin{gather} 
+    \mu_{m,j}=gelu(h_{j}W_{m,\mu}  +  y^{\prime}_{j}W_{m,\mu}), \\ 
+    \eta_{m,j}=gelu(h_{j}W_{m,\eta}  +  y^{\prime}_{j}W_{m,\eta}), \\
+    \gamma_{m,j}=gelu(h_{j}W_{m,\gamma}  +  y^{\prime}_{j}W_{m,\gamma}),
+    \end{gather}
 
@@
 
-\begin{multline*}
-    \lambda_m(t)=\text{softplus}(\mu_{m,i+1}+\\
-    (\eta_{m,i+1}- \mu_{m,i+1}) \exp(-\gamma_{m,i+1}(t-t_i))    ),
-\end{multline*}
+\begin{multline}
+    \lambda_m(t)=\text{softplus}(\mu_{m,j}+\\
+    (\eta_{m,j}- \mu_{m,j}) \exp(-\gamma_{m,j}(t-t_j))    ),
+\end{multline}
 
-for $t \in (t_i, t_{i+1}]$,where the $softplus$ is used to constrain thei intensity function to be positive.
+for $t \in (t_j, t_{j+1}]$,where the $softplus$ is used to constrain the intensity function to be positive.
 
@@
 
-We define a multi-objective loss function 
-$ \mathcal{L}   =   \mathcal{L}_{CIF}   +   \mathcal{L}_{mark}  +   \mathcal{L}_{state}$.
+We define a multi-objective loss function as $ \mathcal{L}   =   \mathcal{L}_{CIF}   +   \beta \mathcal{L}_{state}$., where $\mathcal{L}_{CIF}$ could be one of equations \ref*{eq:CIF-mc},\ref*{eq:CIF-ml} or \ref*{eq:CIF-marked}.
+$\mathcal{L}_{state}$ could be any desired loss functioin such as mean square error or corss entropy depending on the task.
+$\beta$ is a coefficient that can be optimized during hyperparamter tuning.
 
@@
 \subsection*{Datasets}
-
-To show the utility of time concatenation and marked loss, we consider three datasets:
 
@@
 
-\subsection*{Scenarios}
+\subsection*{Experimental Setup}
 
-To show the utility of time concatenation and marked-shp we report the metrics for SO, RT and SYN and compare it with 3 baselines: SAHP, THP, and GRU-CP.
+In the first series of experiments, we compare our proposed model for point process modeling in three scenarios:
 
-To show the utility of additional information for CIF modeling, we report NLL/events for P12 and P19 in the following conditions: TE, TE+DAM, TE+noise
+\begin{itemize}
+    \item AE(next mark) is a simple auto-encoder for predicting the next event from event embeddings without CIF characterization.
+As mentioned before, the log-likelihood of the point process will reduce to this one if we ignore the integral term.
+    \item PP(marked) is a marked point process (equation \ref*{eq:CIF-marked}) where we assume marks and time stamps are independent.
+    \item PP(multi-class or multi-label) uses multi-class or multi-label loss.
+\end{itemize}
 
-Finally, we investigate whether point process characterization can be useful in a downstream task or not.
-As each dataset consists of different hospitals and the pattern of irregular sampling might differ from hosptial to hospital, we assume different settings: single-center, multi-center and external evaluation.
+To show the utility of time concatenation, we also report the metrics for the summation case.
 
-\subsection*{Baselines}
+In the second series, we want to investigate the effectiveness of state encoding in addition to the events for real-world EHR datasets P12 and P19 in a multi-label setting.
+We regard the occurrence of certain laboratory variables as events that are based on clinicians' decisions.
+As a result, we consider two scenarios:
 
-we compare our model against existing models: THP, SAHP, GRU-CP.
+\begin{itemize}
+    \item TEE: Here, we only use the transformer event encoder for CIFs characterization.
+    \item TEE+DAM: we further encode time stamps and values of all clinical variables through the deep attention module (DAM).
+\end{itemize}
+
+In the last series of experiments, we investigate the utility of event encoding in a supervised learning task.
+In particular, we try to predict mortality and sepsis shock as a binary outcome in P12 and P19 respectively.
+As a result, we compare (TEE+DAM) against the baseline (DAM) for outcome prediction.
 
 \subsection*{Metrics}
-We report the weighted AUPRC, AUROC of next predicted event as well as root mean sqaure error (RMSE) of next measurement interval.
-For evaluating the goodness of fit for the parameterized point process, we report normalized negative likelihood normalized by number of ocurred event (NLL/events).
-Furthermore, we can also evaluate the learned representation of each patient to predict the sepsis label in a binary classification task.
+We report log-likelihood normalized by the number of events (LL/\#events) as a goodness of fit for CIFs characterization \cite*{zhangSelfAttentiveHawkesProcess2020,zuoTransformerHawkesProcess2020a}.
+For the next event type prediction, we report the weighted measure of F1-score and area under the receiver operating characteristic curve (AUROC) in the multi-class and multi-label setting respectively.
+In the supervised learning task for binary prediction, we report F1-score and area under the precision-recall curve (AUPRC).
 
@@
 To be completed.
+
 \section{Results and Discussion}
@@
 
-\subsection{Effect of minor improvements}
-
-[TB1] shows the performance metrics of different datasets.
-
-In general, time concatenation has lead to better F1-score/AUROC as well as LL/events compared to adding time. This idea can be used for future research in NTPPs for better characterization of CIFs.
-
-Another interesting fact is that in some datasets (A, B) next mark prediction has competetive performance compared to the point process loss.
-However, in many cases, point process loss significantly improves the performance metrics which indicates that modeling non-event likelihood could be beneficial.
-In addition, we propose to use next-mark prediction case as a baseline.
-
-Another interesting results is that independent assumption in the marked loss may lead to worse performance compared to MC/ML loss.
+\subsection{Preliminary comparison}
 
@@
       \midrule
-    \multirow{2}[2]{*}{SO} & LL/\#events & ND    & ND    &       & -0.56 & -0.57 &       &       &       & -1.54 & -1.86 & -1.84 & NR \\
-          & F1-score & 27.86 & 28.79 &       & 29.10 & 28.95 &       &       &       & 28.34(0.19) & 24.12 & 23.89 & 26 \\
+      \multirow{2}[2]{*}{SO (MC)} & LL/\#events & ND    & ND    &       & -0.56 & -0.57 &       & -2.04 & -2.05 & -1.54 & -1.86 & -1.84 & NR \\
+            & F1-score & 38.46 & 36.91 &       & 36.04 & 34.91 &       & \multicolumn{1}{r}{32.65} & 31.67 & 28.34(0.19) & 24.12 & 23.89 & 26 \\
       \midrule
       \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & ND    & ND    &       &       &       &       &       &       & -3.89 & -4.56 & -4.57 & NR \\
-          & F1-score &       &       &       &       &       &       &       &       & 58.29 & 53.92 & 53.86 & NR \\
+            & F1-score & 62.48 & 61.65 &       & 58.45 & 35.59 &       & 36.38 &       & 58.29 & 53.92 & 53.86 & NR \\
       \midrule
-    \multirow{2}[2]{*}{Synthea} & LL/\#events & ND    & ND    &       &       &       &       &       &       & ND    & ND    & ND    & NR \\
-          & AUROC &       &       &       &       &       &       &       &       & ND    & ND    & ND    & 0.85(.014) \\
+      \multirow{2}[2]{*}{Synthea (ML)} & LL/\#events & ND    & ND    &       &       &       &       &       &       & ND    & ND    & ND    & NR \\
+            & AUROC & 89.58 & 89.19 &       & 64.98 & 63.95 &       & 60.95 & 60.65 & ND    & ND    & ND    & 0.85(.014) \\
       \midrule
-    \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & ND    & ND    &       & 1.86  & 1.4   &       &       &       & ND    & ND    & ND    & NR \\
-          & AUROC & 63.3  & 61.01 &       & 68.7  & 66.2  &       & 73.87 &       & ND    & ND    & ND    & 0.611(0.001) \\
+      \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & ND    & ND    &       & 1.589 & 1.355 &       & -1.587 & -1.656 & ND    & ND    & ND    & NR \\
+            & AUROC & 69.12 & 67.86 &       & 62.5  & 60.4  &       & 73.9  & 71.8  & ND    & ND    & ND    & 0.611(0.001) \\
       \bottomrule
       \end{tabular}%
-  \label{tab:addlabel}%
+    \label{tab:1}%
   \end{table*}%
 
-\subsection{Negative Likelihood with state encoding}
-
-[TB2] shows the result for estimation of negative likelihood for the two datasets.
-
-In general we can see that using state information results in lower NLL, however, it cannot neccessarily increase the AUROC for next time prediction.
-
-Another observation is that more data will lead to higher LL.
-
 % Table generated by Excel2LaTeX from sheet 'state enc'
-\begin{table}[htbp]
+\begin{table*}[htbp]
     \centering
     \caption{Add caption}
-      \begin{tabular}{ccccc}
+      \begin{tabular}{ccccccccc}
       \toprule
-            &       & \multicolumn{3}{c}{Model} \\
-  \cmidrule{3-5}    Dataset & setting & TE    & TE+DAM & TE+noise \\
+      \toprule
+            &       & \multicolumn{3}{c}{LL/\#Events} &       & \multicolumn{3}{c}{AUROC} \\
+  \cmidrule{3-5}\cmidrule{7-9}    Dataset & setting & TEE   & TEE+DAM & TEE+noise &       & TEE   & TEE+DAM & TEE+noise \\
       \midrule
-      \multirow{3}[2]{*}{P12} & sc    & -1.99 & -1.77 & -1.92 \\
-            & mc1   & -0.99 & -1.14 & -0.89 \\
-            & mc2   & -1.5  & -1.33 & -1.47 \\
+      \multirow{3}[2]{*}{P12} & sc    & -0.2345 & \textbf{-0.0019} & 0.2365 &       & 67.76 & \textbf{71.66} & 69.99 \\
+            & mc1   & 0.1984 & \textbf{0.3623} & 0.2252 &       & 74.88 & \textbf{81.39} & 74.52 \\
+            & mc2   & 0.1634 & \textbf{0.2337} & 0.0858 &       & 78.67 & \textbf{80.24} & 72.03 \\
       \midrule
-      \multirow{3}[2]{*}{P19} & sc    & 0.55 (0.02) & 0.55 (0.02) & 0.55 (0.02) \\
-            & mc1   & -1.81 & -1.62 & -1.69 \\
-            & mc2   & -2.77 & -2.38 & -2.67 \\
+      \multirow{3}[2]{*}{P19} & sc    & -0.9734 & \textbf{-0.8531} & -1.047 &       & 79.3  & \textbf{82.52} & 72.13 \\
+            & mc1   & -0.9182 & \textbf{-0.7641} & -0.937 &       & 77.58 & \textbf{87.09} & 73.11 \\
+            & mc2   & -1.199 & \textbf{-1.04} & -1.135 &       & 64    & \textbf{73.31} & 64.96 \\
+      \bottomrule
       \bottomrule
       \end{tabular}%
-    \label{tab:addlabel}%
-  \end{table}%
+    \label{tab:2}%
+  \end{table*}%
 
-can u provide one example patient?
+Table \ref{tab:1} shows the performance metrics of TEEDAM for different datasets in various scenarios.
+We can see that in most cases time concatenation leads to better results in terms of LL/\#events and F1-score/AUROC while adding time encodings to the event embeddings is the default practice in the literature \cite*{zhangSelfAttentiveHawkesProcess2020,zuoTransformerHawkesProcess2020a}.
 
-\subsection{Downstream task}
+Another interesting fact is that the simple auto-encoder for next event type prediction achieves better results for SO,RT(MC) and SYN(ML) compared to the point process scenario.
+These datasets have been widely used in the point process literature to show the effectiveness of point process, however, we show that this simple baseline may perform better for some datasets.
 
-[TB3] indicates the result for Mortality/sepsis prediction in different settings and hospital centers.
+RT(MC) is the only dataset in which the point process with multi-label loss (equation \ref*{eq:CIF-ml}) performs better indicating the real advantage of modeling non-event likelihood of point process in this dataset.
+In general, we would recommend to compare future works in this area with this simple baseline.
 
-In some cases, incorporating event embeddings can lead to better performance.
+\subsection{State encoding for CIF characterization}
+
+Table \ref*{tab:2} shows the result for the utility of state encodings for the estimation of negative likelihood.
+We see that in all cases, state encoding leads to higher AUROC for the next event type prediction and LL/\#event as well.
+It is intuitive that in a hospital, the absolute value of patient states can be useful for ordering future laboratory events.
+We have evaluated our model in a healthcare database, however, it can be further evaluated on other event sequence data where additional information is available.
+
+\subsection{Event encoding in supervised learning}
+
+Table \ref*{tab:3} indicates the result for Mortality/sepsis prediction task across different settings and hospital centers.
+Generally, we can see that in the mc1 setting, has the best performance because of the larger training dataset from other centers and the test center.
+In addition, event encoding module is useful for 2/5 cases while in 1/5 cases degrades the performance.
+
+In general, it might seem problematic to rely on the missingness pattern for outcome prediction as it can hurt generalizability when transferring to a new environment with a different pattern.
+We regard this approach as a double-edged sword that could improve performance in some cases, especially in a new environment with a similar pattern, but it can also degrade the performance if the target environment has a completely different pattern.
 
@@
       \end{tabular}%
-    \label{tab:addlabel}%
+    \label{tab:3}%
   \end{table*}%
@@
 \subsection{Learned representions}
-Fig 1 visualizes the tsne plot for the two scenarios.
+
+\begin{figure*}[!t]
+    \centerline{\includesvg[inkscapelatex=false]{images/tsne.svg}}
+    \caption{Magnetization as a function of applied field.
+    It is good practice to explain the significance of the figure in the caption.}
+    \label{fig1}
+    \end{figure*}
+
+Fig 1 visualizes the t-SNE plot for the learned embeddings of [] with and without event encodings.
+Besides, we visualize the sampling pattern of an example patient and its neighbors.
+As can be seen, in TEEDAM neighbors have more similar pattern compared to the baseline.
+Although event encoding does not necessarily leads to better performance, we can learn a better representation that considers both absolute values and the patterns.
+One application could be time series generation.
 
@@
 
-\subsection{Likelihood estimation}
-
-Although CIF does not improve mark prediction, it has led to better representation of patient for downstream task such as sepsis prediction.
-
-In addition, we can interpret some of learned CIF patterns.
-
-explain the effect of time concatenation in SO dataset
-
-tsne of learned representation. 4 modes:
-
-\begin{itemize}
-    \item (DA,TE)->(Mark, CIF)
-\end{itemize}
-
-attention of DA for sepsis prediction
-
-attention matrix of events for SO dataset
-
 \section{Conclusion}
 \label{sec:Conclusion}
+
+data generation with similar patterns
+
+pattern similarity index
 

### 2023-03-03 Manuscript_v1.tex
update
update ol

@@
 
+% https://github.com/James-Yu/LaTeX-Workshop/wiki/FAQ#how-to-pass--shell-escape-to-latexmk
 \usepackage{svg}
@@
 
-    Irregular sampling of time series in electronic health records (EHRs) is challenging for model development.
-In addition, the pattern of missingness for certain clinical variables are not at random as it is determined by clinicians decision and the state of patient.
-Point process is a mathematical framework for handling event sequence data which is also consistent with the irregular sampling.
-We propose \emph{TEEDAM}, a deep neural network that can learn patient representation from irregulary sample time series as well as informative missingness pattern of certain laboratory variables.
-We performed various experiments to show the effectiveness of event and state encoding for characterization of conditional intensity functions as well as downstream prediction task.
-Results show that in some cases learning from patterns may improve the performance of prediction task.
+    Irregular sampling of time series in electronic health records (EHRs) presents a challenge for model development.
+In addition, the pattern of missingness for certain clinical variables is not random, but rather determined by clinicians' decisions and the patient's state.
+Point process is a mathematical framework for handling event sequence data that is consistent with irregular sampling. 
+We propose \emph{NPP4EHR}, a deep neural network that can learn patient representations from irregularly sampled time series using the neural point process framework.
+We conducted various experiments to demonstrate the effectiveness of event and state encoding for characterizing conditional intensity functions, as well as for downstream prediction tasks, using two healthcare databases.
+Our proposed transformer event encoder outperforms state-of-the-art models in commonly used event sequence databases, as evidenced by its superior performance in negative log-likelihood and next-event prediction.
+Furthermore, we demonstrated the usefulness of state encoding for improved conditional intensity function characterization, as well as event encoding for better representation learning in downstream tasks, using two real-world healthcare datasets.
+We propose that our model could be leveraged for better representation learning in EHRs.
+
 \end{abstract}
@@
 
-[ML in healthcare]Machine learning has the potential to revolutionize healthcare by leveraging vast amounts of data available in electronic health records (EHRs) to develop more accurate clinical decision support systems.
+[ML in healthcare] Machine learning has the potential to revolutionize healthcare by leveraging vast amounts of data available in electronic health records (EHRs) to develop more accurate clinical decision support systems.
 EHRs store patient health information, such as medical history, medications, lab results, and diagnostic images, which can be used as input for machine learning algorithms to identify patterns and associations that could inform more precise diagnoses, better treatment plans, and earlier interventions.
@@
 
-[irregular sampling challenge] One of data challenges for machine learning (ML) when using electronic health records (EHRs) is irregular sampling.
+[irregular sampling challenge] One of the data challenges for machine learning (ML) when using electronic health records (EHRs) is irregular sampling.
 EHR data is often collected at different times and frequencies, depending on a patient's healthcare needs and visit schedules, which can result in uneven and irregularly sampled time series data.
+Asynchronous and incomplete observation of certain clinical variables can also be regarded as missingness in the data.
 
@@
 
-[sources of missingness] Sources of missingness in EHRs must be carefully understood.
-For example, lab measurements are typically ordered as part of a diagnostic work-up, meaning that the presence of a datapoint conveys information about the patient's state
-% \cite*{ghassemiReviewChallengesOpportunities2020}.
+[sources of missingness] The sources of missing data in EHRs must be carefully understood.
+For instance, lab measurements are usually ordered as part of a diagnostic work-up, and the presence or absence of a data point conveys information about the patient's state \cite*{ghassemiReviewChallengesOpportunities2020}.
 
-[conseqs of IRR] A typical EHR consist of tabular and time series data.
-Depending on the desired outcome, we can formulate the problem in static or longitudinal setting.
-There are various consequences asscociated with irregularly sampled data.
-Many machine learning models are not inherently compatible with
-
-imputation for static setting
-
-imputation for time series.
-
-inherently compatible methods.
-
-imputation methods/// Imputation techniques are widely used in electronic health records (EHRs) to handle missing data.
+[imputation methods] Imputation techniques are widely used in electronic health records (EHRs) to handle missing data.
 Imputation refers to the process of filling in the missing data with plausible values based on the available information.
-There are several imputation techniques available, each with their own strengths and weaknesses.
+There are several imputation techniques available, each with its strengths and weaknesses.
 For example, mean imputation replaces missing values with the mean of the available values in the same column, while regression imputation uses a regression model to estimate the missing values based on the relationship between the variables.
-Other popular imputation methods include multiple imputation, hot deck imputation, and k-nearest neighbors imputation.
+Other popular imputation methods include multiple imputations, hot deck imputation, and k-nearest neighbors imputation.
 The choice of imputation technique depends on the type and amount of missing data, as well as the goals of the analysis.
@@
 
-There exist several ML frameworks that are inherently compatible with the missig or irregularly-sampled data such as Gaussian process and recurrent neural networks.
-Deep learning models that are compatible with irregularly sampled time series are important in many real-world applications, such as stock market predictions, speech recognition, and medical diagnosis.
+There exist several ML frameworks that are inherently compatible with the missing or irregularly-sampled data such as Gaussian process and recurrent neural networks.
+Deep learning models that can handle irregularly sampled time series are crucial in many real-world applications, including stock market predictions, speech recognition, and medical diagnosis.
+These models must be able to process data that is not evenly spaced, which is common in time-sensitive applications.
+One popular deep learning architecture that can handle irregularly sampled time series is the Recurrent Neural Network (RNN).
+RNNs are well-suited for this task because they can process sequential data, taking into account not only the current input but also the previous inputs.
+Another architecture that can handle irregularly sampled time series is the Convolutional Neural Network (CNN), which can extract features from the data and then pass the processed data to an RNN for further analysis.
+Additionally, Attention Mechanisms can be integrated into RNNs and CNNs to help the model focus on important features in the data.
+These deep learning models offer a powerful toolset for handling irregularly sampled time series and providing useful insights into this type of data.
+However, it is important to note that these models do not explicitly account for the fact that the absence of some data points can convey unique information, and it is important to carefully consider the sources of missingness in the data.
+
+% Deep learning models that are compatible with irregularly sampled time series are important in many real-world applications, such as stock market predictions, speech recognition, and medical diagnosis.
 These models need to be able to handle data that is not evenly spaced, as is often the case in time-sensitive applications.
 One popular deep learning architecture that can handle irregularly sampled time series is the Recurrent Neural Network (RNN).
-RNNs are well-suited for this task because they have the ability to process sequential data, taking into account not only the current input but also the previous inputs.
+RNNs are well-suited for this task because they can process sequential data, taking into account not only the current input but also the previous inputs.
 Another architecture that can handle irregularly sampled time series is the Convolutional Neural Network (CNN), which can be used to extract features from the data and then pass the processed data to an RNN for further analysis.
@@
 Overall, these deep learning models offer a powerful toolset for handling irregularly sampled time series and provide useful insights into this type of data.
-These models does not explicitly take into account the fact wether the absence of some datapoints could convey some unique information.
+However, these models do not explicitly take into account the fact wether the absence of some datapoints could convey some unique information.
 
-Point processes are mathematical models used to describe the distribution of events in time or space.
+Point process is the mathematical framework for describing the distribution of events in time or space.
 These events can be occurrences of earthquakes, nerve impulses, customer purchases, or anything else that can be counted or measured.
 In healthcare, lab measurements can be regarded as sequence of events that are ordered by clinicians.
-At the core of point process is the defintion of conditional intensity function and the corresponding log likelihood which simultanously models the occurence of events using the history of past events.[GAP]
-More recently, neural Point Processes have been developed for better characteriztion of CIFs by leveraing the power of deep neural networks.
+At the core of point process is the defintion of conditional intensity functions and the corresponding log-likelihood which simultanously models the occurence of events using the history of past events.
+
+Point process is a mathematical framework for describing the distribution of events in time or space.
+These events can be occurrences of earthquakes, nerve impulses, customer purchases, or anything else that can be counted or measured.
+In healthcare, lab measurements can be regarded as a sequence of events that are ordered by clinicians.
+At the core of the point process is the definition of conditional intensity functions and the corresponding log-likelihood, which simultaneously models the occurrence of events using the history of past events.
+The conditional intensity function represents the instantaneous rate of event occurrence, given the history of past events.
+This function can be used to predict the probability of future events, which is particularly useful in healthcare applications such as predicting disease onset or detecting adverse events.
+
+% More recently, neural Point Processes (NPP) have been developed for better characteriztion of CIFs by leveraing the power of deep neural networks.
 These models can be used for tasks such as predicting future events, estimating the rate of event occurrence, or identifying correlations between events.
 They offer a flexible and powerful way to analyze point process data, as they can handle complex dependencies between events and incorporate prior knowledge about the process.
-Traditioaly, a point process encodes timestample and type of events, however, it is common in healtcare to have additional sources of information that could be useful for CIF characterization.
-For example the absolute value of patients vital signs might be predictive for next event prediction.[GAP]
+Traditioaly, a point process encodes timestamp and type of events, however, it is common in healtcare to have additional sources of information that could be useful for CIF characterization.
+For example the absolute value of patients vital signs might be predictive for next event prediction.
 
-our model TEEDAM is a deep learning model for EHR that can simultanously 
-encodes all of events and values and optimizes for a point process objective function as well as a desired outcome.
-We have performed extensive experiments on existing event sequence data to show the effectivness of our improved transformer event encoder against the existing methods.
-In addition, we show that encoding states can reduce point process NLL in two EHRs dataset (p12,p19) and in some cases it can lead to better performance for next event prediction.
-Finally, we provide explanations regarding the interpretablity of out model thanks to the utilized attention mechanisms.
+More recently, Neural Point Processes (NPP) have been developed to better characterize Conditional Intensity Functions (CIFs) by leveraging the power of deep neural networks.
+These models can be used for tasks such as predicting future events, estimating the rate of event occurrence, or identifying correlations between events.
+They offer a flexible and powerful way to analyze point process data, as they can handle complex dependencies between events and incorporate prior knowledge about the process.
+Traditionally, a point process encodes the timestamp and type of events.
+However, in healthcare, additional sources of information can be available that could be useful for CIF characterization.
+For example, the absolute value of a patient's vital signs might be predictive for next-event prediction.
+NPP models can incorporate such information by using event and state encoders that can learn to extract relevant features from the input data.
 
-the contributions of this paper are:
+Our proposed model, NPP4EHR, is a deep learning model for electronic health records (EHRs) that is based on point process theory.
+It consists of two modules: a transformer event encoder, which is designed to handle the missingness patterns of certain laboratory values, and a deep attention module for encoding all other available data.
+Both modules can be jointly learned by using appropriate loss functions for conditional intensity function (CIF) characterization and any downstream task.
+
+We have conducted extensive experiments on existing event sequence data to demonstrate the effectiveness of our improved transformer event encoder in comparison to existing methods.
+Furthermore, we have shown that encoding states can reduce the negative log-likelihood (NLL) in two EHR datasets and, in some cases, lead to better performance for next event prediction.
+Overall, our proposed model offers a powerful tool for handling the irregularly sampled time series data present in EHRs and has the potential to improve representation learning in this domain.
+The contributions of this paper are:
 \begin{itemize}
-    \item cont1
-    \item cont2
+    \item An improved transformer event encoder that can effectively handle missingness patterns in certain laboratory values and has been validated in baseline datasets.
+    \item A deep attention module for encoding all available patient data, which has been shown to improve CIF characterization for laboratory values.
+    \item Leveraging transformer event encoder in downstream tasks for better representation learning of patient's EHR
 \end{itemize}
-
-explain structure of paper
 
@@
 
-A temporal point process [13, 14] is a stochastic process that models a sequence of events in a continuous time domain.
-Consider an event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where $N$ is total number of samples and each sample $\mathcal{S}_{i}$ is represented as a sequence of events $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is the total number of occured events, $t_j$ is the event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary represntation of event marks (multi-class or multi-label).
+A temporal point process is a stochastic process that models a sequence of events in a continuous time domain.
+Consider event sequence data  $\mathcal{D}=\{\mathcal{S}_{i}\}_{i=1}^N$ where $N$ is the total number of samples and each sample $\mathcal{S}_{i}$ is represented as a sequence of events $\mathcal{S}_{i}=\{(t_i,e_i)\}_{j=1}^L$, where $L$ is the total number of occurred events, $t_j$ is the event's timestamp, and $e_j \in \mathbb{R}^M $ is the binary representation of event marks (multi-class or multi-label).
 Furthermore, the history of events at time $t$ is denoted as $\mathcal{H}_t=\{(t_j,e_j):t_j<t \}$.
 
-The core idea of the point process framework is the definition of conditional intensity functions (CIFs) which is the probability of the occurence of an event of type $m$ in an infinitesimal time window $[t,t+dt)$:
+The core idea of the point process framework is the definition of conditional intensity functions (CIFs) which is the probability of the occurrence of an event of type $m$ in an infinitesimal time window $[t,t+dt)$:
 
-\begin{equation} 
+\begin{equation} \label{eq:lambda_def}
    \lambda_m^{*}(t)=\lim \frac{P(  \text{event of type m in } [t,t+\Delta t) |   \mathcal{H}_t  )}{\Delta t} 
@@
 
-Here, $*$ denotes conditioning on the history of events.
-Multivariate Hawkes process is the traditional approach to characterize CIFs by assumming a fixed form of intensity to account for the additive influence of an history event:
+Here, $*$ denotes conditioning on the history of events$\mathcal{H}_t$.
+The Multivariate Hawkes process is the traditional approach to characterize CIFs by assuming a fixed form of intensity to account for the additive influence of an event in the past:
 
@@
  where $\mu \ge 0 $, aka base intensity, is an exogenous component that is independent of history, while $\phi(t)>0$, excitation function, is an endogenous component depending on the history that shows the mutual influences.
-The excitation function can ge characterized using exponentials, linear combination of $M$ basis functions \cite*{xuLearningGrangerCausality2016}.
-
-\subsection{Neural temporal point process}
-
-Encoder-decoder architectures have proven to be effective in many applications.
-The main idea of a neural temporal point process (NTPP) is to first encode the history of events until $t_j$ using a neural network architecture $h_j=Enc(\mathcal{H}_{j+1};\theta)$.
-Then it tries to estimate CIFs using a different decoder architectures $\lambda^*_m(t)=Dec(h_j;\phi)$ for $ t \in (t_j,t_{j+1}]$.
-
-Initial works has used recurrent encoders such as RNN, GRU or LSTM [refs] in which the hidden state gets updated after arrival of new event as $h_{j+1}=Update(h_j,(t_j,e_j))$.
-The main advantage is that they allow us to compute history embeddings in $O(L)$ time, however, they are prone to neglect long-term inter-event influences.
-On the other hand, set aggregation encoders directly encodes all past events into a history embedding.
-adopting attention mechanism is one way that can capture long-term influences and can be trained in parallel as well.
-For example, \cite*{zuoTransformerHawkesProcess2020a} proposed transformer hawkes process (THP) that adopts a transformer architecture for event encoding.
-
-The learned event embeddings can be later used for estimating conditional intensity functions, comulative conditional intensity functions or event probability density functions.
+The excitation function can be characterized using exponentials, or linear combination of $M$ basis functions \cite*{xuLearningGrangerCausality2016}.
 
@@
 
-Based on conditional intensity function, it is straightforward to derive conditional probability density function $p^{*}_{m}(t)$ in $(t_j, t_{j+1}]$:
+ Based on the conditional intensity function \ref*{eq:lambda_def}, it is straightforward to derive conditional probability density function $p^{*}_{m}(t)$ in the interval $(t_j, t_{j+1}]$:
 
  \begin{equation} 
-    p_m^{*}(t)=\lambda_m^{*}(t) \exp \left[-\sum_{m=1}^{M} \int_{t_j}^{t_{j+1}}\lambda_m^{*}(t^{\prime})dt^{\prime}\right]
+     p_m^{*}(t)=\lambda_m^{*}(t) \exp \left[-\sum_{m=1}^{M} \int_{t_j}^{t_{j+1}}\lambda_m^{*}(s)ds\right]
   \end{equation}
 
- The paramters of point process can be learnt by Maximum Likeihood Estimation (MLE) framework.
-However, more advanced methods such as adversarial learning, reinforcement learning have also been proposed.
+  The parameters of the point process can be learned by Maximum Likelihood Estimation (MLE) framework.
+However, more advanced methods such as adversarial learning, and reinforcement learning have also been proposed.
 
-In the multi-class setting, the log-likelihood (LL) of a point process for a single event sequnce $\mathcal{S}_{i}$ is defined as:
+ In the multi-class setting, the log-likelihood (LL) of a point process for a single event sequence $\mathcal{S}_{i}$ is defined as:
 
@@
      = \sum_{j = 1}^{L}\sum_{m = 1}^{M} \mathbbm{1}(e_j=m)   \log \lambda_m^{*}(t_j)\\
-    -\sum_{m = 1}^{M} \left(   \int_{0}^{T}  \lambda_m^{*}(s) \,ds  \right)
+     -\sum_{m = 1}^{M} \left(   \int_{t_1}^{t_L}  \lambda_m^{*}(s) \,ds  \right)
  \end{multline}
@@
 First, the quantity $\lambda_k^{*}(t)dt$ corresponds to the probability of observing an event of type $k$ in the infinitesimal interval $[t_j,t_j+dt]$ conditioned on the past events $\mathbb{H}_{t_j}$.
-Second, we can compute the probability of not observing any eevents of type $k$ in the rest of the interval $[0,T]$ as $\exp \left(-\int_{0}^{T}  \lambda_m^{*}(s) ds\right)$.
+Second, we can compute the probability of not observing any events of type $k$ in the rest of the interval $[t_1,t_L]$ as $\exp \left(-\int_{t_1}^{t_L}  \lambda_m^{*}(s) ds\right)$.
 
-However, in many cases such as EHRs it is common to have co-occuring events. \cite*{enguehardNeuralTemporalPoint2020} proposed to use a binary cross entropy function:
+ However, in many cases, such as in EHRs, it is common to have co-occurring events. To handle this issue, \cite*{enguehardNeuralTemporalPoint2020} proposed using a binary cross-entropy function:
 
@@
 
-It should be noted that the real advantage of point process is modeling non-event likelihoods in the form of integrals.
+%  It should be noted that the real advantage of point process is modeling non-event likelihoods in the form of integrals.
 If we neglect the integrals, we would achieve the cross-entropy and binary cross entropy loss in the multi-class and multi-label settings respectively for the prediction of next mark given history of events.
 
-Another approach is the marked case which assumes that marks and timestamps are conditionally independent given $\mathcal{H}_t$.
+ It is important to note that the main advantage of point processes lies in their ability to model non-event likelihoods in the form of integrals.
+If we neglect the integrals, we would end up with the cross-entropy and binary cross-entropy loss in the multi-class and multi-label settings, respectively, for predicting the next mark given the history of events.
+
+ Another approach is the marked case, which assumes that the marks and timestamps are conditionally independent given $\mathcal{H}_t$:
 
@@
      \log p_{marked}(\mathcal{S}_{i}) =    
-    \sum_{j = 1}^{L}\sum_{m = 1}^{M} 1(e_j=m)   \log p^{*}(e_j=m) \\
-    +\sum_{j = 1}^{L}\lambda^{*}(t_j) - \int_{t_0}^{t_{L}}\lambda^{*}(t^{\prime})dt^{\prime}  
+     \sum_{j = 1}^{L}\sum_{m = 1}^{M} \mathbbm{1}(e_j=m)   \log p^{*}(e_j=m) \\
+     +\sum_{j = 1}^{L}\lambda^{*}(t_j) - \int_{t_1}^{t_{L}}\lambda^{*}(t^{\prime})dt^{\prime}  
  \end{multline}
@@
  This marked case is basically an autoencoder for next mark prediction with a single dimension point process for timestamps only.
+
+\subsection{Neural temporal point process}
+
+Encoder-decoder architectures have proven to be effective in many applications.
+The main idea of a neural temporal point process (NTPP) is to first encode the history of events until $t_j$ using a neural network architecture $h_j=Enc(\mathcal{H}_{j+1};\theta)$.
+Then it tries to estimate $m$ CIF using a different decoder architecture $\lambda^*_m(t)=Dec(h_j;\phi)$ for $ t \in (t_j,t_{j+1}]$.
+
+% Initial works have used recurrent encoders such as RNN, GRU or LSTM [refs] in which the hidden state gets updated after the arrival of a new event as $h_{j+1}=Update(h_j,(t_j,e_j))$.
+The main advantage is that they allow us to compute history embeddings in $O(L)$ time, however, they are prone to neglect long-term inter-event influences.
+On the other hand, set aggregation encoders directly encode all past events into a history embedding.
+Adopting an attention mechanism is one such way that can capture long-term influences and can be trained in parallel as well which makes it more computationally efficient.
+For example, \cite*{zuoTransformerHawkesProcess2020a} proposed Transformer Hawkes Process (THP) that adopts a transformer architecture for event encoding.
+
+Initial works have utilized recurrent encoders such as RNN, GRU, or LSTM, where the hidden state is updated after the arrival of a new event as $h_{j+1}=Update(h_j,(t_j,e_j))$.
+The main advantage of these models is their ability to compute history embeddings in $O(L)$ time.
+However, they are susceptible to ignoring long-term inter-event dependencies.
+In contrast, set aggregation encoders encode all past events directly into a history embedding.
+One approach to capture long-term dependencies is to use an attention mechanism, which can be trained in parallel and is more computationally efficient.
+For instance,\cite{zuoTransformerHawkesProcess2020a} proposed Transformer Hawkes Process (THP), which adopts a transformer architecture for event encoding.
+
+The learned event embeddings can be later used for estimating conditional intensity functions, cumulative conditional intensity functions or probability density functions.
 
@@
 
-We focus on Neural point process based on attention mechanism that can potentially addresses the problems of slow serial computing and loss of long-term information.
+We focus on Neural point process models based on attention mechanisms that can potentially address the problems of slow serial computing and loss of long-term information.
 In addition, attention weights bring interpretability and can show peer influences between events.
-Self-attentive Hawkes Process (SAHP) \cite*{zhangSelfAttentiveHawkesProcess2020} proposes a multihead attention network as the history encoder.
-In addition, they use a sofplus function that can capture both excitaion and inhibition effects.
+Self-attentive Hawkes Process (SAHP) \cite*{zhangSelfAttentiveHawkesProcess2020} proposes a multi-head attention network as the history encoder.
+In addition, they use a sofplus function that can capture both excitation and inhibition effects.
 Similarly, Transformer Hawkes Process (THP) \cite*{zuoTransformerHawkesProcess2020a} adopts the transformer architecture \cite*{vaswaniAttentionAllYou2017} with time encodings for event embeddings.
 They have introduced softplus function that can only capture mutual excitations between events.
-In a interesting study, reseraches studied different combination of encoders (self attention and GRU) and decoders and simulated on various datasets for comparison \cite*{enguehardNeuralTemporalPoint2020}.
-They demonsterated that attention-based TPPs appear to transmit pertinent EHR information and perform favourably compared to existing models.
-One gap in the current literature is that they don't have any mean for encoding additional information that can be useful for characterization of CIFs.
+In an interesting study \cite*{enguehardNeuralTemporalPoint2020}, researchers studied different combinations of encoders (self-attention and GRU) and decoders and simulated on various datasets for comparison.
+They demonstrated that attention-based TPPs appear to transmit pertinent EHR information and perform favorably compared to existing models.
+One gap in the current literature is that they don't have any means for encoding additional information that can be useful for the characterization of CIFs.
+For example, in EHRs there are many information in addition to clinical events that could be useful
 
-Recurrent neural networks have been modified to consider irregulary sampled time series.
-For example, GRU-D \cite*{cheRecurrentNeuralNetworks2018} adapts GRU to consider missingness pattern in the form of feature masks and time intervals to achieve better prediction results.
+Recurrent neural networks have been modified to consider irregularly sampled time series.
+For example, GRU-D \cite*{cheRecurrentNeuralNetworks2018} adapts GRU to consider missingness patterns in the form of feature masks and time intervals to achieve better prediction results.
 RETAIN \cite*{choiRETAINInterpretablePredictive2017} is based on a two-level neural attention model that is specifically designed for EHRs data.
 SeFT \cite*{hornSetFunctionsTime2020} is based on recent advancements in differentiable set function learning, extremely parallelizable with a beneficial memory footprint, thus scaling well to large datasets of long time series and online monitoring scenarios.
-Their use of aggregation function is similar to Transformers that computes the embeddings of set elements independently, leading to lower runtime and memory complexity of O(n).
-Although, these models are nearly end-to-end that eliminates the need for imputation pipeline, it is still unclear how much they are affected by the missigness pattern in EHRs.
+Their use of aggregation function is similar to transformers that compute the embeddings of set elements independently, leading to lower runtime and memory complexity of $O(n)$.
+Although these models are nearly end-to-end that eliminate the need for an imputation pipeline, it is still unclear how much they are affected by the missingness pattern in EHRs.
 In addition, they do not explicitly model the missingness pattern in the data.
@@
 
-The proposed model, TEEDAM, consists of two modules for encoding a dataset with both event sequence data and irregularly sampled time series. The schematic of the model is depicted in \ref*{fig1}.
+The proposed model, NPP4EHR, consists of two modules for encoding a dataset with both event sequence data and irregularly sampled time series. The schematic of the model is depicted in \ref*{fig1}.
 
- The key advantage of our proposed model is to combine a transformer-based event encoder (TEE) with a deep attention module (DAM) that can handle an irregularly sampled time series for a any downstream prediciton task.
-The data is represented as $\mathcal{D}=\{(\mathcal{S}_i,\mathcal{U}_i)\}_{i=1}^{N}$, where $\mathcal{S}_i$ and $\mathcal{U}_i$ are event sequence data and irregular time series for the $i$-th sample.
+ The key advantage of our proposed model is to combine a transformer-based event encoder (TEE) with a deep attention module (DAM) that can handle an irregularly sampled time series for any downstream prediction task.
+The data is represented as $\mathcal{D}=\{(\mathcal{S}_i,\mathcal{U}_i)\}_{i=1}^{N}$, where $\mathcal{S}_i$ and $\mathcal{U}_i$ are event sequence data and irregular time series for the $i$-th sample as described in previous section.
 
 \subsection{Event Encoder}
-We use a transformer event encoder similar to THP \cite*{zuoTransformerHawkesProcess2020a} with minor modifications for event embedding.
-
-In the first step, we embed all event marks $E_{emb}=E \times W_{emb}$ where $E_i \in \mathbb{R}^{L \times M}$ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{M \times d_{emb} }  $ is the trainable embedding matrix.
+We use a transformer event encoder similar to THP \cite*{zuoTransformerHawkesProcess2020a} with minor modifications for time encoding.
+In the first step, we embed all event marks as $E_{emb}=E \times W_{emb}$, where $E_i \in \mathbb{R}^{L \times M}$ is the binary encoding matrix of all event marks (multi-label or multi-class), and  $ W_{emb} \in  \mathbb{R}^{M \times d_{emb} }  $ is the trainable embedding matrix.
 In addition, we encode vector of timestamps $t=[t_1,t_2,...,t_L] $ to $Z=[z(t_1),z(t_2),...,z(t_L)]\in \mathbb{R}^{L \times d_{time}}$ using following transformation formula:
@@
 \end{equation} 
-Here, $\mathcal{T}$ is representing maximum time scale.
-This is very similar to positional encodings in transformers where the index is substitued by the time stamp.
-In contrary to THP and original positional encoding that considers $d_{emb}=d_{time}$ and adds up time encoding to the event embedding, we propose to concatenate this two vectors:
+Here, $\mathcal{T}$ is representing the maximum time scale.
+This transformation is very similar to positional encodings in transformers \cite*{vaswaniAttentionAllYou2017}, where the index is substituted by the timestamp.
+Contrary to THP and original positional encoding that considers $d_{emb}=d_{time}$ and adds up time encoding to the event embedding, we propose to concatenate these two vectors:
 
@@
 
-Finally, we use the standard transformer encoder to encode embedded events $X_{emb}$ tothe encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $.
-It is important to use approporiate mask matrix to prevent information leakage from future to the past.
-In this case, $h_j$ containes the all available information until occurence of $j$-th event.
+Finally, we use the standard transformer architecture to encode embedded events $X_{ev}$ into the encoded matrix  $ H=(h_1, ..., h_j, ..., h_L) $.
+It is important to use an appropriate mask matrix to prevent information leakage from the future to the past.
+In this case, $h_j$ should contain all available information until the occurrence of $j$-th event.
 
@@
 
-We use the a deep attention module \cite*{hornSetFunctionsTime2020} for encoding all additional information including irregularly sampled time series.
-Each observation of a sample can be represented by $u_k=(z(t_k), v_k, m_k)$ where $z(t_k)$ is the same time encoding similar to equation \ref{eq:time-encoding}.% we define attention $a(\mathcal{U}_k,s_k )$
+We use a deep attention module \cite*{hornSetFunctionsTime2020} for encoding all additional information including irregularly sampled time series.
+Each observation is represented by $u_p=(z(t_p), v_p, m_p)$ where $z(t_p)$ is the same transformation for time encoding in equation \ref{eq:time-encoding}.% we define attention $a(\mathcal{U}_k,s_k )$
 
 We define $\mathcal{U}_p$ to be the set of the first $p$ observations.
-The goal is to calculate the attention weight $a(\mathcal{U}_p,u_k ), k \leq p$ that is the relevance of $k$-th observation $s_k$ to the first $p$ observed values $\mathcal{U}_p$.
+The goal is to calculate the attention weight $a(\mathcal{U}_p,u_k ), k \leq p$ that is the relevance of $k$-th observation $u_k$ to the first $p$ observed values $\mathcal{U}_p$.
 This is achieved by computing an embedding of the set elements using a smaller set functions $f^{\prime}$, and projecting the concatenation of the set representation and the individual set element into d-dimensional space:
@@
 \begin{equation}\label{eq:fprime}
-    y^{\prime}_{p}=f^{\prime}(\mathcal{U}_p) = g^{\prime} \left(  \frac{1}{|p|} \sum_{u_k \in \mathcal{U}_p}  h^{\prime}(u_k;\theta^{\prime}) ;\rho^{\prime}\right) 
+   f^{\prime}(\mathcal{U}_p) = g^{\prime} \left(  \frac{1}{|p|} \sum_{u_k \in \mathcal{U}_p}  h^{\prime}(u_k;\theta^{\prime}) ;\rho^{\prime}\right) 
 \end{equation}
 
-Here, we compute the mean of first $p$ observations after passing first $p$ observations through a multilayer perceptron (MLP) nueral network ($h^{\prime}(u_k;\theta^{\prime})$).
+Here, we compute the mean of the first $p$ observations after passing the first $p$ observations through a multilayer perceptron (MLP) neural network ($h^{\prime}(u_k;\theta^{\prime})$).
 Finally, a second transformation using $g^{\prime}$ is performed to obtain embeddings $f^{\prime}(\mathcal{U}_p) \in \mathbb{R}^{d_{g^{\prime}}}$.
@@
 \begin{equation}
-    a(\mathcal{U}_p,u_p)=  softmax(\frac{K_p.w^q}{\sqrt{d}  })
+    a(\mathcal{U}_p,u_k)=  softmax(\frac{K_p.w^q}{\sqrt{d}  })
 \end{equation}
@@
 
-\subsection{All formulas}
+% \subsection{All formulas}
 
@@
 In neural point process literature, many approaches have been propose to decode either conditional or cumulative intensity function.
-We will use a decoder similar to [sahp] as it can model both exciting and inhibiting effects for modeling CIFs.
+We will use a decoder similar to \cite*{zhangSelfAttentiveHawkesProcess2020} as it can model both exciting and inhibiting effects for modeling CIFs.
 
@@
 
-We define a multi-objective loss function as $ \mathcal{L}   =   \mathcal{L}_{CIF}   +   \beta \mathcal{L}_{state}$., where $\mathcal{L}_{CIF}$ could be one of equations \ref*{eq:CIF-mc},\ref*{eq:CIF-ml} or \ref*{eq:CIF-marked}.
-$\mathcal{L}_{state}$ could be any desired loss functioin such as mean square error or corss entropy depending on the task.
-$\beta$ is a coefficient that can be optimized during hyperparamter tuning.
+We define a multi-objective loss function as $ \mathcal{L}   =   \mathcal{L}_{CIF}   +   \beta \mathcal{L}_{state}$., where $\mathcal{L}_{CIF}$ could be one of equations \ref*{eq:CIF-mc},\ref*{eq:CIF-ml} or \ref*{eq:CIF-marked}, and $\mathcal{L}_{state}$ could be any desired loss function such as mean square error or cross-entropy depending on the task.
+$\beta$ is a coefficient that can be optimized during hyperparameter tuning.
 
@@
 
-We perform various experiments to show the effectiveness of each component in out model.
+We perform various experiments to show the effectiveness of each component in our model.
 \subsection*{Datasets}
 
-\textbf{Synthea(\emph{Syn})}.
-We used the Synthea simulator (Walonoski et al., 2018) which generates patient-level EHRs using human expert curated Markov processes.
-Here, we reused the already processed version of this data by [ntpp].
+\textbf{Synthea(\emph{SYN})}.
+We used the Synthea simulator \cite*{walonoskiSyntheaApproachMethod2018} which generates patient-level EHRs using human expert-curated Markov processes.
+Here, we reused the already processed version of this data by \cite*{enguehardNeuralTemporalPoint2020}.
 
@@
 We record the time and the user tag of each tweet.
-Further, users are grouped into three categories based on the number of their followers: “small”, “medium”, and “large”
+Further, users are grouped into three categories based on the number of their followers: “small”, “medium”, and “large”.
+We use two versions of this dataset for multi-class \cite*{zhangSelfAttentiveHawkesProcess2020} and multi-label \cite*{enguehardNeuralTemporalPoint2020} scenario.
 
-\textbf{Stackoverflow (\emph{SO})}.
-is a question-answering website.
+\textbf{Stackoverflow (\emph{SO})} is a question-answering website.
 The website rewards users with badges to promote engagement in the community, and the same badge can be rewarded multiple times to the same user.
@@
 Each event in the sequence signifies receipt of a particular medal.
+We used the same processed dataset in the literature \cite*{zuoTransformerHawkesProcess2020a}.
 
-Furthermore, we consider two EHRs provided by physionet challenge to investigate the advantage of irregular sample and point process modeling in the same time.
+Furthermore, we consider two EHRs provided by physionet:
 
 \textbf{Physionet 2012 Mortality Prediction Challenge (\emph{P12})}.
-The 2012 Physionet challenge dataset (Goldberger et al., 2000),contains 12, 000 ICU stays each of which lasts at least 48 h.
+The 2012 Physionet challenge dataset \cite*{silvaPredictingInHospitalMortality},contains 12, 000 ICU stays each of which lasts at least 48 hours.
 For each stay, a set of general descriptors (such as gender or age) are collected at admission time.
@@
 \textbf{Physionet 2019 Sepsis Early Prediction Challenge (\emph{P19})}.
-This dataset contains clinical data of about 40k patients in ICU.
+This dataset contains clinical data of about 40,000 patients in ICU \cite*{reynaEarlyPredictionSepsis2020}.
 Clinical data consist of demographics, vital signs and laboratory values as well as sepsis label in a one-hour time grid.
-Our objective is to predict the timestamp of next lab sampling events as well as measured variables (event marks) given the patient history.
 
@@
 \begin{itemize}
-    \item AE(next mark) is a simple auto-encoder for predicting the next event from event embeddings without CIF characterization.
-As mentioned before, the log-likelihood of the point process will reduce to this one if we ignore the integral term.
-    \item PP(marked) is a marked point process (equation \ref*{eq:CIF-marked}) where we assume marks and time stamps are independent.
-    \item PP(multi-class or multi-label) uses multi-class or multi-label loss.
+    \item \emph{AE(next mark)} is a simple auto-encoder for predicting the next event from event embeddings without CIF characterization.
+The loss function is cross-entropy or binary cross-entropy for multi-class and multi-label datasets respectively.
+This scenario provides a baseline for other point process-based scenarios.
+    \item \emph{PP(marked)} is a marked point process (\ref*{eq:CIF-marked}) where we assume marks and time stamps are independent.
+    \item \emph{PP(multi-class or multi-label)} uses multi-class (\ref*{eq:CIF-mc}) or multi-label loss (\ref*{eq:CIF-ml}).
 \end{itemize}
@@
 In the second series, we want to investigate the effectiveness of state encoding in addition to the events for real-world EHR datasets P12 and P19 in a multi-label setting.
-We regard the occurrence of certain laboratory variables as events that are based on clinicians' decisions.
+We regard the occurrence of certain laboratory variables as events that are based on clinicians' decisions (see Appendix).
 As a result, we consider two scenarios:
@@
 \begin{itemize}
-    \item TEE: Here, we only use the transformer event encoder for CIFs characterization.
-    \item TEE+DAM: we further encode time stamps and values of all clinical variables through the deep attention module (DAM).
+    \item \emph{TEE}: Here, we only use the transformer event encoder (TEE) for CIFs characterization.
+    \item \emph{TEE+DAM}: We further encode time stamps and values of all clinical variables through the deep attention module (DAM).
+    \item \emph{TEE+noise}: To further investigate that the improved performance in the second scenario is not due to larger vector representation, we replace the DAM representations by a Gaussian noises vector ($\epsilon \sim \mathcal{N} (0,1)$).
 \end{itemize}
 
-In the last series of experiments, we investigate the utility of event encoding in a supervised learning task.
+In the last series of experiments, we investigate the utility of event encoding in a downstream supervised learning task.
 In particular, we try to predict mortality and sepsis shock as a binary outcome in P12 and P19 respectively.
 As a result, we compare (TEE+DAM) against the baseline (DAM) for outcome prediction.
+
+We further split healthcare datasets (P12 and P19) into three settings to compare for generalizability across different centers:
+
+\begin{itemize}
+    \item single-center (sc): Training data is from the same center as the test data.  
+    \item external evaluation (mc2): Training data is from all centers excluding the test data center:
+    \item multi-center (mc1): Training data is from all available center
+\end{itemize}
 
@@
 
+We use t-SNE (t-Distributed Stochastic Neighbor Embedding) for showing learned representations in the downstream tasks which is a machine learning algorithm used for visualizing high-dimensional data in a lower-dimensional space \cite*{maatenVisualizingDataUsing2008}.
+
+To show the similarity of events in the learned representation, we first compute the measurement density of each laboratory variable during patient stay:
+
+\begin{equation}
+    d_{m}^{i}=  \frac{1}{t_L} \sum_{j=1}^{L} \mathbbm{1}(e_j=m),
+\end{equation}
+
+For each desired patient, we can compute the mean of cosine similarity between the 10 nearest neighbors (in the embedding space) and the desired patients.
 \subsection*{Training Details}
@@
 
-In this section, we present our results regarding the advantage of state and event encoding.
+% In this section, we present our results regarding the advantage of state and event encoding.
 
@@
       \toprule
+      \toprule
             &       & \multicolumn{8}{c}{TEEDAM}                                    &       &       &       &  \\
@@
       \midrule
-      \multirow{2}[2]{*}{SO (MC)} & LL/\#events & ND    & ND    &       & -0.56 & -0.57 &       & -2.04 & -2.05 & -1.54 & -1.86 & -1.84 & NR \\
-            & F1-score & 38.46 & 36.91 &       & 36.04 & 34.91 &       & \multicolumn{1}{r}{32.65} & 31.67 & 28.34(0.19) & 24.12 & 23.89 & 26 \\
+      \multirow{2}[2]{*}{SO (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -0.56 & -0.57 &       & -2.04 & -2.05 & -1.54 & -1.86 & -1.84 & \textit{NR} \\
+            & F1-score & \textbf{38.46} & 36.91 &       & 36.04 & 34.91 &       & 32.65 & 31.67 & 28.34 & 24.12 & 23.89 & 26 \\
       \midrule
-      \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & ND    & ND    &       &       &       &       &       &       & -3.89 & -4.56 & -4.57 & NR \\
-            & F1-score & 62.48 & 61.65 &       & 58.45 & 35.59 &       & 36.38 &       & 58.29 & 53.92 & 53.86 & NR \\
+      \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -6.91 & -8.036 &       & -11   &       & -3.89 & -4.56 & -4.57 & \textit{NR} \\
+            & F1-score & \textbf{62.69} & 61.39 &       & 53.91 & 38.63 &       & 53    &       & 58.29 & 53.92 & 53.86 & \textit{NR} \\
       \midrule
-      \multirow{2}[2]{*}{Synthea (ML)} & LL/\#events & ND    & ND    &       &       &       &       &       &       & ND    & ND    & ND    & NR \\
-            & AUROC & 89.58 & 89.19 &       & 64.98 & 63.95 &       & 60.95 & 60.65 & ND    & ND    & ND    & 0.85(.014) \\
+      \multirow{2}[2]{*}{Synthea (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & -2.405 & -2.599 &       & -7.254 & -6.58 & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
+            & AUROC & \textbf{89.58} & 89.19 &       & 64.98 & 63.95 &       & 60.95 & 60.65 & \textit{ND} & \textit{ND} & \textit{ND} & 0.85(.014) \\
       \midrule
-      \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & ND    & ND    &       & 1.589 & 1.355 &       & -1.587 & -1.656 & ND    & ND    & ND    & NR \\
-            & AUROC & 69.12 & 67.86 &       & 62.5  & 60.4  &       & 73.9  & 71.8  & ND    & ND    & ND    & 0.611(0.001) \\
+      \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & 2.045 & 1.863 &       & -1.517 & -1.601 & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
+            & AUROC & 69.63 & 68.54 &       & 69.31 & 68.05 &       & \textbf{74.22} & 71.27 & \textit{ND} & \textit{ND} & \textit{ND} & 0.611(0.001) \\
+      \bottomrule
       \bottomrule
@@
 
-Table \ref{tab:1} shows the performance metrics of TEEDAM for different datasets in various scenarios.
-We can see that in most cases time concatenation leads to better results in terms of LL/\#events and F1-score/AUROC while adding time encodings to the event embeddings is the default practice in the literature \cite*{zhangSelfAttentiveHawkesProcess2020,zuoTransformerHawkesProcess2020a}.
+Table \ref{tab:1} shows the performance metrics of NPP4EHR for different datasets in various scenarios.
+We can see that in most cases time concatenation leads to better results in terms of LL/\#events and F1-score/AUROC.
+While adding time encodings to the event embeddings is the default practice in the point process literature \cite*{zhangSelfAttentiveHawkesProcess2020,zuoTransformerHawkesProcess2020a} as well as natural language processing, we can see that time concatenation would allow model to achieve better results.
 
@@
 These datasets have been widely used in the point process literature to show the effectiveness of point process, however, we show that this simple baseline may perform better for some datasets.
+We remark again that point process loss functions (\ref*{eq:CIF-mc},\ref*{eq:CIF-ml},\ref*{eq:CIF-marked}) will reduce to the loss function of AE(next mark) if we neglect the integral term.
 
-RT(MC) is the only dataset in which the point process with multi-label loss (equation \ref*{eq:CIF-ml}) performs better indicating the real advantage of modeling non-event likelihood of point process in this dataset.
-In general, we would recommend to compare future works in this area with this simple baseline.
+RT(MC) is the only dataset in which the point process with multi-label loss (equation \ref*{eq:CIF-ml}) performs better indicating the real advantage of modeling non-event likelihood of the point process in this dataset.
+For future research in the neural point process literature, we would recommend comparing model performance with this simple autoencoder model.
 
@@
       \toprule
+      \toprule
             &       &       & \multicolumn{2}{c}{F1} &       & \multicolumn{2}{c}{AUPRC} &       & \multicolumn{2}{c}{AUROC} \\
@@
       \midrule
-      \multirow{8}[6]{*}{P12} & \multirow{3}[2]{*}{sc} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc1} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-  \cmidrule{2-11}          & mc2   & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            & seft  & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+      \multirow{8}[6]{*}{P12} & \multirow{3}[2]{*}{mc1} & 1     & 0.53 (0.04) & 0.52 (0.03) &       & 0.53 (0.06) & 0.53 (0.04) &       & 0.85 (0.03) & 0.85 (0.03) \\
+            &       & 2     & 0.56 (0.02) & 0.55 (0.03) &       & 0.56 (0.05) & 0.56 (0.03) &       & 0.86 (0.01) & 0.87 (0.01) \\
+            &       & 3     & 0.55 (0.04) & \textbf{0.57 (0.03)} &       & 0.59 (0.05) & 0.58 (0.03) &       & 0.86 (0.02) & 0.87 (0.01) \\
+  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc2} & 1     & 0.52  & 0.52  &       & 0.53  & 0.52  &       & 0.85  & 0.85 \\
+            &       & 2     & 0.54  & 0.54  &       & 0.53  & \textbf{0.55} &       & 0.85  & 0.86 \\
+            &       & 3     & 0.52  & 0.54  &       & 0.56  & 0.56  &       & 0.85  & 0.86 \\
+  \cmidrule{2-11}          & sc    & -     & \textbf{0.49 (0.05)} & 0.46 (0.01) &       & \textbf{0.48 (0.05)} & 0.44 (0.02) &       & 0.82 (0.03) & 0.81 (0.02) \\
+            & seft  & -     & \textbf{0.52 (0.03)} & 0.5 (0.03) &       & \textbf{0.51 (0.05)} & 0.47 (0.03) &       & 0.83 (0.01) & 0.82 (0.01) \\
       \midrule
-      \multirow{8}[6]{*}{P19} & \multirow{3}[2]{*}{sc} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc1} & 1     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            &       & 2     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            &       & 3     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-  \cmidrule{2-11}          & mc2   & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
-            & seft  & -     & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) &       & 0.55 (0.02) & 0.55 (0.02) \\
+      \multirow{8}[6]{*}{P19} & \multirow{3}[2]{*}{mc1} & 1     & 0.53 (0.03) & 0.52 (0.04) &       & 0.51 (0.05) & \textbf{0.53 (0.05)} &       & 0.84 (0.03) & 0.84 (0.02) \\
+            &       & 2     & 0.51  & 0.51  &       & 0.51  & 0.51  &       & 0.85  & 0.85 \\
+            &       & 3     & 0.7 (0.02) & 0.69 (0.01) &       & 0.78 (0.02) & 0.78 (0.02) &       & 0.92 (0.02) & 0.93 (0.01) \\
+  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc2} & 1     & 0.62 (0.04) & 0.62 (0.03) &       & 0.7 (0.03) & 0.69 (0.03) &       & 0.94 (0.01) & 0.94 (0.01) \\
+            &       & 2     & 0.45  & \textbf{0.59} &       & 0.68  & 0.67  &       & 0.88  & 0.87 \\
+            &       & 3     & 0.62  & 0.61  &       & 0.53  & \textbf{0.55} &       & 0.92  & 0.91 \\
+  \cmidrule{2-11}          & sc    & -     & 0.68 (0.02) & 0.68 (0.03) &       & 0.77 (0.02) & 0.76 (0.03) &       & 0.93 (0.0) & 0.93 (0.0) \\
+            & seft  & -     & 0.61 (0.02) & 0.61 (0.01) &       & 0.68 (0.01) & 0.67 (0.01) &       & 0.94 (0.0) & 0.94 (0.0) \\
+      \bottomrule
       \bottomrule
@@
 
+% \begin{figure*}[!t]
+%     \centerline{\includesvg[inkscapelatex=false,width=2\columnwidth]{images/tsne.svg}}
+%     \caption{Magnetization as a function of applied field.
+%     It is good practice to explain the significance of the figure in the caption.}
+%     \label{fig1}
+%     \end{figure*}
+
 \begin{figure*}[!t]
-    \centerline{\includesvg[inkscapelatex=false]{images/tsne.svg}}
+    \centerline{\includegraphics{images/tsne.png}}
     \caption{Magnetization as a function of applied field.

### 2023-03-06 Manuscript_v1.tex
update

@@
 
-For each desired patient, we can compute the mean of cosine similarity between the 10 nearest neighbors (in the embedding space) and the desired patients.
+The mean of cosine similarities between the measurement density of a desired patient and its 10 nearest neighbors (in the embedding space) is computed (10-nnps).
 \subsection*{Training Details}
@@
       \midrule
-      \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -6.91 & -8.036 &       & -11   &       & -3.89 & -4.56 & -4.57 & \textit{NR} \\
-            & F1-score & \textbf{62.69} & 61.39 &       & 53.91 & 38.63 &       & 53    &       & 58.29 & 53.92 & 53.86 & \textit{NR} \\
+      \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -6.91 & -8.036 &       & -11   & -23.38 & -3.89 & -4.56 & -4.57 & \textit{NR} \\
+            & F1-score & \textbf{62.69} & 61.39 &       & 53.91 & 38.63 &       & 53    & 35.17 & 58.29 & 53.92 & 53.86 & \textit{NR} \\
       \midrule
@@
       \midrule
-      \multirow{8}[6]{*}{P12} & \multirow{3}[2]{*}{mc1} & 1     & 0.53 (0.04) & 0.52 (0.03) &       & 0.53 (0.06) & 0.53 (0.04) &       & 0.85 (0.03) & 0.85 (0.03) \\
+      \multirow{10}[8]{*}{P12} & \multirow{3}[2]{*}{mc1} & 1     & 0.53 (0.04) & 0.52 (0.03) &       & 0.53 (0.06) & 0.53 (0.04) &       & 0.85 (0.03) & 0.85 (0.03) \\
             &       & 2     & 0.56 (0.02) & 0.55 (0.03) &       & 0.56 (0.05) & 0.56 (0.03) &       & 0.86 (0.01) & 0.87 (0.01) \\
@@
             &       & 3     & 0.52  & 0.54  &       & 0.56  & 0.56  &       & 0.85  & 0.86 \\
-  \cmidrule{2-11}          & sc    & -     & \textbf{0.49 (0.05)} & 0.46 (0.01) &       & \textbf{0.48 (0.05)} & 0.44 (0.02) &       & 0.82 (0.03) & 0.81 (0.02) \\
-            & seft  & -     & \textbf{0.52 (0.03)} & 0.5 (0.03) &       & \textbf{0.51 (0.05)} & 0.47 (0.03) &       & 0.83 (0.01) & 0.82 (0.01) \\
+  \cmidrule{2-11}          & \multirow{3}[2]{*}{sc} & 1     & \textbf{0.49 (0.05)} & 0.46 (0.01) &       & \textbf{0.48 (0.05)} & 0.44 (0.02) &       & 0.82 (0.03) & 0.81 (0.02) \\
+            &       & 2     & \textbf{0.52 (0.03)} & 0.5 (0.03) &       & \textbf{0.51 (0.05)} & 0.47 (0.03) &       & 0.83 (0.01) & 0.82 (0.01) \\
+            &       & 3     & 0.53 (0.03) & 0.52 (0.04) &       & 0.51 (0.05) & \textbf{0.53 (0.05)} &       & 0.84 (0.03) & 0.84 (0.02) \\
+  \cmidrule{2-11}          & seft  & -     & 0.51  & 0.51  &       & 0.51  & 0.51  &       & 0.85  & 0.85 \\
       \midrule
-      \multirow{8}[6]{*}{P19} & \multirow{3}[2]{*}{mc1} & 1     & 0.53 (0.03) & 0.52 (0.04) &       & 0.51 (0.05) & \textbf{0.53 (0.05)} &       & 0.84 (0.03) & 0.84 (0.02) \\
-            &       & 2     & 0.51  & 0.51  &       & 0.51  & 0.51  &       & 0.85  & 0.85 \\
-            &       & 3     & 0.7 (0.02) & 0.69 (0.01) &       & 0.78 (0.02) & 0.78 (0.02) &       & 0.92 (0.02) & 0.93 (0.01) \\
-  \cmidrule{2-11}          & \multirow{3}[2]{*}{mc2} & 1     & 0.62 (0.04) & 0.62 (0.03) &       & 0.7 (0.03) & 0.69 (0.03) &       & 0.94 (0.01) & 0.94 (0.01) \\
-            &       & 2     & 0.45  & \textbf{0.59} &       & 0.68  & 0.67  &       & 0.88  & 0.87 \\
-            &       & 3     & 0.62  & 0.61  &       & 0.53  & \textbf{0.55} &       & 0.92  & 0.91 \\
-  \cmidrule{2-11}          & sc    & -     & 0.68 (0.02) & 0.68 (0.03) &       & 0.77 (0.02) & 0.76 (0.03) &       & 0.93 (0.0) & 0.93 (0.0) \\
-            & seft  & -     & 0.61 (0.02) & 0.61 (0.01) &       & 0.68 (0.01) & 0.67 (0.01) &       & 0.94 (0.0) & 0.94 (0.0) \\
+      \multirow{6}[6]{*}{P19} & \multirow{2}[2]{*}{mc1} & 1     & 0.7 (0.02) & 0.69 (0.01) &       & 0.78 (0.02) & \textbf{0.78 (0.02)} &       & 0.92 (0.02) & 0.93 (0.01) \\
+            &       & 2     & 0.62 (0.04) & 0.62 (0.03) &       & 0.7 (0.03) & 0.69 (0.03) &       & 0.94 (0.01) & 0.94 (0.01) \\
+  \cmidrule{2-11}          & \multirow{2}[2]{*}{mc2} & 1     & 0.45  & \textbf{0.59} &       & 0.68  & 0.67  &       & 0.88  & 0.87 \\
+            &       & 2     & 0.62  & 0.61  &       & 0.53  & \textbf{0.55} &       & 0.92  & 0.91 \\
+  \cmidrule{2-11}          & \multirow{2}[2]{*}{sc} & 1     & 0.68 (0.02) & 0.68 (0.03) &       & 0.77 (0.02) & 0.76 (0.03) &       & 0.93 (0.0) & 0.93 (0.0) \\
+            &       & 2     & 0.61 (0.02) & 0.61 (0.01) &       & 0.68 (0.01) & 0.67 (0.01) &       & 0.94 (0.0) & 0.94 (0.0) \\
       \bottomrule
@@
 As can be seen,  in TEEDAM neighbors have more similar pattern compared to the baseline.
-Although event encoding does not necessarily leads to better performance, we can learn a better representation that considers both absolute values and the patterns.
+Although event encoding does not necessarily leads to better performance, we can learn a better representation that considers both absolute values and patterns.
+The 10 nearest neighbor pattern similarity (10-nnps) of positive labels for TEE+DAM ($0.903 \pm 0.058 $) is signifcantly higher than (DAM) ($0.832 \pm 0.095$)($pvalue \leq 0.001$).
 One application could be time series generation.

### 2023-03-08 Manuscript_v1.tex
update

@@
 
+% Table generated by Excel2LaTeX from sheet 'datasets'
+\begin{table}[htbp]
+    \centering
+    \caption{Add caption}
+      \begin{tabular}{ccccc}
+      \toprule
+      \toprule
+      Dataset & Task  & \# classes & \# samples & Avg. length \\
+      \midrule
+      Stack Overflow & Multi-class & 22    & 6633  & 72 \\
+      ReTweets & Multi-class & 3     & 24000 & 109 \\
+      ReTweets & Multi-label & 3     & 20000 & 104 \\
+      Synthea & Multi-label & 357   & 12000 & 43 \\
+      Physionet 2012 & Multi-label & 25    & 12000 & 21 \\
+      Physionet 2019 & Multi-label & 25    & 12000 & 21 \\
+      \bottomrule
+      \bottomrule
+      \end{tabular}%
+    \label{tab:data}%
+
+  \end{table}%
+
 \textbf{Synthea(\emph{SYN})}.
@@
 % Table generated by Excel2LaTeX from sheet 'time concat'
+% Table generated by Excel2LaTeX from sheet 'time concat'
 \begin{table*}[htbp]
@@
     \caption{Add caption}
-      \begin{tabular}{cccccccccccccc}
+      \begin{tabular}{ccccccccccccccc}
       \toprule
       \toprule
-            &       & \multicolumn{8}{c}{TEEDAM}                                    &       &       &       &  \\
-  \cmidrule{3-10}          &       & \multicolumn{2}{c}{AE (next mark)} &       & \multicolumn{2}{c}{PP(single+mark)} &       & \multicolumn{2}{c}{PP (MC/ML)} &       &       &       &  \\
-  \cmidrule{3-4}\cmidrule{6-7}\cmidrule{9-10}    Dataset & Metric & concat & sum   &       & concat & sum   &       & concat & sum   & Latent & SAHP  & THP   & GRU-CP \\
+            &       & \multicolumn{8}{c}{TEEDAM}                                    &       &       &       &       &  \\
+  \cmidrule{3-10}          &       & \multicolumn{2}{c}{AE (next-mark)} &       & \multicolumn{2}{c}{PP(single+mark)} &       & \multicolumn{2}{c}{PP (MC/ML)} &       & \multicolumn{4}{c}{Baselines} \\
+  \cmidrule{3-4}\cmidrule{6-7}\cmidrule{9-10}\cmidrule{12-15}    Dataset & Metric & concat & sum   &       & concat & sum   &       & concat & sum   &       & Latent & SAHP  & THP   & GRU-CP \\
       \midrule
-      \multirow{2}[2]{*}{SO (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -0.56 & -0.57 &       & -2.04 & -2.05 & -1.54 & -1.86 & -1.84 & \textit{NR} \\
-            & F1-score & \textbf{38.46} & 36.91 &       & 36.04 & 34.91 &       & 32.65 & 31.67 & 28.34 & 24.12 & 23.89 & 26 \\
+      \multirow{2}[2]{*}{SO (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -0.56 & -0.57 &       & -2.04 & -2.05 &       & -1.54 & -1.86 & -1.84 & \textit{NR} \\
+            & F1-score & \textbf{38.46} & 36.91 &       & 36.04 & 34.91 &       & 32.65 & 31.67 &       & 28.34 & 24.12 & 23.89 & 26 \\
       \midrule
-      \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -6.91 & -8.036 &       & -11   & -23.38 & -3.89 & -4.56 & -4.57 & \textit{NR} \\
-            & F1-score & \textbf{62.69} & 61.39 &       & 53.91 & 38.63 &       & 53    & 35.17 & 58.29 & 53.92 & 53.86 & \textit{NR} \\
+      \multirow{2}[2]{*}{ReTweet (MC)} & LL/\#events & \textit{ND} & \textit{ND} &       & -6.91 & -8.036 &       & -11   & -23.38 &       & -3.89 & -4.56 & -4.57 & \textit{NR} \\
+            & F1-score & \textbf{62.69} & 61.39 &       & 53.91 & 38.63 &       & 53    & 35.17 &       & 58.29 & 53.92 & 53.86 & \textit{NR} \\
       \midrule
-      \multirow{2}[2]{*}{Synthea (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & -2.405 & -2.599 &       & -7.254 & -6.58 & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
-            & AUROC & \textbf{89.58} & 89.19 &       & 64.98 & 63.95 &       & 60.95 & 60.65 & \textit{ND} & \textit{ND} & \textit{ND} & 0.85(.014) \\
+      \multirow{2}[2]{*}{Synthea (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & -2.405 & -2.599 &       & -7.254 & -6.58 &       & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
+            & AUROC & \textbf{89.58} & 89.19 &       & 64.98 & 63.95 &       & 60.95 & 60.65 &       & \textit{ND} & \textit{ND} & \textit{ND} & 0.85(.014) \\
       \midrule
-      \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & 2.045 & 1.863 &       & -1.517 & -1.601 & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
-            & AUROC & 69.63 & 68.54 &       & 69.31 & 68.05 &       & \textbf{74.22} & 71.27 & \textit{ND} & \textit{ND} & \textit{ND} & 0.611(0.001) \\
+      \multirow{2}[2]{*}{ReTweet (ML)} & LL/\#events & \textit{ND} & \textit{ND} &       & 2.045 & 1.863 &       & -1.517 & -1.601 &       & \textit{ND} & \textit{ND} & \textit{ND} & \textit{NR} \\
+            & AUROC & 69.63 & 68.54 &       & 69.31 & 68.05 &       & \textbf{74.22} & 71.27 &       & \textit{ND} & \textit{ND} & \textit{ND} & 0.611(0.001) \\
       \bottomrule
@@
 We can see that in most cases time concatenation leads to better results in terms of LL/\#events and F1-score/AUROC.
-While adding time encodings to the event embeddings is the default practice in the point process literature \cite*{zhangSelfAttentiveHawkesProcess2020,zuoTransformerHawkesProcess2020a} as well as natural language processing, we can see that time concatenation would allow model to achieve better results.
+While adding time encodings to the event embeddings is the default practice in the point process literature \cite*{zhangSelfAttentiveHawkesProcess2020,zuoTransformerHawkesProcess2020a} as well as natural language processing, we can see that time concatenation would allow our model to achieve better results in terms of log-likelihood and next event type prediction.
 
-Another interesting fact is that the simple auto-encoder for next event type prediction achieves better results for SO,RT(MC) and SYN(ML) compared to the point process scenario.
-These datasets have been widely used in the point process literature to show the effectiveness of point process, however, we show that this simple baseline may perform better for some datasets.
-We remark again that point process loss functions (\ref*{eq:CIF-mc},\ref*{eq:CIF-ml},\ref*{eq:CIF-marked}) will reduce to the loss function of AE(next mark) if we neglect the integral term.
+Another interesting fact is that the simple auto-encoder for next event type prediction (AE(next-mark)) achieves better results for SO, RT(MC) and SYN(ML) compared to other point process models.
+These datasets have been widely used in the point process literature to show the effectiveness of NTPP models, however, we show that this simple baseline may perform better for some datasets.
+We remark again that point process loss functions (\ref*{eq:CIF-mc},\ref*{eq:CIF-ml},\ref*{eq:CIF-marked}) will reduce to the loss function of AE(next-mark) if we neglect the integral term.
 
-RT(MC) is the only dataset in which the point process with multi-label loss (equation \ref*{eq:CIF-ml}) performs better indicating the real advantage of modeling non-event likelihood of the point process in this dataset.
-For future research in the neural point process literature, we would recommend comparing model performance with this simple autoencoder model.
+RT(MC) is the only dataset in which the point process with multi-label loss (\ref*{eq:CIF-ml}) performs better indicating the real advantage of modeling the non-event likelihood of the point process in this dataset.
+For future research in the NTPP literature, we would recommend AE(next-mark) as a simple baseline for NTPP architectures.
 
@@
 
-Table \ref*{tab:2} shows the result for the utility of state encodings for the estimation of negative likelihood.
+% Table \ref*{tab:2} shows the results for the utility of state encodings for the estimation of negative likelihood.
 We see that in all cases, state encoding leads to higher AUROC for the next event type prediction and LL/\#event as well.
-It is intuitive that in a hospital, the absolute value of patient states can be useful for ordering future laboratory events.
-We have evaluated our model in a healthcare database, however, it can be further evaluated on other event sequence data where additional information is available.
+The noisy model (TEE+noise) does not improve performance reflecting the fact that the observed performance gain comes from unique information from patient's states.
+
+The results for the utility of state encodings for the estimation of negative likelihood are presented in Table \ref*{tab:2}.
+Our analysis indicates that state encoding consistently results in higher AUROC for predicting the next event type and a higher LL/\#event ratio in all cases.
+Notably, incorporating noise into the model (TEE+noise) does not yield any improvements in performance.
+This finding suggests that the observed performance gain is attributable to unique information derived from patient states, rather than extraneous factors.
+
+% It is also intuitive that in a hospital, the absolute value of patient states can be useful for ordering future laboratory events.
+We have evaluated our model on a healthcare database, however, it can be further evaluated on other event sequence data where additional information is available.
+
+It is also intuitive that in a hospital setting, the absolute value of patient states could prove useful in determining the order of future laboratory events.
+While we have evaluated our model using a healthcare database, we believe that further assessment could be conducted on other event sequence data that includes additional information.
 
@@
 
-Table \ref*{tab:3} indicates the result for Mortality/sepsis prediction task across different settings and hospital centers.
-Generally, we can see that in the mc1 setting, has the best performance because of the larger training dataset from other centers and the test center.
+Table \ref*{tab:3} indicates the result for the mortality/sepsis prediction task across different settings and hospital centers.
+Generally, we can see that the mc1 setting has the best performance because of the larger training dataset from other centers and the test center.
 In addition, event encoding module is useful for 2/5 cases while in 1/5 cases degrades the performance.
@@
     It is good practice to explain the significance of the figure in the caption.}
-    \label{fig1}
+    \label{fig2}
     \end{figure*}
 
-Fig 1 visualizes the t-SNE plot for the learned embeddings of [] with and without event encodings.
-Besides, we visualize the sampling pattern of an example patient and its neighbors.
-As can be seen,  in TEEDAM neighbors have more similar pattern compared to the baseline.
-Although event encoding does not necessarily leads to better performance, we can learn a better representation that considers both absolute values and patterns.
-The 10 nearest neighbor pattern similarity (10-nnps) of positive labels for TEE+DAM ($0.903 \pm 0.058 $) is signifcantly higher than (DAM) ($0.832 \pm 0.095$)($pvalue \leq 0.001$).
-One application could be time series generation.
+We have visualized the learned patients' embeddings for an example dataset and setting (P12-seft).
+Fig.
+1-a and b visualize the t-SNE plot for the learned representations of TEE+DAM and TEE, respectively.
+
+The 10 nearest neighbor pattern similarity (10-nnps) of positive labels for TEE+DAM ($0.903 \pm 0.058 $) is significantly higher than (DAM) ($0.832 \pm 0.095$)($pvalue \leq 0.001$).
+The sampling pattern of an example patient and its 4 nearest neighbors (in the embedding space) is visualized in Fig.
+\ref*{fig2}-c and d for TEE+DAM and TEE, respectively.
+As can be seen, in TEE+DAM, the 4 neighbors have a more similar pattern compared to the example patient.
+
+We argue that event encoding could lead to better representation learning in EHRs although it does not necessarily leads to better performance.
+Patient representation that can respect the irregular sampling pattern in EHRs could have various applications in synthetic data generation.
 

### 2023-03-10 Manuscript_v1.tex
update
Updates from Overleaf

@@
 The choice of imputation technique depends on the type and amount of missing data, as well as the goals of the analysis.
-Regardless of the technique used, it is important to carefully consider the impact of imputed values on the analysis and to report the imputation method used in the results.
+Regardless of the technique used, it is important to carefully consider the impact of imputed values on the analysis and to report the imputation method used in the results.https://www.overleaf.com/project/63c00678bca7836b32514494
 
@@
 We argue that event encoding could lead to better representation learning in EHRs although it does not necessarily leads to better performance.
-Patient representation that can respect the irregular sampling pattern in EHRs could have various applications in synthetic data generation.
+Patient representation that can respect the irregular sampling pattern in EHRs could have various applications in synthetic data generation..
 
